### 2024-04-29 docs/papers.tex
Set up research workspace.

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{tcolorbox}
\usepackage{ifthen}

\tcbuselibrary{skins,breakable}

\title{Journal Article Summaries}
\author{Your Name}
\date{\today}

\colorlet{ArticleBoxBackgroundColorOdd}{white}
\colorlet{ArticleBoxBackgroundColorEven}{gray!10}

\newbool{isEven}
\setbool{isEven}{false}

\newtcolorbox{ArticleColorBox}[1]{%
    breakable, 
    enhanced, 
    colback=\ifbool{isEven}{ArticleBoxBackgroundColorEven}{ArticleBoxBackgroundColorOdd}, 
    colframe=black, 
    fonttitle=\bfseries, 
    title=Article #1
}

\newcounter{ArticleSummaryCounter}
\setcounter{ArticleSummaryCounter}{0}

\newenvironment{ArticleSummary}[1]{
    \refstepcounter{ArticleSummaryCounter} % Increment the counter
    \begin{tcolorbox}[breakable, enhanced, colback=\ifbool{isEven}{ArticleBoxBackgroundColorEven}{ArticleBoxBackgroundColorOdd}, colframe=black, fonttitle=\bfseries, title=Summary \theArticleSummaryCounter: #1]
    \setlength\parskip{0.2cm}
}{
    \end{tcolorbox}
    \vspace*{0.5cm}
    \ifbool{isEven}{\setbool{isEven}{false}}{\setbool{isEven}{true}}
}

\begin{document}

\maketitle

\begin{ArticleSummary}{dfdsf}
\textbf{Citation:} sdfsd

\textbf{Summary:} sdfasf

\textbf{Key Points:}
\begin{itemize}
    \item : sdfs
    \item : 
\end{itemize}

\end{ArticleSummary}
\end{document}

### 2024-06-09 docs/papers.tex
Read more papers

@@
 \documentclass{article}
+\usepackage[margin=1in]{geometry}
 \usepackage[utf8]{inputenc}
-\usepackage{amsmath,amsfonts,amssymb}
+\usepackage{amsmath,amsfonts,amssymb, bm}
 \usepackage{graphicx}
+\usepackage[dvipsnames]{xcolor}
 \usepackage{hyperref}
@@
 \usepackage{ifthen}
+\usepackage{etoolbox}
+\usepackage{physics}
+\input{commands}
 
@@
 \title{Journal Article Summaries}
-\author{Your Name}
+\author{Jeffrey Liang}
 \date{\today}
@@
 \colorlet{ArticleBoxBackgroundColorOdd}{white}
-\colorlet{ArticleBoxBackgroundColorEven}{gray!10}
+\colorlet{ArticleBoxBackgroundColorEven}{BlueGreen!4}
 
 \newbool{isEven}
-\setbool{isEven}{false}
-
-\newtcolorbox{ArticleColorBox}[1]{%
-    breakable, 
-    enhanced, 
-    colback=\ifbool{isEven}{ArticleBoxBackgroundColorEven}{ArticleBoxBackgroundColorOdd}, 
-    colframe=black, 
-    fonttitle=\bfseries, 
-    title=Article #1
-}
+\setbool{isEven}{false}  % Start with true to begin with the first condition
 
@@
 
+\newtcolorbox{ArticleSummaryBox}[2][]{%
+    breakable, 
+    enhanced, 
+    colback=\ifbool{isEven}{ArticleBoxBackgroundColorEven}{ArticleBoxBackgroundColorOdd}, 
+    colframe=BlueGreen, 
+    fonttitle=\bfseries, 
+    title=Summary \theArticleSummaryCounter: #2,
+    #1 % for additional optional parameters
+}
+
 \newenvironment{ArticleSummary}[1]{
     \refstepcounter{ArticleSummaryCounter} % Increment the counter
-    \begin{tcolorbox}[breakable, enhanced, colback=\ifbool{isEven}{ArticleBoxBackgroundColorEven}{ArticleBoxBackgroundColorOdd}, colframe=black, fonttitle=\bfseries, title=Summary \theArticleSummaryCounter: #1]
-    \setlength\parskip{0.2cm}
+    \ArticleSummaryBox{#1}
+    % \setlength\parskip{0.2cm}
 }{
-    \end{tcolorbox}
+    \endtcolorbox % This ends the 'tcolorbox' defined in 'ArticleSummaryBox'
     \vspace*{0.5cm}
-    \ifbool{isEven}{\setbool{isEven}{false}}{\setbool{isEven}{true}}
+    \ifbool{isEven}{\global\boolfalse{isEven}}{\global\booltrue{isEven}}
 }
@@
 
-\begin{ArticleSummary}{dfdsf}
-\textbf{Citation:} sdfsd
+\begin{ArticleSummary}{MeshSDF: Differentiable Iso-Surface Extraction}
+    \textbf{Citation:} Remelli \textit{et al.}, 2020
 
-\textbf{Summary:} sdfasf
+    \textbf{Summary:} Using a neural network to model a signed distance function, we sample points from the zero-level set of the watertight shape and then use Marching Cubes (MC) to create a mesh between points.
+Rather than differentiating with respect to the MC mesh, we differentiate directly with the neural network since the loss function only cares about zero-level set.
+There do exist differentiable MC variants but they can be complex/apply only in certain conditions.
 
@@
     \begin{itemize}
-    \item : sdfs
-    \item : 
+        \item Vertices are sampled from zero level set usually, so if the loss function only cares about the zero level set then you can directly differentiate with respect to the neural net.
+When the loss function cares about the facets, then we need a differentiable Marching Cubes (but the points are still sampled from the zero level set).
+        \item When a point $s$ undergoes an infinitesimal perturbation $\Delta s$, the local surface is perturbed in the opposite direction of the surface normal: $\pdv{v}{s} v = -n(v) = -\nabla s(v)$.
+        \item a negative $\Delta s$ causes points with slightly positive distance values move closer to the zero set (as you subtract), that is, the surface is inflated.
+        \item Marching Cubes (MC) converts implicit functions to 3D surface meshes $\mathcal M=(V,F)$.
+It samples the network on a discrete 3D grid, detecting zero-crossing of the field along grid edges and builds a surface mesh with a lookup table.
+The position of vertices on grid edges involves linear interpolation, which is modelled by a function discontinuous at $s_i = s_j$, so cannot allow topology changes through backpropagation.
+
+        \item This method is used for tasks where we deform a 3D mesh $\mathcal{M} = (V,F)$, where $V = \lrbrace{\bm v_1, \bm v_2,...}$ denotes vertices in $\R^3$ and $F$ facets to minimise a task-specific loss function $\mathcal L_{\text {task}}(\mathcal M)$.
+        \item Signed distance function (SDF): $s: \R^3 \to \R$. Neural network $f_\theta$ is tarined on watertight surfaces $\mathcal S$ to approximate SDF $s$ by minimising
+        \begin{equation*}
+            \mathcal{L}_{\text{sdf}}\lr(){\lrbrace{\bm z_S}_{S\in \mathcal S}, \theta}
+                = \sum_{S\in \mathcal S} \frac{1}{|X_S|}\sum_{\bm x \in X_S} \lr||{f_\theta(\bm x, \bm z_S) - s(\bm x)} + \lambda_{\text{reg}}\sum_{S\in \mathcal S} \norm*{\bm z_S}^2_2
+        \end{equation*}
+        where $\bm z_S \in \R^Z$ is a $Z$-dimensional encoding of surface $S$, $\theta$ denotes network parameters, $X_S$ represents 3D point samples we use to train our netowrk, and $\lambda_{\text{reg}}$ affects regularisation.
+
+        When given a mesh, we want to be able to evaluate:
+        \begin{equation*}
+            \pdv{\mathcal L_{\text{task}}}{\bm z} = \sum_{\bm v \in V} \pdv{\mathcal L_{\text{task}}}{\bm v} \pdv{\bm v}{f_\theta} \pdv{f_\theta} {\bm z}.
+        \end{equation*}
+        As $f_\theta$ approximates an SDF, we can replace $\pdv{\bm v}{f_\theta}$ with $- \nabla f_\theta(\bm v, \bm z)$.
+        \item Two loss functions are considered between surfaces $S$ and $T$:
+        \begin{align*}
+            \mathcal L_{\text{task1}} &= \min_{s \in S} d(s,T) + \min_{t\in T} d(S,t), \tag*{surface-to-surface distance}\\
+            \mathcal L_{\text{task2}} &= \norm*{\text{DR}(S) - \text{DR}(T)}_1 \tag*{image-to-image distance}
+        \end{align*}
+
+        \item For shape optimisation, we can use the following loss function (aerodynamics):
+        \begin{equation*}
+            \mathcal L_{\text{task}}(\mathcal M) = \iint_{\mathcal M} g_\beta \bm n_x \, \text d\mathcal{M} + \mathcal L_{\text{constraint}}(\mathcal M).
+        \end{equation*}
     \end{itemize}
+\end{ArticleSummary}
 
+\begin{ArticleSummary}{Structure-from-Motion Revisited}
+    \textbf{Citation:} Schönberger and Frahm, 2016
+
+    \textbf{Summary:} A new (incremental) SfM technique which looks to address image registration and triangulation more efficiently and robustly.
+It augments scene graph using geometric verification strategy.
+Next best view selection for incremental reconstruction process.
+Robust triangulation to produce more complete scene structure with reduced computational cost.
+An iterative BA, re-triangulation, and outlier filtering strategy to improve completeness.
+More efficient BA parametrisation for dense photo collections.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item SfM starts with correspondence search given a set of images $\mathcal{I} = \lrbrace{I_i \mid i=1,...,N_I}$:
+        \begin{enumerate}
+            \item Feature extraction: $\mathcal F_i = \lrbrace{(x_j, \bm f_j) \mid j=1,...,N_{F_i}}$ at location $\bm x_j\in\R^2$
+            \item Matching ($\mathcal{O}(N_I^2N_{F_i}^2)$): find most similar feature in the other image through brute force to give set of potentially matching image pairs $\mathcal C = \lrbrace{\lrbrace{I_a,I_b} \mid I_a, I_b \in \mathcal I, a<b}$ and associated feature correspondences $\mathcal{M}_{ab} \in \mathcal F_a \times \mathcal F_b$.
+            \item Geometric verification: estimate projection transformation to map between feature points (use homography or epipolar geometry) + RANSAC.
+Outputs a scene-graph, with images as nodes and verified pairs of images as edges.
+        \end{enumerate}
+        \item Incremental reconstruction: outputs pose estimates $\mathcal P = \lrbrace{\bm P_c \in \bm{SE}(3) \mid c = 1,...,N_P}$ for registered images and the reconstructed scene structure as a set of points $\mathcal X = \lrbrace{\bm X_k \in \R^3 \mid k=1,...,N_X}$.
+        \begin{enumerate}
+            \item Initialisation: done with carefully selected two-view reconstruction - can be from densse location in the image graph with many overlapping cameras. 
+            \item Image registration: can register to current model by solving Perspective-n-Point (PnP) problem suing using correspondences to triangulated points already in the image.
+Each new registered image extends the set $\mathcal P$.
+This paper attempts to improve in this part.
+            \item Triangulation: a new scene point can be triangulated and added to $\mathcal X$ as soon as one more image also covering the same scene but from a different viewpoint is registered.
+The redundancy and additional correspondences increases stability.
+This paper addresses more robust/efficient method.
+            \item Bundle adjustment: A joint non-linear refinement of camera parameters $\bm P_c$ and point parameters $\bm X_k$ which minimise the reprojection error $E = \sum_{j} \rho_j \lr(){\norm*{\pi (\bm P_c, \bm X_k) - \bm x_j}_2^2}$, where $\pi$ projects scene points to the image space and $\rho_j$ is a loss function to down-weight outliers.
+Levenberg-Marquardt is used to solve BA problems.
+Can use Schur complement trick.
+This paper attempts to improve on this part.
+        \end{enumerate}
+        \item Current challenges for SfM: failure to register images which should be, produce broken models due to mis-registrations or drift.
+Caused by incomplete scene graphs, or in failures in reconstruction stage due to inaccurate scene structure.
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{CAPNet: Continuous Approximation Projection for 3D Point Cloud Reconstruction Using 2D Supervision}
+    \textbf{Citation:} Navaneet \textit{et al.}, 2019
+
+    \textbf{Summary:} Constructs a 3D point cloud by projecting them into multiple 2D views and comparing with ground-truth 2D masks in each view (weak supervision).
+A continuous approximation of points in the point cloud produces smooth projections in a differentiable manner.
+Uses encoder-decoder architecture.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Point cloud representations are better than volumetric representations since the latter suffers from information sparsity.
+Surface voxels provide structural information, however internal voxels increase the computational complexity with minimal addition to information.
+        \item \textit{Goal}: From a single image of an object, reconstruct a 3D point cloud representation of the object.
+If $I$ is an image from the training set and $f$ is a trained network, let $p=f(I)$ be the corresponding 3D point cloud reconstruction.
+A projection $P(p,v)$ is from an arbitrary view point $v$ is obtained by performing a perspective transformation and projecting the transformed point cloud onto a plane.
+The transformed point $\hat{p}_n=(\hat x_n, \hat y_n, \hat z_n)$ in camera coordinates is obtained from: $\hat{p}_n = K(R_v p_n + t_v)\quad  \forall n \in \lrbrace{1,...,N}$, where $K, R_v, t_n$ are the camera intrinsics and extrinsics.
+Training uses ground truth 2D masks $M$ to supervise projection $\hat{M} = P(p,v).$
+        \item The encoder takes in a 2D image and the decoder reconstructs the point cloud.
+To compute the loss, the predicted point cloud is projected from $V$ different view-points and compared with corresponding ground truth projections.
+        \item The final loss function is $\mathcal L = \mathcal L_{bce} + \lambda \cdot \mathcal L_{\text{aff}}$.
+Let $\hat{M}^v_{i,j} = \tanh\lr(){\sum_{n=1}^N \phi(\hat{x}_n-i)\cdot\phi(\hat{y}_n-j)}$, where $\phi$ is the kernel function $\phi(k) = \exp\lr(){\frac{-k^2}{2\sigma^2}}$ (this Gaussian kernel allows smooth, accurate projections with no holes).
+The loss function is binary cross-entropy loss:
+        $$\mathcal{L}_{bce} = \sum_{v=1}^V -M^v \log (\hat M^v) - (1-M^v)\log(1-\hat M^v)$$
+        where $M^v$ and $\hat M^v$ are the ground truth and predicted masks, respectively, of dimension $(H,W)$.
+This loss function results in reconstructions with many outlier points, so a \textit{nearest point affinity loss} is imposed which minimises the nearest neighbour distance between two pixel maps weighted by pixel confidence:
+        $$\mathcal{L}_{\text {aff}} = \sum_{v=1}^V\sum_{i,j}^{H,W} \min_{(k,l)\in M_+^v} ((i-k)^2 +(j-l)^2)\hat M^v_{i,j} M^v_{k,l} + \sum_{v=1}^V\sum_{i,j}^{H,W} \min_{(k,l)\in \hat M_+^v} ((i-k)^2 +(j-l)^2) M^v_{i,j} \hat M^v_{k,l}$$
+        where $M_+^v$ and $\hat M_+^v$ are sets of pixel coordinates of the ground truth and predicted projections whose values are non-zero.
+        \item The resultant point cloud is evaluated with the ground truth by computing the Chamfer distance:
+        $$d_{\text{Chamfer}}(\hat{P}, P) = \sum_{x\in \hat P} = \min_{y\in P} \norm{x-y}_2^2 + \sum_{y\in P} = \min_{x \in \hat P} \norm{y-x}_2^2.$$ Ground truths were obtained by randomly sampling 16,384 points on the object surface and then performing farthest point sampling to obtain 1024 points.
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{Occupancy Networks: Learning 3D Reconstruction in Function Space}
+    \textbf{Citation:} Mescheder \textit{et al.}
+
+    \textbf{Summary:} Uses a neural network classifier to represent the continuous decision boundary of a 3D surface.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Existing 3D representations can be broadly categorised into three categories: voxel-based representations, point-based representations, and mesh representations.
+Point clouds lack connectivity structure of underlying mesh and hence require extra post-processing steps to extract 3D geometry.
+        \item Let the occupancy function $o: \R^3 \to \{0,1\}$ denote whether that point is occupied by the object.
+The paper's occupancy function outputs a probability between 0 and 1 for every point $p\in \R^3$.
+We want to condition the reconstructed 3D object output on the input $x \in \mathcal X$.
+Hence, we train an \textit{occupancy network} $f_\theta: \R^3 \times \mathcal X \to [0,1]$, which takes a pair $(p,x)$ and outputs a real number representing the probability of occupancy.
+        \item We randomly sample points in the 3D bounding volume of the object. The loss function for a batch $\mathcal B$ is:
+        $$\mathcal L_{\mathcal B}^{\text{gen}} = \frac{1}{|\mathcal B|} \sum_{i=1}^{|\mathcal B|} \lr[] {\sum_{j=1}^{K} \mathcal L (f_\theta(p_{ij}, z_i), o_{ij}) + \text{KL}(q_\psi(z | (p_{ij}, o_{ij})_{j=1:K}) \lVert\, p_0(z))}$$
+        \item Creates a mesh using coarse voxel grid and evaluating occupancy network for cells in this grid.
+For boundary voxels, we further subdivide them and repeat, until we reach the desired resolution, before applying Marching Cubes.
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{DRWR: A Differentiable Renderer without Rendering for Unsupervised 3D Strucutre Learning }
+    \textbf{Citation:} Han \textit{et al.}
+
+    \textbf{Summary:} A unsupervised differentiable renderer using losses in a clever way as opposed to pixel-wise losses.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item This is unsupervised, so only requires images for training.
+        \item Existing pixel-wise loss functions render images from 3D structures 
+    \end{itemize}
 \end{ArticleSummary}

### 2024-08-02 docs/logbook.tex
Implemented PCA for initialisation

\documentclass{project-logbook}

\usepackage{csquotes}
\usepackage{lastpage}
% You can add the maintainer and the contributors in the preamble of the document

% This is how you add the maintainer (there is only one) the maintainer is automatically a contributor, the format is
% name, id (to appear in short messages), the affiliation.
\SetMaintainer{Jeffrey Liang}{u7013004}{ANU School of Computing}

% This is how you add a contributor, you can have as many contributors as you like (althought
% there is a limited space on the front page, at least 25 contributors fit in the front page, which should
% be more than enought).
% The format is similar to Maintainer: key, name, id, affiliation. The key field is the difference. It is used
% to internally identify the contributor, such that one can access the data of each contributor. This is discussed
% below.

% The same goes for the titles of the project.
% Three levels can be used. The project title sets the main title of the project, which appears in the middle of the front page.
% The project subtitle appears below the project title in the front page and serves as additional information to the title.
%  Additionally, a shorter title must be set to appear on the header of the document, the header name.
\SetProjectTitle{Honours Logbook}
\SetProjectSubtitle{The subtitle of the project}
\SetProjectHeaderName{The very fancy project}

% Below the project there is space to add a brief summary of the project for easy understanding when looking at the front page.
\SetProjectSummary{Here you can add a brief summary of your project such that someone can quickly see what it is about, do not spend more than two lines on it.}

% The institution logo (typically the one of the maintainer, but a logo file combining different logos can be used, but needs to
% be in one single pdf or png file) can be set using the following macro
\SetInstitutionLogo{figures/mcod_logo.png}

% Set the bibliography file
\addbibresource{mybib.bib}

% Start the document
\begin{document}

%----------------------------------------------------------------------------------------
%  Front page
%----------------------------------------------------------------------------------------

% The front page is constructed like this, it simply adds the logo, title, subtitle, summary, the maintainer's information, along
% with the list of all contributors.
\MakeFrontPage

% Standard LaTeX can be used throughout the document, but custom functionality has been added for simplifying some recurrent tasks and
% ensuring a consistent format.

%----------------------------------------------------------------------------------------
%  Body of the document
%----------------------------------------------------------------------------------------
\newpage

% The document is separated into sections, which in turn can be subdivided into subsections
\section{Overview} \label{sec:overview}
	% Sections start in a new page, for this reason sections contain clearly disjoint parts of the document
	This document shows an example of how to use the \texttt{project-logbook} \LaTeX~class. This class is based upon the \texttt{article} class and may be used together with other packages for additional functionality. This class implements functionality for simplifying writing project/research logbooks and unifying its layout. We will  go over the different functionalities to display how to use them.

	\subsection{Preamble}
		In the preamble of the document you may set several parameters of this document.

		\subsubsection{Maintainer and contributors}
			Everyone who contributes to this document, should be added as a contributor. In this way it is know who contributed and we can also use this to identify who added some elements. Adding a contributor is done with \texttt{\textbackslash CreateContributor\string{<key>\string}\string{<name>\string}\string{<id>\string}\string{<affiliation>\string}}, where
			\begin{description}
				\item[\texttt{<key>}] is the shorthand used to internally identify a contributor and to retrieve its data, this should be a short word without special characters.
				\item[\texttt{<name>}] is the full name of the contributor to display on the front page (but may also be used in other situations if desired).
				\item[\texttt{<id>}] is the abbreviation of the contributor name, to display when associating the contributor to a note (see below) or a meeting minutes box (see below).
				\item[\texttt{<affiliation>}] is the affiliation of the contributor, mainly to display on the front page.
    			\end{description}

			This document is maintained by a unique maintainer, which is automatically defined as a contributor with key \texttt{maintainer}. The maintainer can be set in the preamble with  \texttt{\textbackslash SetMaintainer\string{<name>\string}\string{<id>\string}\string{<affiliation>\string}}. The meaning of the different input parameters is identical to what was seen above for the contributor.

			The maintainer name, id, and affiliation can be accessed via the commands \texttt{\textbackslash MaintainerName}, \texttt{\textbackslash Maintainerid} and \texttt{\textbackslash MaintainerInsitution}, which for this case gives: \MaintainerName~(name), \Maintainerid~(id), and \MaintainerInstitution~(affiliation).

			In a similar way, a contributor can be accessed via the commands \texttt{\textbackslash ContributorName\string{<key>\string}}, \texttt{\textbackslash Contributorid\string{<key>\string}}, and \texttt{\textbackslash ContributorInstitution\string{<key>\string}}, which for this case gives (for key \texttt{chalmers}):  \ContributorName{chalmers}~(name), \Contributorid{chalmers}~(id), and \ContributorInstitution{chalmers}~(affiliation).

		\subsubsection{Project title, subtitle, and header name}
			The project title (which appears in the center of the front page) can be set in the preamble of the document with \texttt{\textbackslash SetProjectTitle\string{<title>\string}}, where \texttt{<title>} is the title we wish to set.

			Below the project title we can have a project subtitle, which can also be set in the preamble of the document with \texttt{\textbackslash SetProjectSubtitle\string{<subtitle>\string}}, where \texttt{<subtitle>} is the subtitle we wish to set.

		\subsubsection{Maintainer institution logo}
			On the top right corner of the front page you can add the logo of the institution of the maintainer. This can be done in the preamble with the command \texttt{\textbackslash SetInstitutionLogo\string{<path\_to\_file>\string}}, where \texttt{<path\_to\_file>} is the path to a .pdf or .png file. If you do not want to show a logo, use a white figure.

	\subsection{Body}
		The front page is generated with the command \texttt{\textbackslash MakeFrontPage}. This automatically adds the table of contents on the second page of the document, with direct links to the sections, for quick access.

		One important section is the Meetings section, which is subdivided into External and Internal. The External subsection contains the minutes of the meetings with external partners and the Internal sections contains the internal meetings associated to the institution of the Maintainer. Any contributor (including the maintainer) can be an author of Meeting minutes.

		Following the Meetings section, any number of sections can be added focussing on specific topics. The goal of these is to discuss the current state of the development of a particular part of the project. More discussed below.

		Following these sections, there is an Appendix, containing the list of references (papers to use as reference later when writing a paper), a list of resources such as a link to a video presentation, a tutorial, or something else that is less formal than a reference. In the end, you have a list of TODOs that have been defined in the document. They appear in the order in which they appeared in the document, not by writing date.

\section{Progress Logbook}
	\subsection{Progress}
	\begin{MeetingMinutes}{Sun 12/05/2024}{u7013004}
		Reading MeshSDF paper.
	\end{MeetingMinutes}

	\begin{MeetingMinutes}{Sat 18/05/2024}{u7013004}
		Reading MeshSDF, the Learning Implicit Fields papers, and Structure from Motion. Need to look up what the following mean:
		\begin{itemize}
			\item differentiable rendering, surface reconstruction
		\end{itemize}
	\end{MeetingMinutes}

	\begin{MeetingMinutes}{Sun 19/05/2024}{u7013004}
		Reading more from Structure-from-Motion Revisited. Didn't really understand much of SfM last night.
	\end{MeetingMinutes}

	\begin{MeetingMinutes}{Thu 27/06/2024}{u7013004}
		Attempting 3D version of the ellipse problem. \\
		Question: Why do we invert the $u_i$s? Answer: you bring the 1 over and multiply by some expression on both sides.
	\end{MeetingMinutes}

	\begin{MeetingMinutes}{Fri 5/07/2024}{u7013004}
		Having been making headway on the non-axis-aligned ellipsoid.
		I've been able to fit an ellipse to one but I haven't been able to do the bi-level optimisation.
		I'm going to try a couple approaches:
		\begin{itemize}
			\item reparametrise everything to radians and $a,b,c$ rather than squared inverses of semi-axes 
			\item try original objective function rather than focusing on optimising rotation 
			\item last resort is to figure out what the SqrtBackward error is
		\end{itemize}
	\end{MeetingMinutes}

	\begin{MeetingMinutes}{Sun 7/07/2024}{u7013004}
		Was able to switch implementation to radians and $a,b,c$ rather than squared inverses $1/a^2,...,1/c^2$. 
		It's still temperamental though and will hopefully be finetuned when meeting with Chamin who hopefully has a few tricks.
		Going to read some papers today about DRWR and the DDN and take notes to try and fully understand derivations etc.
	\end{MeetingMinutes}

	\begin{MeetingMinutes}{Sat 20/07/2024}{u7013004}
		Reading additional papers and waiting for Steve's meeting. Reading "Small Steps and Level Sets" by Chamin and will document.
	\end{MeetingMinutes}

	\begin{MeetingMinutes}{Mon 29/07/2024}{u7013004}
		Trying to do experiments to ensure I know the cause of my errors and that it's not some implementation error.
		Will do:
		\begin{itemize}
			\item small volume initialisations
			\item close to ground truth initialisations
			\item tightening the constraint slowly
			\item PCA?
		\end{itemize} 
		Ideal semi-axes lengths is: 0.28209
		Trials:
		\begin{enumerate}
			\item initialisation: 0.2, 0.4, 0.3, 30°, 25°, -120°. final: 0.271, 0.296,0.28, 29.98°, 24.99°, 60.0°
			\item initialisation: 0.1, 0.2, 0.3, 30°, 25°, -120°. final: 0.479, 0.29,0.0913, 83.83°, 128.2°, 47.02° (no improvement).
				 \\ the issue here is that the fitted ellipse is too far away from a sphere for it to approach a sphere?
			\item initialisation: close to ideal. result: optimised fine and as expected.
		\end{enumerate}
		Perhaps farthest point sampling would prevent the ellipse from over concentrating at the poles.

		Issue! For Trial 2, a close reinitialisation of 0.1, 0.3, 0.5 and angles 30°, 25°, -120° worked well. In fact, an exact reinitialisation worked?! But then when reverting back so that the fitted ellipse had those params, the gradients calculated were zero.

		Steps to reproduce error:
		\begin{enumerate}
			\item Initialise with 0.1, 0.2, 0.3 and 30°, 25°, -120°. Run the first cell and then second cell, and we should see no deformation of the fitted ellipsoid to a sphere - the parameters remain as 0.479, 0.29,0.0913, 83.83°, 128.2°, 47.02°. The gradients are very close to 0.
			\item Initialise with the parameters 0.479, 0.29,0.0913 and 83.83°, 128.2°, 47.02°. We see that the gradients are non-zero and we optimise to a sphere.
		\end{enumerate}
		a, b, c = 0.4786546028, 0.2901328092,0.09132074619 \\ 
		yaw, pitch, roll = 83.82856741, 128.3492915, 47.0228292

	\end{MeetingMinutes}

\section{Topic A to focus on} \label{sec:topic_A}
	We can use sections organize the text into different parts  such that specific topics are addressed in their own section.

	\begin{HighlightedNote}{}
		Sections always start at the top of a new page in order to generate clearly separated parts of the document. In this first section (Overview) one should introduce what this project is about, in a high level so you can do a simple explanation of the topic and the research challenges, who is involved, etc.

		Since this information is important, we have added it within a \texttt{HighlightedNote} environment. This is the version without a title.
	\end{HighlightedNote}

	\begin{HighlightedNote}{You can add multiple things inside a \texttt{HighlightedNote}}
		And this is the version with a title. You can use the title for example if you want to clearly state what is is, or if you want to make a larger note and give an idea of what it is about.

		You can add equations to clarify your ideas
		\begin{equation}
			e^{i\pi} + 1 = 0
		\end{equation}

		Some tables

		\begin{center}
			\begin{tabular}{ccccc}
            			{}  & \multicolumn{1}{c}{$A$} & \multicolumn{1}{c}{$N$} & \multicolumn{1}{c}{$S$} & \multicolumn{1}{c}{$A/S$} \\
            			\midrule
            			a & 292 & 117   & 409   & 0.714 \\
            			b & 104 & 386   & 490   & 0.212 \\
            			c & 24  & 7     & 31    & 0.774 \\
            			d & 28  & 27    & 55    & 0.509 \\
            			e & 183 & 1958  & 2141  & 0.085 \\
            			\midrule
            			& 631 & 2495 & 3126 & 0.202 \\
        			\end{tabular}
    		\end{center}
		and other elements.

	\end{HighlightedNote}

	\begin{HighlightedNote}{With a fancy title}
		The \texttt{HighlightedNotes} environment that generates these fancy boxes, can be used with a title \texttt{\textbackslash begin\string{HighlightedNotes\string}\string{the\_title\_text\string}}, as in this case, or without a title \texttt{\textbackslash begin\string{HighlightedNotes\string}\string{\string}}, as in the previous case.
	\end{HighlightedNote}

	\subsection{A subsection}
		You can subdivide your section in a finer level using \texttt{\textbackslash subsection}.

		\subsubsection{A subsubsection}
			The subsections can be subdivided with \texttt{\textbackslash subsubsection}.

			\subsubsubsection{A subsubsubsection}
				And an additional level was added to subdivide subsubsections named \texttt{\textbackslash subsubsubsection}. This should give more than enough subdivision levels to your text for fine grained organisation.

				\begin{HighlightedNote}{}
					In most Machine Learning research projects you will be using some kind of samples from a Dataset to learn a model. Thus it is extremely important that you carefully describe the dataset and why you believe is a good dataset for the project and what type of preprocessing are you going to apply.
				\end{HighlightedNote}

				As usual, you can use your references to cite relevant articles as \cite{einstein} and \cite{knuth-fa}.

				You can use tables and, of course, make references to them, like the amazing \tabref{tab:populations}.

				\begin{table}[ht]
					\centering
						\begin{subtable}{0.48\textwidth}
							\centering
        								\begin{tabular}{*{5}{r}}
									{}  & \multicolumn{1}{c}{$A$} & \multicolumn{1}{c}{$N$} & \multicolumn{1}{c}{$S$} & \multicolumn{1}{c}{$A/S$} \\
									\midrule
									a & 292 & 117   & 409   & 0.714 \\
									b & 104 & 386   & 490   & 0.212 \\
									c & 24  & 7     & 31    & 0.774 \\
									d & 28  & 27    & 55    & 0.509 \\
									e & 183 & 1958  & 2141  & 0.085 \\
									\midrule
									& 631 & 2495 & 3126 & 0.202 \\
								\end{tabular}
							\caption{Training Set}
						\end{subtable}
    						%
    						\begin{subtable}{0.48\textwidth}
        							\centering
        								\begin{tabular}{*{6}{r}}
									{}  & \multicolumn{1}{c}{$A'$} & \multicolumn{1}{c}{$N'$} & \multicolumn{1}{c}{$S'$} & \multicolumn{1}{c}{$A'/S'$} & $S'/S$ \\
									\midrule
									a & 40  & 40    & 80    & 0.5   & 0.20 \\
									b & 49  & 49    & 98    & 0.5   & 0.20 \\
									c & 4   & 3     & 7     & 0.57  & 0.23 \\
									d & 5   & 5     & 10    & 0.5   & 0.18 \\
									e & 53  & 53    & 106   & 0.5   & 0.05 \\
									\midrule
									& 151 & 150 & 301 & 0.50 & 0.10 \\
								\end{tabular}
								\caption{Validation Set}
						\end{subtable}

    						\caption{Population properties $A \equiv$ Abnormal, $N\equiv$ Normal, $S\equiv A+N$}
    						\label{tab:populations}
				\end{table}

	\subsection{Another subsection} \label{sub:another_subsection}
		Relevant equations can be added

		\begin{equation}
    			x_i' \gets \frac{x_i-\bar{x}}{\upsigma}
		\end{equation}
		\begin{equation}
    			\bar{x} = \frac{1}{N}\sum_{i=1}^{N} x_i \qquad \upsigma^2 = {\frac{1}{N}\sum_{i=1}^{N} (x_i-\bar{x})}
		\end{equation}

		\begin{HighlightedNote}{Strike-through}
 			Often you will be wrong on your assumptions, but do not throw them away completely, just cross them out in case you need them later using the \texttt{\textbackslash sout} command and it will look like \sout{this} or \texttt{\textbackslash soutthick} command and it will look like \soutthick{this}
		\end{HighlightedNote}

		\sout{Medoid computation is perfomed at $f_s'' = f_s'/5 = \SI{200}{\kilo\hertz}$ to speed computation. Simple analysis was performed to check that the features extracted from these 200Hz-medoids were approximately the same as the ones extracted from the  1khz-medoids.}

		\soutthick{Downsampling can be seen as a problem of information loss in the frequency spectrum. If the frequency content $f > f_s/(2N)$ is mostly empty for $N$ when we downsample by said $N$ we will only be losing information in that range.}

		\begin{HighlightedNote}{TODOs}
			You will often have pending tasks that you need to track. This project logbook allows you to include both high and low priority todos that will be summarised in a list at the end of the file. Use the commands \texttt{\textbackslash hightodo\string{<date\_added>\string}\string{<author\_key>\string}} and \texttt{\textbackslash lowtodo}, including a date is recommended for tracking purposes.

			Note: define your \texttt{userId} in the preamble
		\end{HighlightedNote}

		\lowtodo{2016-05-23}{chalmers}{Try the code in the newest version of numpy compiled with optimized BLAS.}

		\hightodo{2016-05-27}{maintainer}{Perform convergence analysis with the latest version of the code.}

\section{Topic B to focus on} \label{sec:topic_B}

	\subsection{Algorithms} \label{sub:algorithms}

		\begin{HighlightedNote}{}
			Sometimes the most straightforward way to explain a procedure is just to give it in a algorithmic format, it takes a little time but it will force you to go through the steps and you will most likely be able to reuse it on you paper. For the full documentation see here \url{https://texdoc.org/serve/algorithmicx/0}.
		\end{HighlightedNote}

		\begin{algorithm}
			\caption{Euclid’s algorithm}\label{alg:euclid}
			\begin{algorithmic}[1]
				\Procedure{Euclid}{$a,b$}\Comment{The g.c.d. of a and b}
					\State $r\gets a\bmod b$
					\While{$r\not=0$}\Comment{We have the answer if r is 0}
						\State $a\gets b$
						\State $b\gets r$
						\State $r\gets a\bmod b$
					\EndWhile\label{euclidendwhile}
					\State \textbf{return} $b$\Comment{The gcd is b}
				\EndProcedure
			\end{algorithmic}
		\end{algorithm}

	\subsection{Code} \label{sub:code}
		\begin{HighlightedNote}{}
			If the algorithm is too vague and you feel like you need the source code you can also insert it. You can put LaTeX code inside by using \texttt{<@ @>} delimiters, highlight small pieces of the code with \texttt{<| |>} delimiters, and highlight full lines (see the source code). For now only Python colored syntax highlighting is available, but most languages are supported and the color scheme can be added. More information available here \url{https://www.overleaf.com/learn/latex/Code_listing}.

			Note that these pieces of code in the \LaTeX~document must start with no indent.
		\end{HighlightedNote}

\begin{lstlisting}[style=Python, linebackgroundcolor={
	\ifnum \value{lstnumber}=8 \color{orange!30} \fi
	\ifnum \value{lstnumber}=10 \color{orange!30} \fi
	\ifnum \value{lstnumber}=17 \color{orange!30}\fi
	\ifnum \value{lstnumber}=18 \color{orange!30}\fi}]
def DTW_distance(s1, s2):
	"""
	Function to compute the Dynamic Time Warping in Python between two signals
	"""
	DTW={}

	for i in range(len(s1)):
		DTW[(i, -1)] = float('inf') # By default <@\color{pgreen} $\infty$@>
	for i in range(len(s2)):
		DTW[(-1, i)] = float('inf') # By default <@\color{pgreen} $\infty$@>
	DTW[(-1, -1)] = 0

	for i in range(len(s1)):
		for j in range(len(s2)):
			dist= <|(s1[i]-s2[j])**2|>
			DTW[(i,j)] = dist + min(DTW[(i-1, j)],DTW[(i, j-1)], \
				DTW[(i-1, j-1)])

	return sqrt(DTW[len(s1)-1, len(s2)-1])
\end{lstlisting}

	\subsection{Diagrams} \label{sub:diagrams}
		\begin{HighlightedNote}{}
			For simple diagrams I highly recommend learning TiKZ, you will be drawing the diagrams in pure \LaTeX which has a steep learning curve but once you get used to it, it can be quite easy to display and do \texttt{for} loops to draw multiples line at once.
		\end{HighlightedNote}

		\begin{figure}[!h]
			\centering
				\tikzset{%
  					every neuron/.style={
    						circle,
    						draw,
    						minimum size=1cm
  					},
  					neuron missing/.style={
    						draw=none,
    						scale=4,
    						text height=0.333cm,
    						execute at begin node=\color{black}$\vdots$
  					},
				}

				\begin{tikzpicture}[x=1.5cm, y=1.5cm, >=stealth]
					\foreach \m/\l [count=\y] in {1,2,3,4,missing,5}
  						\node [every neuron/.try, neuron \m/.try] (input-\m) at (0,2.5-\y) {};

					\foreach \m [count=\y] in {1,2,missing,3}
  						\node [every neuron/.try, neuron \m/.try ] (hidden-\m) at (2,2-\y*1.25) {};

					\foreach \m [count=\y] in {1,2,missing,3}
  						\node [every neuron/.try, neuron \m/.try ] (output-\m) at (4,1.5-\y) {};

					\foreach \l [count=\i] in {1,2,3,4,n}
  						\draw [<-] (input-\i) -- ++(-1,0)
    							node [above, midway] {$I_\l$};

					\foreach \l [count=\i] in {1,2,n}
  						\node [above] at (hidden-\i.north) {$H_\l$};

					\foreach \l [count=\i] in {1,2,n}
  						\draw [->] (output-\i) -- ++(1,0)
    							node [above, midway] {$O_\l$};

					\foreach \i in {1,...,5}
  						\foreach \j in {1,...,3}
    							\draw [->] (input-\i) -- (hidden-\j);

					\foreach \i in {1,...,3}
  						\foreach \j in {1,...,3}
    							\draw [->] (hidden-\i) -- (output-\j);

					\foreach \l [count=\x from 0] in {Input, Hidden, Ouput}
  						\node [align=center, above] at (\x*2,2) {\l \\ layer};

				\end{tikzpicture}
		\end{figure}

		\begin{HighlightedNote}{}
			However, sometimes you will need more complicated diagrams (or maybe you do not like TiKZ, in that case I recommend a vector drawing tool such as Inkscape which allows \LaTeX \,embedding)
		\end{HighlightedNote}

		% \begin{figure}[h]
    	% 		\includegraphics[width=0.9\textwidth]{figures/LSTM}
    	% 		\centering
    	% 		\caption{Gate Recurrent Unit in a Long Short Term Memory Neural Netwok (GRU-LSTM). \\ \emph{Credit to \url{https://colah.github.io/posts/2015-08-Understanding-LSTMs/}}}
    	% 		\label{fig:LSTM}
		% \end{figure}

\section{Topic C to focus on} \label{sec:topic_c}
	\subsection{Figures} \label{sub:figures}

		\begin{HighlightedNote}{}
			In general the best way to visualize your results will be some figures, I recommend Python's matplotlib for generating them or any other tool you are familiar with.
		\end{HighlightedNote}

		% \begin{figure}[htp]
    	% 		\centering
		% 		\begin{subfigure}{0.48\textwidth}
        % 					\centering
        % 						\includegraphics[width=\textwidth]{figures/sfem_1D_p3.pdf}
        % 						\caption{One-dimensional spectral element basis functions of polynomial degree $p = 3$ for subspace $G_{3} \subset H^{1}(\Omega)$, $\Omega\in\mathbb{R}$.}
    	% 			\end{subfigure}
    	% 			\hspace{.35cm}
    	% 			\begin{subfigure}{0.48\textwidth}
        % 					\centering
        % 						\includegraphics[width=\textwidth]{figures/sfem_1D_p5.pdf}
        % 						\caption{One-dimensional spectral element basis functions of polynomial degree $p = 5$ for the subspace $G_{5} \subset H^{1}(\Omega)$, $\Omega\in\mathbb{R}$.}
    	% 			\end{subfigure}

		% 		\vspace{1cm}

		% 		\begin{subfigure}{0.48\textwidth}
        % 					\centering
        % 						\includegraphics[width=\textwidth]{figures/sfem_p3.png}
        % 						\caption{Two-dimensional spectral element basis functions of polynomial degree $p = 3$ for subspace $G_{3} \subset H^{1}(\Omega)$, $\Omega\in\mathbb{R}^{2}$.}
    	% 			\end{subfigure}
    	% 			\hspace{.35cm}
    	% 			\begin{subfigure}{0.48\textwidth}
        % 					\centering
        % 						\includegraphics[width=\textwidth]{figures/sfem_p5.png}
        % 						\caption{Two-dimensional spectral element basis functions of polynomial degree $p = 5$ for the subspace $G_{5} \subset H^{1}(\Omega)$, $\Omega\in\mathbb{R}^{2}$.}
    	% 			\end{subfigure}

   		% 		\vspace{.25cm}
    	% 			\caption{Example basis functions for the spectral finite element method in 1D and 2D.}
    	% 			\label{fig:plots}
		% \end{figure}

% subsection figures (end)

\subsection{Tables} % (fold)
\label{sub:tables}

\begin{HighlightedNote}{}
\LaTeX \; booktab environments are really good to showcase and track your results, however they can get fairly messy. My suggestion is to generate them via Python automatically and store the results in either a plain text file or a spreadsheet (there are packages to read spreadsheets with Python)
\end{HighlightedNote}

\begin{table}[ht]
     \centering
     \begin{tabular}{rrrrrrrrrr}
         $\upsigma \setminus \tau$ & \multicolumn{1}{c}{0} &\multicolumn{1}{c}{1} &\multicolumn{1}{c}{2} &\multicolumn{1}{c}{3} &\multicolumn{1}{c}{4} &\multicolumn{1}{c}{5} &\multicolumn{1}{c}{6} &\multicolumn{1}{c}{7} &\multicolumn{1}{c}{8}\\
          \midrule
         0.0 & 100.0  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0 \\
         0.2 & 100.0  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0 \\
         0.4 & 100.0  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0 \\
         0.6 & 98.6  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0 \\
         0.8 & 84.7  &99.5  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0 \\
         1.0 & 28.1  &98.3  &99.9  &100.0  &100.0  &100.0  &100.0  &100.0  &100.0 \\
         1.2 & 1.3  &88.7  &99.4  &99.9  &99.8  &99.9  &100.0  &99.9  &100.0 \\
         1.4 & 0.0  &57.1  &96.2  &99.3  &99.0  &99.3  &99.4  &99.8  &99.7 \\
         1.6 & 0.0  &18.6  &81.2  &93.0  &93.7  &94.8  &95.6  &92.3  &93.3 \\
         1.8 & 0.0  &2.4  &42.8  &67.0  &70.1  &72.1  &69.0  &69.1  &68.6 \\
         2.0 & 0.0  &0.1  &9.0  &23.1  &24.5  &26.9  &28.2  &27.3  &27.3 \\
         \midrule
          $t$(ms) &27.92 &40.23 &77.30 &157.27 &252.05 &342.18 &381.46 &399.85 &413.72\\
     \end{tabular}
     \caption{Performance of the algorithm for 128-bit key and with multiple readings per key}
     \label{tab:error_128_avg}
 \end{table}

% subsection tables (end)

% section results (end)

%----------------------------------------------------------------------------------------
%   REFERENCE LIST
%----------------------------------------------------------------------------------------
\appendix

%% References (bibliography)
\section{References}
	\begin{HighlightedNote}{}
		Do not forget to cite the papers that you are using in your research, this way your work will be infinitely easier to write down and to review when the time comes.
	\end{HighlightedNote}

	% Add the bibliography without the header. You can remove the heading=none and \section{References} to have it in a different way.
	% Here we keep the references as a section from the Appendix, as an example
	\printbibliography[heading=none]

\section{Resources}
	\begin{HighlightedNote}{}
		It is a good idea to record sources that explain concepts or provide tools so the research is both better documented and if someone has to continue it there is enough supporting documentation.
	\end{HighlightedNote}
	\begin{itemize}
    		\item Quick read in DTW and Keogh Lower Bounding. \newline
			\url{http://alexminnaar.com/time-series-classification-and-clustering-with-python.html} \newline
			\url{http://nbviewer.jupyter.org/github/alexminnaar/time-series-classification-and-clustering/blob/master/Time%20Series%20Classification%20and%20Clustering.ipynb}

		\item Parallelizing DTW -- Good article on making a parallel version of DTW. Uses Keogh lower bound not as a linear approximation but as a pruning device. \newline
			\url{https://www.andrew.cmu.edu/user/mmohta/15418Project/finalreport.html}

    		\item Deep Learning
    			\begin{itemize}
        				\item Intro to LSTM \newline
					\url{https://colah.github.io/posts/2015-08-Understanding-LSTMs}
        				\item Intro to CNN \newline
					 \url{https://colah.github.io/posts/2014-07-Conv-Nets-Modular/}
				 \item Why are LSTMs are so useful, impressive result in character pattern and syntax learning \newline
					\url{https://karpathy.github.io/2015/05/21/rnn-effectiveness/}
    			\end{itemize}
	\end{itemize}

%% Add the list of todos in the end
\section{TODOS}
	\begin{HighlightedNote}{}
		Here you will have all your TODOs grouped with anchor links to the parts of the document where they are. Really handy if you do not know where to continue with your project.
	\end{HighlightedNote}

	\AddListOfTodos

\end{document}

### 2024-08-02 docs/papers.tex
Implemented PCA for initialisation

@@
 \usepackage[utf8]{inputenc}
-\usepackage{amsmath,amsfonts,amssymb, bm}
+\usepackage{amsmath,amsfonts,amssymb, bm, mathtools}
 \usepackage{graphicx}
@@
 
-    \textbf{Summary:} A unsupervised differentiable renderer using losses in a clever way as opposed to pixel-wise losses.
+    \textbf{Summary:} An unsupervised differentiable renderer uses losses to evaluate the alignment of projections of reconstructed 3D point clouds with ground truth object silhouettes, instead of pixel-wise losses.
+Idea is to pull points to the surface of the silhouette and then push them far away from each other.
 
@@
         \item This is unsupervised, so only requires images for training.
-        \item Existing pixel-wise loss functions render images from 3D structures 
+        \item Existing pixel-wise loss functions render images from 3D structures but require rasterisation or pixel-wise interpolation commonly.
+        \item DRWR: doesn't require visibility handling, shading, or pixel-wise interpolation.
+        \item Loss function: smooth unary loss pulls projection of each 3D point into the foreground (image we're interested in) and binary/pairwise loss pushes each pair of projections lying inside the foreground far away (structure-aware since only considers repulsion once both points appear in foreground).
+This ensures the entire foreground is covered and prevents clumping.
     \end{itemize}
 \end{ArticleSummary}
+
+\begin{ArticleSummary}{Deep Declarative Networks: A New Hope}
+    \textbf{Citation:} Gould, Hartley, and Campbell, 2020
+
+    \textbf{Summary:} Networks are defined in terms of desired behaviour rather than an explicit forward function, with the desired behaviour implicitly defined as a solution to an optimisation problem.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item We only need the optimal solution to an optimisation problem and twice-differentiable objective and constraint functions to perform backpropagation - the method used to find the solution can be non-differentiable such as RANSAC.
+        \item Identities and solutions use the Schur complement and block matrix inverses (on Wikipedia).
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{Small Steps and Level Sets: Fitting Neural Surface Models with Point Guidance}
+    \textbf{Citation:} Koneputugodage \textit{et al.}, 2024
+
+    \textbf{Summary:} Spherical initialisations of point clouds for neural SDFs often get trapped in local minima due to the biased initialisation.
+This approach uses intermediary "guiding points" during optimisation to allow the point cloud to fit complex shapes.
+Uses percentage of self-intersecting outward normals (SION\%)  to provide a metric of the diffence between surface geometries of the current and target shapes.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item The idea behind SION\% is that if there are large concavities in the target shape, the surface normals will self-intersect and few will make it out to the current surface which doesn't fit inside the cavity - however, if it did "fit" inside, then the surface normals would not self-intersect.
+        \item Performance decreases on IoU and Chamfer distance metrics as the shapes get more complex (increasing SION\%), though at much slower rate than other SotA.
+        \item In each iteration, the guiding points are determined by taking the guiding points from the previous iteration by moving it towards the closest target point.
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{DeepSDF: Learning Continuous Signed Distance Functions
+    for Shape Representation}
+    \textbf{Citation:} Park \textit{et al.}, 2019
+
+    \textbf{Summary:} Revolutionary paper about using neural networks to implicitly model the signed distance function of a 3D watertight shape.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Trained using set of pairs $X$ composed of 3D point samples and the SDF value, $X \coloneq \lrbrace{(\bm x, s) \mid SDF(\bm x) = s}$, such that the trained network is a good approximator of the SDF, that is, $f_\theta(\bm x) \approx SDF(\bm x)$.
+Loss is an $L_1$ loss function which uses clamping.
+        \item Auto-decoder which has the latent code as input and directly optimises that, involves taking the maximum-a-priori estimate of the loss function + a regularising term over all points over all shapes
+        \item To make the network flexible and adaptable to many shapes, a latent code describing the shape is inputted with the 3D point into the network, rather than a single point input with the network containing information about the shape.
+        \item The SDF value is the signed distance from the shape's surface, with negative values indicating the point is inside the shape, zero values meaning the point lies on the surface, and positive values meaning the point is outside the shape.
+The magnitude of the value is the distance to the surface.
+    \end{itemize}
+\end{ArticleSummary}
+
 \end{document}

### 2024-08-15 docs/logbook.tex
Cleaned up comments

@@
 		yaw, pitch, roll = 83.82856741, 128.3492915, 47.0228292
+	\end{MeetingMinutes}
 
+	\begin{MeetingMinutes}{Thu 15/08/2024}{u7013004}
+		Steve told me that I needed to update the DDN package and it worked!\\
+		I spent the past week a bit confused on how to do projection but Steve gave me some clear instructions. I've switched the volume and surface area constraints, but the optimisation is being worse.
 	\end{MeetingMinutes}

### 2024-09-04 docs/logbook.tex
Updates to logbook and projection

@@
 		I spent the past week a bit confused on how to do projection but Steve gave me some clear instructions. I've switched the volume and surface area constraints, but the optimisation is being worse.
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Sun 18/08/2024}{u7013004}
+		Going to work on optimising the volume to be a certain target now and read a paper. Then, I'll write some things in the template. Efficiency!
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Wed 4/09/2024}{u7013004}
+		For the past few days, I've been trying to mathematically formulate the projection.
+Just tested it in code today and it seems that I am minimising the area of the cross-section with the $x$-$y$ plane instead.
+So, just need to take Schur complement instead of projection with matrix P.
 	\end{MeetingMinutes}

### 2024-09-04 docs/projection.tex
Updates to logbook and projection

\documentclass{article}
\input{commands}
\usepackage{amsmath, amssymb}

\newcommand{\x}{\mathbf{x}}
\title{Mathematical Formulation of Projection Problem}
\author{Jeffrey Liang}
\setlength{\parskip}{\baselineskip}%
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}

\begin{document}
\maketitle

\section{Goal}
To project a set of points $\mathcal X$ onto the nearest ellipsoid with volume 1, such that the projection of this ellipsoid in two different views produces ellipses which minimise the error between their areas and two target areas.

\section{Parametrising Ellipsoids}

An ellipsoid $\mathcal E$ can be parametrised by a positive definite, symmetric matrix $A \in \R^{3\times 3}$ as follows:
\begin{equation}
    \mathcal E = \lrbrace{\x \in \R^3 | \x^\top A \x = 1, A \succ 0}.
\end{equation}
By performing an eigen-decomposition on $A$, we have that $A$ can be represented as $\mathbf A= \mathbf Q^\top \mathbf  \Lambda \mathbf  Q$, where $\mathbf Q\in SO(3)$ is an orthonormal rotation matrix and $\mathbf \Lambda$ is a diagonal matrix with eigenvalues that are the inverse square of the semi-axes lengths of $\mathcal E$:
\begin{align}
    A &= \mathbf Q^\top \mathbf \Lambda \mathbf Q \\
        & = \mathbf Q^\top \begin{bmatrix}
            \lambda_1 & & \\
            & \lambda_2 & \\
            & & \lambda_3
        \end{bmatrix}\mathbf Q \\
        &= \mathbf Q^\top \begin{bmatrix}
            1/a^2 & & \\
            & 1/b^2 & \\
            & & 1/c^2
        \end{bmatrix}\mathbf Q,
\end{align}
where $a, b,$ and $c$ are the semi-axes lengths of the ellipsoid $\mathcal E$. As the rotation matrix can be parametrised by three Euler angles -- roll $\psi$, pitch $\theta$, and yaw $\phi$ -- in addition to the three semi-axes, an ellipsoid in 3D can thus be parametrised by six degrees of freedom which we can represent as a vector $\mathbf u \in \R^6$.

\section{Inner-Level Problem}
Let $\mathcal{X} = \lrbrace{\mathbf x_i}$ be a set of sampled points $\mathbf x_i \in \R^3$ from a noisy ellipsoid. Moreover, let $Q: \R^3 \to SO(3)$ map three angles (roll, pitch, yaw) to the orthonormal 3D rotation matrix $\mathbf Q$ defined above and $\Lambda: \R^3 \to \R^{3\times 3}$ map three lengths to the diagonal matrix $\mathbf \Lambda$ described above.

To project these points to the nearest ellipsoid with a volume of 1 unit cubed, we can minimise the least squares error subject to the volume constraint.
\begin{equation}
    \begin{array}{ll}
        \underset{\text {over } \mathbf u\in \R^6}{\text{minimize}} & \sum_{i=1}^{m} \left({\mathbf x_i}^\top Q(\mathbf u_{4:6}) ^\top \Lambda(\mathbf u_{1:3}) Q(\mathbf u_{4:6}) \mathbf x_i - 1\right)^2 \\[0.8em]
        \text{subject to} & \frac{4}{3} \pi \mathbf u_1 \mathbf u_2 \mathbf u_3 = 1
      \end{array}          
\end{equation}
Or, in matrix form, letting $$\mathbf X = \begin{bmatrix}
    | & \cdots & | \\
    \mathbf x_1 & \cdots & \mathbf x_m \\
    | & \cdots & |
  \end{bmatrix},$$
  we have

\begin{equation}
    \begin{array}{ll}
        \underset{\text {over } \mathbf u\in \R^6}{\text{minimize}} & \lr\Vert\Vert {\text{diag}(\mathbf X^\top Q(\mathbf u_{4:6}) ^\top \Lambda(\mathbf u_{1:3}) Q(\mathbf u_{4:6}) \mathbf X) - \mathbf 1_m} ^2 \\[0.8em]
        \text{subject to} & \frac{4}{3} \pi \mathbf u_1 \mathbf u_2 \mathbf u_3 = 1
      \end{array}          
\end{equation}

\section{Outer-Level Problem}
There are two approaches to performing the projection of the ellipsoid into two different views: parametrically and via point-sampling.

\subsection{Parametric Approach}
To perform the projection using parameters obtained from the inner-level problem, we employ orthographic projections. We simulate orthographic projections into two different views by rotating the projected ellipse produced from the inner-level problem differently for each view before projecting the rotated ellipsoids both onto the $x$-$y$ plane.

Two views were chosen as we believed this was sufficient to constrain the problem such that the ellipsoid matched the desired properties.

The output of the inner-level problem gives us a vector 
$$
\begin{array}{ll}
    \mathbf u^* = &\argmin \lr\Vert\Vert {\text{diag}(\mathbf X^\top Q(\mathbf u_{4:6}) ^\top \Lambda(\mathbf u_{1:3}) Q(\mathbf u_{4:6}) \mathbf X) - \mathbf 1_m} ^2 \\[0.5em] &\text{subject to} \frac{4}{3} \pi \mathbf u_1 \mathbf u_2 \mathbf u_3 = 1.
\end{array}          
$$

Using this $\mathbf u^*$, we have the projected ellipsoid with volume 1, and can apply the rotation matrices to the resulting $A(\mathbf u^*) = Q(\mathbf u_{4:6}) ^\top \Lambda(\mathbf u_{1:3}) Q(\mathbf u_{4:6})$. The rotations are performed by two rotation matrices, $\mathbf R_1, \mathbf R_2 \in SO(3)$ respectively. Each are applied as a right conjugation to the matrix $A(\mathbf u^*)$ in a change-of-basis operation.

Thus, we get two rotated ellipsoids, $\mathcal E_1$ and $\mathcal E_2$, parametrised by the following corresponding matrices:
\begin{align}
    \mathcal E_1: \mathbf{R}_1^\top A(\mathbf u^*) \mathbf{R}_1 \qquad \mathcal E_2: \mathbf{R}_2^\top A(\mathbf u^*) \mathbf{R}_2.
\end{align}

To apply the projection onto the $x$-$y$ plane, we can multiply the above matrices by the projection matrix $\mathbf P$,
\begin{equation}
    \mathbf P = \begin{bmatrix}
        1 & 0 & 0\\
        0 & 1 & 0
    \end{bmatrix}.
\end{equation}
To make the matrix multiplication well-defined, we can add a third row of zeroes to $P$.

Thus, the two 2D ellipses $\varepsilon_1$ and $\varepsilon_2$ whose areas we are interested in are parametrised by the matrices $\hat{\mathbf A}_1$ and $\hat{\mathbf A}_2$ (which are effectively 2$\times$2 due to the last row and column being zeroes):
\begin{align}
    \varepsilon_1:\hat{\mathbf A}_1= \mathbf  P \mathbf{R}_1^\top A(\mathbf u^*) \mathbf{R}_1 \qquad \varepsilon_2: \hat{\mathbf A}_2 =\mathbf P \mathbf{R}_2^\top A(\mathbf u^*) \mathbf{R}_2.
\end{align}
The determinants of the upper $2\times 2$ submatrices of $\hat{\mathbf A}_1$ and $\hat{\mathbf A}_2$, denoted as $\hat{\mathbf A}_{1, 2\times2}$ and $\hat{\mathbf A}_{2, 2\times2}$, respectively, would give the product of the eigenvalues of each matrix. As the eigenvalues would be the squared inverses of the semi-axes of $\varepsilon_1$ and $\varepsilon_2$, the inverse square root of the product of eigenvalues gives the product of the semi-axes lengths. Therefore, the area of an ellipse can be expressed as:
\begin{equation}
    \pi ab = \frac{\pi }{\frac{1}{\sqrt{a^2b^2}}}= \frac{\pi }{\sqrt{\lambda _1 \lambda _2}} = \frac{\pi}{\sqrt{\det{\hat{\mathbf A}_i}}}.
\end{equation}

Given two target areas $t_1$ and $t_2$ for the projected ellipses $\varepsilon_1$ and $\varepsilon_2$, our final minimisation problem is thus:
$$
\begin{array}{ll}
  \text{minimize (over $\mathcal X = \{ \mathbf x_i \}$)} &  \lr\Vert\Vert { \frac{\pi}{\sqrt{\det \hat{A}_1(\mathbf u^*)}} - t_1} ^2 + \lr\Vert\Vert{ \frac{\pi}{\sqrt{\det \hat{A}_2(\mathbf u^*)}} - t_2} ^2  \\[0.5em]
  \text{subject to} & \mathbf u^* = \text{argmin} \; \|\lr\Vert\Vert {\text{diag}(\mathbf X^\top Q(\mathbf u_{4:6}) ^\top \Lambda(\mathbf u_{1:3}) Q(\mathbf u_{4:6}) \mathbf X) - \mathbf 1_m} ^2 \\[0.5em]
  & \hphantom{y =} \text{subject to} \;  \frac{4}{3} \pi \mathbf u_1 \mathbf u_2 \mathbf u_3 = 1
\end{array}.
$$

\end{document}

### 2024-09-15 docs/logbook.tex
Got parametric approach done, working on point-sampling now

@@
 \SetProjectTitle{Honours Logbook}
-\SetProjectSubtitle{The subtitle of the project}
+\SetProjectSubtitle{Semester 1 2024 - Semester 1 2025}
 \SetProjectHeaderName{The very fancy project}
@@
 % Below the project there is space to add a brief summary of the project for easy understanding when looking at the front page.
-\SetProjectSummary{Here you can add a brief summary of your project such that someone can quickly see what it is about, do not spend more than two lines on it.}
+\SetProjectSummary{This logbook contains notes and updates on my progress and findings throughout my Honours year.}
 
@@
 % be in one single pdf or png file) can be set using the following macro
-\SetInstitutionLogo{figures/mcod_logo.png}
+\SetInstitutionLogo{assets/ANU_Secondary_Horizontal_GoldBlack.png}
 
@@
 
-% The document is separated into sections, which in turn can be subdivided into subsections
-\section{Overview} \label{sec:overview}
-	% Sections start in a new page, for this reason sections contain clearly disjoint parts of the document
-	This document shows an example of how to use the \texttt{project-logbook} \LaTeX~class.
-This class is based upon the \texttt{article} class and may be used together with other packages for additional functionality.
-This class implements functionality for simplifying writing project/research logbooks and unifying its layout.
-We will  go over the different functionalities to display how to use them.
-
-	\subsection{Preamble}
-		In the preamble of the document you may set several parameters of this document.
-
-		\subsubsection{Maintainer and contributors}
-			Everyone who contributes to this document, should be added as a contributor.
-In this way it is know who contributed and we can also use this to identify who added some elements.
-Adding a contributor is done with \texttt{\textbackslash CreateContributor\string{<key>\string}\string{<name>\string}\string{<id>\string}\string{<affiliation>\string}}, where
-			\begin{description}
-				\item[\texttt{<key>}] is the shorthand used to internally identify a contributor and to retrieve its data, this should be a short word without special characters.
-				\item[\texttt{<name>}] is the full name of the contributor to display on the front page (but may also be used in other situations if desired).
-				\item[\texttt{<id>}] is the abbreviation of the contributor name, to display when associating the contributor to a note (see below) or a meeting minutes box (see below).
-				\item[\texttt{<affiliation>}] is the affiliation of the contributor, mainly to display on the front page.
-    			\end{description}
-
-			This document is maintained by a unique maintainer, which is automatically defined as a contributor with key \texttt{maintainer}.
-The maintainer can be set in the preamble with  \texttt{\textbackslash SetMaintainer\string{<name>\string}\string{<id>\string}\string{<affiliation>\string}}.
-The meaning of the different input parameters is identical to what was seen above for the contributor.
-
-			The maintainer name, id, and affiliation can be accessed via the commands \texttt{\textbackslash MaintainerName}, \texttt{\textbackslash Maintainerid} and \texttt{\textbackslash MaintainerInsitution}, which for this case gives: \MaintainerName~(name), \Maintainerid~(id), and \MaintainerInstitution~(affiliation).
-
-			In a similar way, a contributor can be accessed via the commands \texttt{\textbackslash ContributorName\string{<key>\string}}, \texttt{\textbackslash Contributorid\string{<key>\string}}, and \texttt{\textbackslash ContributorInstitution\string{<key>\string}}, which for this case gives (for key \texttt{chalmers}):  \ContributorName{chalmers}~(name), \Contributorid{chalmers}~(id), and \ContributorInstitution{chalmers}~(affiliation).
-
-		\subsubsection{Project title, subtitle, and header name}
-			The project title (which appears in the center of the front page) can be set in the preamble of the document with \texttt{\textbackslash SetProjectTitle\string{<title>\string}}, where \texttt{<title>} is the title we wish to set.
-
-			Below the project title we can have a project subtitle, which can also be set in the preamble of the document with \texttt{\textbackslash SetProjectSubtitle\string{<subtitle>\string}}, where \texttt{<subtitle>} is the subtitle we wish to set.
-
-		\subsubsection{Maintainer institution logo}
-			On the top right corner of the front page you can add the logo of the institution of the maintainer.
-This can be done in the preamble with the command \texttt{\textbackslash SetInstitutionLogo\string{<path\_to\_file>\string}}, where \texttt{<path\_to\_file>} is the path to a .pdf or .png file.
-If you do not want to show a logo, use a white figure.
-
-	\subsection{Body}
-		The front page is generated with the command \texttt{\textbackslash MakeFrontPage}.
-This automatically adds the table of contents on the second page of the document, with direct links to the sections, for quick access.
-
-		One important section is the Meetings section, which is subdivided into External and Internal.
-The External subsection contains the minutes of the meetings with external partners and the Internal sections contains the internal meetings associated to the institution of the Maintainer.
-Any contributor (including the maintainer) can be an author of Meeting minutes.
-
-		Following the Meetings section, any number of sections can be added focussing on specific topics.
-The goal of these is to discuss the current state of the development of a particular part of the project.
-More discussed below.
-
-		Following these sections, there is an Appendix, containing the list of references (papers to use as reference later when writing a paper), a list of resources such as a link to a video presentation, a tutorial, or something else that is less formal than a reference.
-In the end, you have a list of TODOs that have been defined in the document.
-They appear in the order in which they appeared in the document, not by writing date.
+\section{Overview}\label{sec:overview}
+\begin{HighlightedNote}{Welcome to my progress logbook!}
+	Hello. Topic. Challenges. Research interest.
+\end{HighlightedNote}
 
@@
 Just tested it in code today and it seems that I am minimising the area of the cross-section with the $x$-$y$ plane instead.
-So, just need to take Schur complement instead of projection with matrix P.
+So, just need to take Schur complement instead of projection with matrix $P$.
+
+		One strange observation is that the gradients seem to increase in magnitude as time goes on; not decreasing.
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Sun 15/09/2024}{u7013004}
+		Finally writing about the point-sampling approach. Have been busy with side project website.
 	\end{MeetingMinutes}
@@
 	We can use sections organize the text into different parts  such that specific topics are addressed in their own section.
-
-	\begin{HighlightedNote}{}
-		Sections always start at the top of a new page in order to generate clearly separated parts of the document.
-In this first section (Overview) one should introduce what this project is about, in a high level so you can do a simple explanation of the topic and the research challenges, who is involved, etc.
-
-		Since this information is important, we have added it within a \texttt{HighlightedNote} environment. This is the version without a title.
-	\end{HighlightedNote}
-
-	\begin{HighlightedNote}{You can add multiple things inside a \texttt{HighlightedNote}}
-		And this is the version with a title. You can use the title for example if you want to clearly state what is is, or if you want to make a larger note and give an idea of what it is about.
-
-		You can add equations to clarify your ideas
-		\begin{equation}
-			e^{i\pi} + 1 = 0
-		\end{equation}
-
-		Some tables
-
-		\begin{center}
-			\begin{tabular}{ccccc}
-            			{}  & \multicolumn{1}{c}{$A$} & \multicolumn{1}{c}{$N$} & \multicolumn{1}{c}{$S$} & \multicolumn{1}{c}{$A/S$} \\
-            			\midrule
-            			a & 292 & 117   & 409   & 0.714 \\
-            			b & 104 & 386   & 490   & 0.212 \\
-            			c & 24  & 7     & 31    & 0.774 \\
-            			d & 28  & 27    & 55    & 0.509 \\
-            			e & 183 & 1958  & 2141  & 0.085 \\
-            			\midrule
-            			& 631 & 2495 & 3126 & 0.202 \\
-        			\end{tabular}
-    		\end{center}
-		and other elements.
-
-	\end{HighlightedNote}
-
-	\begin{HighlightedNote}{With a fancy title}
-		The \texttt{HighlightedNotes} environment that generates these fancy boxes, can be used with a title \texttt{\textbackslash begin\string{HighlightedNotes\string}\string{the\_title\_text\string}}, as in this case, or without a title \texttt{\textbackslash begin\string{HighlightedNotes\string}\string{\string}}, as in the previous case.
-	\end{HighlightedNote}
-
-	\subsection{A subsection}
-		You can subdivide your section in a finer level using \texttt{\textbackslash subsection}.
-
-		\subsubsection{A subsubsection}
-			The subsections can be subdivided with \texttt{\textbackslash subsubsection}.
-
-			\subsubsubsection{A subsubsubsection}
-				And an additional level was added to subdivide subsubsections named \texttt{\textbackslash subsubsubsection}.
-This should give more than enough subdivision levels to your text for fine grained organisation.
-
-				\begin{HighlightedNote}{}
-					In most Machine Learning research projects you will be using some kind of samples from a Dataset to learn a model.
-Thus it is extremely important that you carefully describe the dataset and why you believe is a good dataset for the project and what type of preprocessing are you going to apply.
-				\end{HighlightedNote}
-
-				As usual, you can use your references to cite relevant articles as \cite{einstein} and \cite{knuth-fa}.
-
-				You can use tables and, of course, make references to them, like the amazing \tabref{tab:populations}.
-
-				\begin{table}[ht]
-					\centering
-						\begin{subtable}{0.48\textwidth}
-							\centering
-        								\begin{tabular}{*{5}{r}}
-									{}  & \multicolumn{1}{c}{$A$} & \multicolumn{1}{c}{$N$} & \multicolumn{1}{c}{$S$} & \multicolumn{1}{c}{$A/S$} \\
-									\midrule
-									a & 292 & 117   & 409   & 0.714 \\
-									b & 104 & 386   & 490   & 0.212 \\
-									c & 24  & 7     & 31    & 0.774 \\
-									d & 28  & 27    & 55    & 0.509 \\
-									e & 183 & 1958  & 2141  & 0.085 \\
-									\midrule
-									& 631 & 2495 & 3126 & 0.202 \\
-								\end{tabular}
-							\caption{Training Set}
-						\end{subtable}
-    						%
-    						\begin{subtable}{0.48\textwidth}
-        							\centering
-        								\begin{tabular}{*{6}{r}}
-									{}  & \multicolumn{1}{c}{$A'$} & \multicolumn{1}{c}{$N'$} & \multicolumn{1}{c}{$S'$} & \multicolumn{1}{c}{$A'/S'$} & $S'/S$ \\
-									\midrule
-									a & 40  & 40    & 80    & 0.5   & 0.20 \\
-									b & 49  & 49    & 98    & 0.5   & 0.20 \\
-									c & 4   & 3     & 7     & 0.57  & 0.23 \\
-									d & 5   & 5     & 10    & 0.5   & 0.18 \\
-									e & 53  & 53    & 106   & 0.5   & 0.05 \\
-									\midrule
-									& 151 & 150 & 301 & 0.50 & 0.10 \\
-								\end{tabular}
-								\caption{Validation Set}
-						\end{subtable}
-
-    						\caption{Population properties $A \equiv$ Abnormal, $N\equiv$ Normal, $S\equiv A+N$}
-    						\label{tab:populations}
-				\end{table}
-
-	\subsection{Another subsection} \label{sub:another_subsection}
-		Relevant equations can be added
-
-		\begin{equation}
-    			x_i' \gets \frac{x_i-\bar{x}}{\upsigma}
-		\end{equation}
-		\begin{equation}
-    			\bar{x} = \frac{1}{N}\sum_{i=1}^{N} x_i \qquad \upsigma^2 = {\frac{1}{N}\sum_{i=1}^{N} (x_i-\bar{x})}
-		\end{equation}
-
-		\begin{HighlightedNote}{Strike-through}
- 			Often you will be wrong on your assumptions, but do not throw them away completely, just cross them out in case you need them later using the \texttt{\textbackslash sout} command and it will look like \sout{this} or \texttt{\textbackslash soutthick} command and it will look like \soutthick{this}
-		\end{HighlightedNote}
-
-		\sout{Medoid computation is perfomed at $f_s'' = f_s'/5 = \SI{200}{\kilo\hertz}$ to speed computation.
-Simple analysis was performed to check that the features extracted from these 200Hz-medoids were approximately the same as the ones extracted from the  1khz-medoids.}
-
-		\soutthick{Downsampling can be seen as a problem of information loss in the frequency spectrum.
-If the frequency content $f > f_s/(2N)$ is mostly empty for $N$ when we downsample by said $N$ we will only be losing information in that range.}
 

### 2024-09-15 docs/projection.tex
Got parametric approach done, working on point-sampling now

@@
 
-To apply the projection onto the $x$-$y$ plane, we can multiply the above matrices by the projection matrix $\mathbf P$,
+To apply the projection onto the $x$-$y$ plane, it is \textit{not} as simple as multiplying the projection matrix $\mathbf P$,
 \begin{equation}
@@
         0 & 1 & 0
-    \end{bmatrix}.
+    \end{bmatrix},
 \end{equation}
-To make the matrix multiplication well-defined, we can add a third row of zeroes to $P$.
-
-Thus, the two 2D ellipses $\varepsilon_1$ and $\varepsilon_2$ whose areas we are interested in are parametrised by the matrices $\hat{\mathbf A}_1$ and $\hat{\mathbf A}_2$ (which are effectively 2$\times$2 due to the last row and column being zeroes):
+as this will give us the cross-section with the $x$-$y$ plane instead.
+To obtain the projection, we need to take the Schur complement of the bottom-right entry of the 3D matrix.
+Let $S: \R^{3\times 3} \to \R^{2\times 2}$ return the Schur complement described above.
+Then, the two 2D ellipses $\varepsilon_1$ and $\varepsilon_2$ whose areas we are interested in are parametrised by the matrices $\hat{\mathbf A}_1$ and $\hat{\mathbf A}_2$:
 \begin{align}
-    \varepsilon_1:\hat{\mathbf A}_1= \mathbf  P \mathbf{R}_1^\top A(\mathbf u^*) \mathbf{R}_1 \qquad \varepsilon_2: \hat{\mathbf A}_2 =\mathbf P \mathbf{R}_2^\top A(\mathbf u^*) \mathbf{R}_2.
+    \varepsilon_1:\hat{\mathbf A}_1= S\lr(){ \mathbf{R}_1^\top A(\mathbf u^*) \mathbf{R}_1} \qquad \varepsilon_2: \hat{\mathbf A}_2 = S\lr() {\mathbf{R}_2^\top A(\mathbf u^*) \mathbf{R}_2}.
 \end{align}
-The determinants of the upper $2\times 2$ submatrices of $\hat{\mathbf A}_1$ and $\hat{\mathbf A}_2$, denoted as $\hat{\mathbf A}_{1, 2\times2}$ and $\hat{\mathbf A}_{2, 2\times2}$, respectively, would give the product of the eigenvalues of each matrix.
+The determinants of these $2\times 2$ matrices  $\hat{\mathbf A}_1$ and $\hat{\mathbf A}_2$ would give the product of the eigenvalues of each matrix.
 As the eigenvalues would be the squared inverses of the semi-axes of $\varepsilon_1$ and $\varepsilon_2$, the inverse square root of the product of eigenvalues gives the product of the semi-axes lengths.
@@
 \begin{array}{ll}
-  \text{minimize (over $\mathcal X = \{ \mathbf x_i \}$)} &  \lr\Vert\Vert { \frac{\pi}{\sqrt{\det \hat{A}_1(\mathbf u^*)}} - t_1} ^2 + \lr\Vert\Vert{ \frac{\pi}{\sqrt{\det \hat{A}_2(\mathbf u^*)}} - t_2} ^2  \\[0.5em]
-  \text{subject to} & \mathbf u^* = \text{argmin} \; \|\lr\Vert\Vert {\text{diag}(\mathbf X^\top Q(\mathbf u_{4:6}) ^\top \Lambda(\mathbf u_{1:3}) Q(\mathbf u_{4:6}) \mathbf X) - \mathbf 1_m} ^2 \\[0.5em]
+    \underset{\text{over} \mathcal X = \{ \mathbf x_i \}}{\text{minimize}} &  \lr\Vert\Vert { \frac{\pi}{\sqrt{\det \hat{A}_1(\mathbf u^*)}} - t_1} ^2 + \lr\Vert\Vert{ \frac{\pi}{\sqrt{\det \hat{A}_2(\mathbf u^*)}} - t_2} ^2  \\[0.5em]
+  \text{subject to} & \mathbf u^* = \text{argmin} \; \lr\Vert\Vert {\text{diag}(\mathbf X^\top Q(\mathbf u_{4:6}) ^\top \Lambda(\mathbf u_{1:3}) Q(\mathbf u_{4:6}) \mathbf X) - \mathbf 1_m} ^2 \\[0.5em]
   & \hphantom{y =} \text{subject to} \;  \frac{4}{3} \pi \mathbf u_1 \mathbf u_2 \mathbf u_3 = 1
@@
 
+\subsection{Sampling Approach}
+A point-sampling-based approach is more involved, though is more relevant to the water balloon example which this toy example will graduate to.
+
+After obtaining the projected ellipsoid with unit volume,
+
 \end{document}

### 2024-09-22 docs/projection.tex
Corrected objective function

@@
 \section{Goal}
-To project a set of points $\mathcal X$ onto the nearest ellipsoid with volume 1, such that the projection of this ellipsoid in two different views produces ellipses which minimise the error between their areas and two target areas.
+To project a set of points $\mathcal X$ onto the nearest ellipsoid with volume 1, such that the projection of this ellipsoid in two different views produces ellipses which minimise the error to two target ellipses.
+This error can be measured as the error between the areas of the ellipses or their point clouds.
 
@@
 A point-sampling-based approach is more involved, though is more relevant to the water balloon example which this toy example will graduate to.
+In this approach, we are no longer minimising the error between the ellipses' areas and their targets, but matching the ellipses themselves.
+As a result, we need to sample the points of the projected \textit{ellipses} and minimise the difference between this set of points and a target set of points (such as a set of points sampled from a circle).
 
-After obtaining the projected ellipsoid with unit volume,
+In literature, there have been various proposed metrics to measure the difference between two sets of points.
+In particular, these metrics are important for point cloud registration, where the goal is to align multiple point clouds.
+Hausdorff distance has been used previously to evaluate the difference between two point clouds, however, as it is not traditionally differentiable, we will be using Chamfer distance.
+The Chamfer distance $D_C$ between two sets of points $\mathcal{X}$ and $\mathcal{Y}$ is defined as
+\begin{equation}
+    D_C(\mathcal{X,Y})= \frac{1}{|\mathcal{X}|}\sum_{x \in \mathcal{X}} d(x, \mathcal Y) + \frac{1}{|\mathcal{Y}|}\sum_{y \in \mathcal{Y}} d(y, \mathcal X),
+\end{equation}
+where $d(a, \mathcal S)$ for a point $a$ and a set $\mathcal S$ is defined as $\min d(a,s)$ for all $s \in \mathcal S$.
+In other words, this gives the sum of mean distance between all pairs of points from $\mathcal X$ to $\mathcal Y$ and the opposite.
+As Euclidean distance is known to be sensitive to outliers, other variations like density-aware and hyperbolic Chamfer distance have been researched.
+
+Conceptually, after projected the rotated unit volume ellipsoid onto the $x$-$y$ plane, we sample around these 2D ellipses.
+We then calculate the Chamfer distance of these sampled points to our ground truth points at each step and attempt to minimise this distance through backpropagation.
+
+The matter of sampling the points poses issues if not done carefully as stochastic nodes cannot be backpropagated through.
+We take inspiration from variational autoencoders (VAE) and apply the "reparametrisation trick" such that the random variable can be expressed in terms of deterministic nodes through which the gradient can flow through and a stochastic node to perform the sampling.
+
+Let the projected ellipses $\varepsilon_1$ and $\varepsilon_2$ be represented by the matrices $\hat{\mathbf A}_1$ and $\hat{\mathbf A}_2$, respectively, as above.
+For these ellipses to be well-defined, both these matrices must be symmetric positive definite.
+As proven by Horn \& Johnson, there exists a unique positive definite square root $B=A^{1/2}$ for a positive definite and symmetric matrix $A$ such that $A= A^{1/2} A^{1/2}  = BB$.
+Furthermore, as it is positive definite, $A^{1/2}$ is invertible -- that is, $A^{-1/2}$ exists.
+
+Thus, the reparametrisation trick applied is to decompose $\hat{\mathbf A}_i$ into its square roots like so:
+\begin{align}
+    x^\top \hat{\mathbf A}_i x = 1 &\implies x^\top \hat{\mathbf A}^{1/2}_i \hat{\mathbf A}^{1/2}_i x = 1 \\
+    &\implies z^\top z = 1,
+\end{align}
+where we let $z = \hat{\mathbf A}^{1/2}_i x$.
+Thus, we simply need to sample unit vectors $z$ (or vectors on the unit circle) and can recover the sampled point $x$ by applying the inverse square root.
+That is, $x = \hat{\mathbf A}^{-1/2}z$.
+
+Let the sampled points of the projected ellipses be denoted by $\mathcal X_1(\mathbf u^*)$ and $\mathcal X_2(\mathbf u^*)$ for $\varepsilon_1$ and $\varepsilon_2$, respectively, using the reparametrisation trick.
+Hence, given two sets of ground truth points $\mathcal Y_1, \mathcal Y_2$ for the projected ellipses $\varepsilon_1$ and $\varepsilon_2$, our final minimisation problem is
+$$
+\begin{array}{ll}
+    \underset{\text{over} \mathcal X = \{ \mathbf x_i \}}{\text{minimize}} &  d_C(\mathcal X_1(\mathbf u^*), \mathcal Y_1) + d_C(\mathcal X_2(\mathbf u^*), \mathcal Y_2) \\[0.5em]
+  \text{subject to} & \mathbf u^* = \text{argmin} \; \lr\Vert\Vert {\text{diag}(\mathbf X^\top Q(\mathbf u_{4:6}) ^\top \Lambda(\mathbf u_{1:3}) Q(\mathbf u_{4:6}) \mathbf X) - \mathbf 1_m} ^2 \\[0.5em]
+  & \hphantom{y =} \text{subject to} \;  \frac{4}{3} \pi \mathbf u_1 \mathbf u_2 \mathbf u_3 = 1
+\end{array}.
+$$
 

### 2024-09-23 docs/logbook.tex
Corrected objective function
Changed to float64, achieves slightly more spherical results

@@
 		Finally writing about the point-sampling approach. Have been busy with side project website.
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Fri 20/09/2024}{u7013004}
+		Trying to add a regulariser between iterations for lower problem.
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Sun 22/09/2024}{u7013004}
+		Going to fully do the regulariser and report results. Will also provide warm-starting with previous solution. 
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Mon 23/09/2024}{u7013004}
+		Tried switching to float64 and it's not working. Going to try add more views. For 
 	\end{MeetingMinutes}

### 2024-10-21 docs/logbook.tex
attempting alpha

@@
 \documentclass{project-logbook}
-
+\input{commands}
 \usepackage{csquotes}
@@
 
-\section{Topic A to focus on} \label{sec:topic_A}
+	\begin{MeetingMinutes}{Sun 29/09/2024}{u7013004}
+		Worked on Chamfer distance and it worked! Forgot to take inverse of square root matrix but figured it all out after reading my projection notes.
+
+		Going to write my ellipsoid stuff in this logbook until I have to copy it over into my thesis.
+
+		\lowtodo{2024-09-29}{u7013004}{Figure out Blender stuff, proposal, and whether to use other nearest-neighbour method.}
+
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Sun 6/10/2024}{u7013004}
+		Wrote up some stuff for ellipsoid. Still unsure how to tackle boundary?
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Mon 21/10/2024}{u7013004}
+		Basically spent 2 weeks without doing work. Time to do more! Going to attempt the alpha shapes approach. Hopefully converting back and forth from tensor to numpy won't do weird things.
+	\end{MeetingMinutes}
+
+\section{Ellipsoid Formulation} \label{sec:ellipsoid}
+An ellipsoid is a 3-dimensional surface obtained by deforming a sphere under an affine transformation.
+As ellipsoid surfaces can be represented as the zero set of a quadratic polynomial in three variables, it is considered a quadric surface.
+
+There are multiple mathematical representations of ellipsoids.
+Ellipsoids have three pairwise perpendicular axes of symmetry known as principal axes which intersect at the centre of the ellipsoid.
+The most general ellipsoid is a triaxial ellipsoid, where each of these three principal axes are of different lengths.
+The line segments along these axes from the origin to the surface boundary are named the principal semi-axes as they are half the length of the principal axes.
+An ellipsoid centred at the origin and with its axes aligned to the standard basis of $\R^3$ is commonly represented in Cartesian coordinates as the points $(x,y,z) \in \R^3$ which satisfy
+\begin{equation}
+	\frac{x^2}{a^2} + \frac{y^2}{b^2} + \frac{z^2}{c^2} = 1,
+\end{equation}
+where $a$, $b$, and $c$ are the lengths of the semi-axes. If we let $\mathbf x = [x, y, z]^\top \in \R^3$, then we can represent this expression becomes
+\begin{equation}
+	\mathbf x^\top \mathbf A \mathbf x = 1,
+\end{equation}
+where $$\mathbf A = \begin{bmatrix}
+	1/a^2 & 0 & 0 \\
+	0 & 1/b^2 & 0 \\
+	0 & 0 & 1/c^2
+\end{bmatrix}.$$
+However, ellipsoids need not be axis-aligned with respect to the standard basis.
+Rather than having $\mathbf A$ be a diagonal matrix with the inverse squared semi-axes lengths, we can generalise ellipsoids to be non-axis-aligned by letting $\mathbf A$ be any symmetric positive-definite matrix -- that is, having $\mathbf A \in \mathbb S^3_{++}$ such that $\mathbf A = \mathbf A^\top \succ 0$.
+This represents an ellipsoid $\mathcal E$ as the convex set
+\begin{equation}
+	\mathcal E  = \lrbrace{\mathbf x \in \R^3 \mid \mathbf x^\top \mathbf A \mathbf x = 1, \mathbf A \in \mathbb S^3_{++} }.
+\end{equation}
+
+We can further perform an eigendecomposition on $\mathbf A$ to express it in terms of an orthonormal rotation matrix $\mathbf R \in SO(3)$ and a diagonal matrix $\mathbf \Lambda$ containing the eigenvalues of $\mathbf A$, which are the inverse squares of the semi-axes lengths:
+\begin{equation}
+	\mathbf A = \mathbf Q^\top \mathbf\Lambda \mathbf Q,
+\end{equation}
+where $\mathbf \Lambda = \begin{bmatrix}
+	1/a^2 & 0 & 0 \\
+	0 & 1/b^2 & 0 \\
+	0 & 0 & 1/c^2
+\end{bmatrix}$
+and thus 
+\begin{equation}
+	\mathcal E  = \lrbrace{\mathbf x \in \R^3 \mid \mathbf x^\top \mathbf Q^\top \mathbf\Lambda \mathbf Q \mathbf x = 1, \mathbf Q \in SO(3)}.
+\end{equation}
+
+\section{Project Book Components} \label{sec:topic_B}
+
 	We can use sections organize the text into different parts  such that specific topics are addressed in their own section.
@@
 		\hightodo{2016-05-27}{maintainer}{Perform convergence analysis with the latest version of the code.}
-
-\section{Topic B to focus on} \label{sec:topic_B}
 
@@
 
-		% \begin{figure}[h]
-    	% 		\includegraphics[width=0.9\textwidth]{figures/LSTM}
-    	% 		\centering
-    	% 		\caption{Gate Recurrent Unit in a Long Short Term Memory Neural Netwok (GRU-LSTM). \\ \emph{Credit to \url{https://colah.github.io/posts/2015-08-Understanding-LSTMs/}}}
-    	% 		\label{fig:LSTM}
-		% \end{figure}
-
-\section{Topic C to focus on} \label{sec:topic_c}
 		\subsection{Figures} \label{sub:figures}
@@
 
+		% \begin{figure}[h]
+    	% 		\includegraphics[width=0.9\textwidth]{figures/LSTM}
+    	% 		\centering
+    	% 		\caption{Gate Recurrent Unit in a Long Short Term Memory Neural Netwok (GRU-LSTM). \\ \emph{Credit to \url{https://colah.github.io/posts/2015-08-Understanding-LSTMs/}}}
+    	% 		\label{fig:LSTM}
+		% \end{figure}
+
+\section{Topic C to focus on} \label{sec:topic_c}
 % subsection tables (end)

### 2025-02-23 docs/logbook.tex
Updated mesh explanation

@@
 
+	\begin{MeetingMinutes}{Thu 23/01/2025}{u7013004}
+		Not sure how the mesh will be able to handle concavities if only the boundary points are detected in the upper-level loss, these will only be able to handle convex parts.
+I guess that the lower-level loss will handle that by moving in those concave sections in more.
+	\end{MeetingMinutes}
+
 \section{Ellipsoid Formulation} \label{sec:ellipsoid}
@@
 \end{equation}
+
+\section{Deep Declarative Nodes} \label{sec:ddn}
+Deep neural networks are themselves optimisation problems, where a loss function is minimised over the network weights.
+Backpropagation enables this optimisation to occur efficiently.
+However, incorporating constraints into a neural network complicates this procedure as constraints need to be respected in the optimisation process.
+Furthermore, constrained optimisation is often iterative and not naturally compatible with the differentiable nature of the current neural networks.
+
+Deep declarative networks (DDNs) provide a new framework to achieve backpropagation through certain constrained optimisation problems.
+The conditions for these problems are that we need the optimal solution to the optimisation problem and twice-differentiable objective and constriant functions, and thus the method used to find the optimal solution can be non-differentiable, such as RANSAC.
 

### 2025-02-23 docs/projection.tex
Updated mesh explanation

@@
 
-Conceptually, after projected the rotated unit volume ellipsoid onto the $x$-$y$ plane, we sample around these 2D ellipses.
+Conceptually, after projecting the rotated unit volume ellipsoid onto the $x$-$y$ plane, we sample around these 2D ellipses.
 We then calculate the Chamfer distance of these sampled points to our ground truth points at each step and attempt to minimise this distance through backpropagation.
@@
 
+\section{Adopting a Mesh-Based Approach}
+Now, as our water balloon may not remain conform to an ellipsoid, we must represent it more generally. We do this by using a mesh based representation.
+
+Let $\mathcal M = (V,E)$ represent the mesh.
+Let $\text{Vol}(\mathcal M)$ denote the volume of such a mesh.
+Then, the first constraint which we must change is $\frac{4}{3} \pi \mathbf u_1 \mathbf u_2 \mathbf u_3 = 1$, which now becomes $\text{Vol}(\mathcal M) = 1$.
+
+Next, we replace $\lr\Vert\Vert {\text{diag}(\mathbf X^\top Q(\mathbf u_{4:6}) ^\top \Lambda(\mathbf u_{1:3}) Q(\mathbf u_{4:6}) \mathbf X) - \mathbf 1_m} ^2$ with another least squares function.
+However, this time, we pair up points $V$ of the mesh with vertices $U$ in the ground-truth mesh we obtain from the physical simulation to form corresponding pairs $(v_i, u_i)$.
+The replacement least squares function is thus $\sum_{i=1}^{|V| = |U|}\lr\Vert\Vert {v_i - u_i}^2$.
+
+For our upper-level loss function, we can still perform a Chamfer distance to compare with the ground-truth points of the mesh from $k$ different views, $\mathcal Y_k$, with these same views of our mesh $\mathcal X_k(U)$.
+
+Though this can be confusing, if we take a step back, conceptually, the components of our bi-level optimisation are:
+$$
+\begin{array}{ll}
+    \underset{}{\text{minimize}} &  \text{difference between current points $V^*$ and target} \\[0.5em]
+  \text{subject to} & V^* = \text{argmin} \; \text{projection of current points onto intended shape} \\[0.5em]
+  & \hphantom{y =} \text{subject to} \;  \text{volume constraint}
+\end{array}
+$$
+
+This makes it very easy to substitute in our modifications:
+$$
+\begin{array}{ll}
+    \underset{\text{over } \mathcal X = \lrbrace{\mathbf x_i} }{\text{minimize}} &  \sum_{k'=1}^k d_C(\mathcal X_k(V^*), \mathcal Y_k) \\[0.5em]
+  \text{subject to} & V^* = \text{argmin} \; \sum_{i \in [|\mathcal X|]} ||v_i - u_i||^2, \quad v_i \in V, u_i \in U \\[0.5em]
+  & \hphantom{y =} \text{subject to} \;  \text{Vol}(\mathcal M) = 1
+\end{array}
+$$
+
+So the difference between the upper and lower-level losses are that the upper-level function is moving the points $\mathcal X$ by minimising differences in 2D projections of the mesh (so only boundary points have gradient flowing through them), whereas the lower-level loss minimises the 3D distance between all points on the mesh with the corresponding pairs.
+
+Additional considerations are now whether the mesh is self-intersecting and ensuring it is watertight; two factors which were implicit in assuming an ellipsoid representation.
+We need to handle these somehow later on.
+
 \end{document}

### 2025-02-28 docs/logbook.tex
Updated logbook

@@
 I guess that the lower-level loss will handle that by moving in those concave sections in more.
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Wed 26/02/2025}{u7013004}
+		Have explored many many options for the physics simulation.
+I will list them out.
+The aim is to get good data by Friday.
+I will aim by tonight to have tried various options and have thought about which one is best.
+		The options are:
+		\begin{enumerate}
+			\item NVIDIA Omniverse - pros: minimal effort to set up inflatable, can set up scene easily, cons: unsure how mesh would be extracted, I will need access to Pranav's computer
+			\item Feather - pros: physics looks robust, easily usable interface, should be easily to extract mesh and it's in C++, can simulate interactions, cons: will need to manually set up environment, port to Blender and render there, requires Linux
+			\item Q-Minh: pros: good interface, physics looks good, should be easy to extract mesh, in C++, cons: need to port to Blender
+			\item \sout{liujustin53: unknown, will have to try} BAD!
+			\item PBS-soft-body (ETH): can interact with rigid body if the rigid body is in the scene and not moving
+			\item PBDTaichi: Python, pretty basic, can't handle collisions with other objects
+		\end{enumerate}
+
+		FML. Feather can only work on OpenGL 4.3 so I'll try on Windows at home.
+
+		Can also ask Jiahao about the apt stuff for libgl and other RandR (RandR library and headers were not found).
+
+		Can use the PBAToolkit in Python, so that should be doable on Mac. Same with some of the other ones.
+
+		The Feather one, I have almost no hope for. I can try to run it on the GPU but the OpenGL is so far behind.
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Thu 27/02/2025}{u7013004}
+		This really sucks right now bruh. I am bumping into dead ends and I wonder if the only thing I can use is Pranav's computer atm.
+		Going to try vcpkg again for Feather and Q-Minh (he replied and said he hasn't ported it over to Python yet but can do it next week).
+		In the meantime, I can try liujustin53, PBS-soft-body (ETH).
+		Verdict: liujustin53 can't do collisions.
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Fri 28/02/2025}{u7013004}
+		Ok, need to ask if I can somehow use a computer with a GPU. If not, rent one?
+		NVIDIA Omniverse has everything, I just need to be able to use it.
+
+		My options otherwise are PBDTaichi, PBS-soft-body, and Feather.
+		All can do volume preservation.\\
+
+		Feather is super slow but can do bending and length constraints, and can handle collisions with other objects. Investigate solver iteration.
+
+		PBDTaichi has bending and length constraints, but cannot handle collisions with other objects.
+
+		PBS-soft-body has length and friction constraints, can handle collisions with static objects in the scene, cannot handle collisions with other objects.
+
+		For now, I need some really basic data to get the pipeline working.
+
+		Discussion with Steve: he'll ask IT, I will generate a simple example in Blender first.
+
 	\end{MeetingMinutes}

### 2025-02-28 docs/projection.tex
Updated logbook

@@
 \begin{array}{ll}
-    \underset{}{\text{minimize}} &  \text{difference between current points $V^*$ and target} \\[0.5em]
-  \text{subject to} & V^* = \text{argmin} \; \text{projection of current points onto intended shape} \\[0.5em]
+    \underset{}{\text{minimize}} &  \text{error between projected mesh points $V^*$ and target} \\[0.5em]
+  \text{subject to} & \text{projected mesh points } V^* = \text{argmin} \; \text{nearest points to existing mesh} \\[0.5em]
   & \hphantom{y =} \text{subject to} \;  \text{volume constraint}

### 2025-03-08 docs/logbook.tex
Implemented signed volume function
Updated logbook

@@
 
-		Discussion with Steve: he'll ask IT, I will generate a simple example in Blender first.
+		Discussion with Steve: he'll ask IT, I will generate a simple example in Blender first. Also code up the volume.
 
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Mon 3/3/2025}{u7013004}
+		Yesterday, I worked on the three Blender shapes. I have three equations which produce surfaces of volume 1.4 pi. The next step is to visualise these in Blender.
+
+		I also tested the correctness of my signed volume function and it works correctly. As the mesh resolution increases, the volume approximates the true volume better and better.
+
+		Going to write about my background.
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Tue 4/3/2025}{u7013004}
+		Made the mesh fairly high res, most of the volume is lost lengthways, so focus on upping resolution for that. Was able to render them.
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Wed 5/03/2025}{u7013004}
+		Get the camera matrix.
+		Get the Canny edge detector.
+		After that, tidy up point cloud section and talk about how to perform surface reconstruction.
 	\end{MeetingMinutes}

### 2025-03-12 docs/logbook.tex
Got the projection code finished

@@
 		After that, tidy up point cloud section and talk about how to perform surface reconstruction.
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Sat 8/3/2025}{u7013004}
+		Important to have no nothing days. Today, I will write more about SfM in my thesis. The goal of the long weekend will be to have results of the mesh changing.
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Wed 12/03/2025}{u7013004}
+		Turns out the SciPy was hanging because it takes 4 minutes per iteration when using a subdivision level of 4.
+Takes 10 seconds if subdivision level of 3, and basically instant with subdivision level of 2.
+I think including the manually providing the gradients should speed up the computation.
+
+		Today, I want to write the code for the node and make sure that it works for subdiv levels 2 and 3.
+After that, I want to write the Chamfer distance loss function and hook it up.
+Then I'd like to visualise the results.
+Should be good!
 	\end{MeetingMinutes}

### 2025-03-12 src/edgedetection.py

@@ -1,0 +1,38 @@
+import cv2
+import sys
+
+def apply_canny(input_path, output_path=None, threshold1=100, threshold2=200):
+    # Read the image in grayscale
+    img = cv2.imread(input_path, cv2.IMREAD_GRAYSCALE)
+    
+    if img is None:
+        print(f"Error: Unable to read image {input_path}")
+        return
+
+    # Apply Canny edge detection        
+    img = cv2.equalizeHist(img)
+    edges = cv2.Canny(img, threshold1, threshold2)
+
+    if output_path:
+        # Save the result
+        cv2.imwrite(output_path, edges)
+        print(f"Saved edge-detected image to {output_path}")
+    else:
+        # Show the result in a window
+        cv2.imshow("Canny Edge Detection", edges)
+        cv2.waitKey(0)
+        cv2.destroyAllWindows()
+
+if __name__ == "__main__":
+    if len(sys.argv) < 2:
+        print("Usage: python canny_edge.py <input_image> [output_image] [threshold1] [threshold2]")
+        sys.exit(1)
+    
+    input_file = sys.argv[1]
+    output_file = sys.argv[2] if len(sys.argv) > 2 and sys.argv[2] != "-" else None  # Use "-" to skip saving
+
+    # Optional thresholds
+    thresh1 = int(sys.argv[3]) if len(sys.argv) > 3 else 100
+    thresh2 = int(sys.argv[4]) if len(sys.argv) > 4 else 200
+
+    apply_canny(input_file, output_file, thresh1, thresh2)

### 2025-03-12 src/notebooks/testcrash.py

@@ -1,0 +1,224 @@
+import math
+import numpy as np
+from torch import cos, sin
+import scipy.optimize as opt
+import torch
+import torch.nn as nn
+import matplotlib.pyplot as plt
+from sklearn.decomposition import PCA
+import psutil
+import os
+
+process = psutil.Process(os.getpid())
+
+import warnings
+warnings.filterwarnings('ignore')
+
+import sys
+sys.path.append("../../../ddn/")
+from ddn.pytorch.node import *
+
+from pytorch3d.loss import chamfer_distance
+from pytorch3d.ops import sample_farthest_points
+import alphashape
+from descartes import PolygonPatch
+import pytorch3d
+
+from pytorch3d.io import IO
+import plotly.graph_objects as go
+from pytorch3d.vis.plotly_vis import AxisArgs, plot_batch_individually, plot_scene
+from plotly.subplots import make_subplots
+from pytorch3d.renderer import (
+    HeterogeneousRayBundle,
+    ray_bundle_to_ray_points,
+    RayBundle,
+    TexturesAtlas,
+    TexturesVertex,
+)
+from pytorch3d.renderer.camera_utils import camera_to_eye_at_up
+from pytorch3d.renderer.cameras import CamerasBase
+from pytorch3d.structures import join_meshes_as_scene, Meshes, Pointclouds
+
+import os
+import torch
+from pytorch3d.io import load_obj, save_obj,load_objs_as_meshes
+from pytorch3d.structures import Meshes
+from pytorch3d.utils import ico_sphere
+from pytorch3d.ops import sample_points_from_meshes
+from pytorch3d.loss import (
+    chamfer_distance, 
+    mesh_edge_loss, 
+    mesh_laplacian_smoothing, 
+    mesh_normal_consistency,
+)
+
+import faulthandler
+faulthandler.enable()
+
+def get_signed_tet_volume(face_vertices: torch.Tensor) -> torch.Tensor:
+    """
+    Compute signed tetrahedron volumes for a batch of faces.
+
+    Args:
+        face_vertices (torch.Tensor): Tensor of shape (F, 3, 3), where
+                                      F is the number of faces, and each face
+                                      consists of 3 vertices in 3D.
+
+    Returns:
+        torch.Tensor: A tensor of shape (F,) containing signed volumes.
+    """
[…]

### 2025-03-12 src/test.py

@@ -1,0 +1,260 @@
+import numpy as np
+import torch
+import trimesh
+import matplotlib.pyplot as plt
+from scipy.optimize import minimize
+
+def load_meshes(source_path, target_path):
+    """
+    Load source and target meshes from OBJ files.
+    
+    Args:
+        source_path: Path to source mesh OBJ file
+        target_path: Path to target mesh OBJ file
+    
+    Returns:
+        source_vertices: Vertices of source mesh
+        source_faces: Faces of source mesh
+        target_vertices: Vertices of target mesh
+        target_faces: Faces of target mesh
+    """
+    source_mesh = trimesh.load_mesh(source_path)
+    target_mesh = trimesh.load_mesh(target_path)
+    print(source_mesh.vertices.shape)
+    # Verify that meshes have the same number of vertices
+    if source_mesh.vertices.shape[0] != target_mesh.vertices.shape[0]:
+        raise ValueError("Meshes must have the same number of vertices")
+    
+    # Verify that meshes have the same number of faces
+    if source_mesh.faces.shape[0] != target_mesh.faces.shape[0]:
+        raise ValueError("Meshes must have the same number of faces")
+    return source_mesh.vertices, np.array(source_mesh.faces), target_mesh.vertices, np.array(target_mesh.faces)
+
+def calculate_volume_torch(vertices_torch, faces):
+    """
+    Calculate the volume of a mesh using PyTorch tensors.
+    
+    Args:
+        vertices_torch: Nx3 tensor of vertex coordinates
+        faces: Mx3 array of face indices
+    
+    Returns:
+        volume: Total volume of the mesh as a PyTorch scalar
+    """
+    volume = torch.tensor(0.0, requires_grad=True)
+    
+    for face in faces:
+        v1, v2, v3 = [vertices_torch[idx] for idx in face]
+        # Calculate cross product v2 × v3
+        cross_product = torch.cross(v2, v3)
+        # Calculate dot product v1 · (v2 × v3)
+        triple_product = torch.dot(v1, cross_product)
+        # Volume of tetrahedron = (1/6) * |triple scalar product|
+        tetrahedron_volume = torch.abs(triple_product) / 6.0
+        volume = volume + tetrahedron_volume
+    
+    return volume
+
+def objective_function(x, target_vertices, num_vertices):
+    """
+    Calculate the squared sum of differences between vertices.
+    
[…]

### 2025-03-14 src/edgedetection.py
Finished pytorch chamfer loss module

@@ -1,7 +1,7 @@
 import cv2
 import sys
 
-def apply_canny(input_path, output_path=None, threshold1=100, threshold2=200):
+def apply_canny(input_path, output_path=None, threshold1=20, threshold2=250):
     # Read the image in grayscale
     img = cv2.imread(input_path, cv2.IMREAD_GRAYSCALE)
     

### 2025-03-14 src/notebooks/utils.py

@@ -1,0 +1,89 @@
+import torch
+import numpy as np
+
+
+import os
+import re
+from collections import defaultdict
+import cv2
+
+def read_matrices_from_file(filename):
+    matrices = {}
+    current_cam = None
+    
+    with open(filename, "r") as f:
+        lines = f.readlines()
+        
+    i = 0
+    while i < len(lines):
+        line = lines[i].strip()
+        
+        if line.startswith("Camera:"):
+            current_cam = line.split("Camera: ")[1]
+            matrices[current_cam] = {}
+        elif any(matrix_type in line for matrix_type in ["K (Intrinsic", "RT (Extrinsic", "P (Projection"]):
+            matrix_type = line.split(" ")[0]  
+            i += 1
+            matrices[current_cam][matrix_type] = torch.tensor(np.loadtxt(lines[i:i+3]), dtype=torch.float32)
+            i += 2  
+        i += 1
+        
+    return matrices
+
+def load_camera_matrices(path):
+    cameras = defaultdict(dict)
+    file_pattern = re.compile(r"^(.*)_(K|RT|P)\.npy$")
+    for filename in os.listdir(path):
+        match = file_pattern.match(filename)
+        if match:
+            print(filename)
+            cam_name, matrix_type = match.groups()
+            filepath = os.path.join(path, filename)
+            cameras[cam_name][matrix_type] = np.load(filepath)
+    return cameras
+    
+
+def load_renders(renders_path):
+    renders = defaultdict(list)
+    pattern = re.compile(r"([a-zA-Z]+)(\d+)\.png")
+    for filename in sorted(os.listdir(renders_path), key=lambda x: (re.match(pattern, x).group(1), int(re.match(pattern, x).group(2))) if re.match(pattern, x) else (x, float('inf'))):
+        match = pattern.match(filename)
+        if match:
+            word, number = match.groups()
+            number = int(number)  # Convert number to integer for sorting
+
+            # Load image (PIL) or store filename
+            image_path = os.path.join(renders_path, filename)
+            image = cv2.imread(image_path)  # Load image using OpenCV
+            if image is not None:
+                renders[word].append(image)
+    return renders
+
+
+
+def project_batched_vertices(self, vertices):
+    """
+    Projects batched sets of vertices into multiple views using a shared set of projection matrices.
+
+    Args:
[…]

### 2025-03-18 docs/papers.tex
Updated mesh explanation
Worked on mesh pipeline and read papers

@@
         \item This method is used for tasks where we deform a 3D mesh $\mathcal{M} = (V,F)$, where $V = \lrbrace{\bm v_1, \bm v_2,...}$ denotes vertices in $\R^3$ and $F$ facets to minimise a task-specific loss function $\mathcal L_{\text {task}}(\mathcal M)$.
-        \item Signed distance function (SDF): $s: \R^3 \to \R$. Neural network $f_\theta$ is tarined on watertight surfaces $\mathcal S$ to approximate SDF $s$ by minimising
+        \item Signed distance function (SDF): $s: \R^3 \to \R$. Neural network $f_\theta$ is trained on watertight surfaces $\mathcal S$ to approximate SDF $s$ by minimising
         \begin{equation*}
@@
 \begin{ArticleSummary}{DRWR: A Differentiable Renderer without Rendering for Unsupervised 3D Strucutre Learning }
-    \textbf{Citation:} Han \textit{et al.}
+    \textbf{Citation:} Han \textit{et al.}, 2020
 
@@
         \item This is unsupervised, so only requires images for training.
-        \item Existing pixel-wise loss functions render images from 3D structures but require rasterisation or pixel-wise interpolation commonly.
+        \item Existing pixel-wise loss functions render images from 3D structures but require rasterisation or pixel-wise interpolation commonly, which they claim don't actually contribute to structure learning
         \item DRWR: doesn't require visibility handling, shading, or pixel-wise interpolation.
@@
 
+\begin{ArticleSummary}{Multi-view Convolutional Neural Networks for 3D Shape Recognition}
+    \textbf{Citation:} Su et al., 2015
+
+    \textbf{Summary:} Classifies 3D objects using 2D projections of said object from different views using a CNN, outperforms previous approaches which had access to the 3D information.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item They can outperform 3D based methods as 1) 2D representations are less memory expensive compared to voxel representations and thus training is more efficient, 2) image descriptors have advanced to be more information dense, 3) there exist massive image databases which are well-annotated
+        \item Rather than averaging views, the multi-view CNN is able to identify more informative views and combine these to form the 3D object (think about the pooling layers).
+        \item ``Jittering'' used to augment data and provide analogous samples to changing the view.
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{GRF: Learning a General Radiance Field for 3D Representation and Rendering}
+    \textbf{Citation:} Trevithick and Yang, 2021
+
+    \textbf{Summary:} Takes a set of 2D images with known camera poses and intrinsics, constructs 3D rendering from arbitrary positions using neural network, uses attention mechanism from differing views to account for occlusions.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Builds upon NeRF, can infer from unobserved viewpoints. Can generalise unlike NeRF to any set of images to a 3D structure with geometric details.
+        \item Exploits the underlying 3D geometry unlike previous approaches which generated images from unseen views by learning a 2D manifold.
+        \item Four components: 1) feature extractor from 2D pixels using CNN, 2) reprojector to transform into 3D space, 3) aggreggator to obtain general features for 3D points, 4) NeRF to render internal representation
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{Pixel2Mesh: Generating 3D Mesh Models
+    from Single RGB Images}
+    \textbf{Citation:} Wang et al., 2018
+
+    \textbf{Summary:} Graph-based CNN which learns deformations (not the structure itself) of an ellipsoid tempplate to output a 3D mesh from a single colour image.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Coarse-to-fine strategy for stability by graph unpooling (typical division strategy used in computer graphics where new vertex is midpoint of an edge, generating 4 new triangles), uses many surface and geometric losses to improve realism and fidelity
+        \item graph-based convolutional network used to model the mesh with nodes = vertices and connections = edges, use a pooling layer of 2D images in the GCN to incorporate features/knowledge in 3D construction
+        \item Multi-view geometry is restricted by the coverage of views (cannot generalise to unseen viewpoints) and appearance (cannot generate non-Lambertian surfaces), hence why learning approaches are important
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{AtlasNet: A Papier-M\^ach\'e Approach to Learning 3D Surface Generation}
+    \textbf{Citation:} Groueix et al., 2018
+
+    \textbf{Summary:} From a latent shape vector, sample a point on a unit square, the network produces a 3D point with the square deformed in 3D to be a continuous planar surface (locally 2D manifold) allowing arbitrary resolution, and repeat this for multiple patches.
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{Mesh R-CNN}
+    \textbf{Citation:} Gkioxari, Malik, and Johnson, 2019
+
+    \textbf{Summary:} Mix of 2D perception and 3D shape prediction.
+Takes input image and predicts meshes of variable topology by first predicting coarse voxel structure, converting to mesh, and then refining the mesh with graph convolution network.
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{Deep Marching Tetrahedra: a Hybrid Representation for High-Resolution 3D Shape Synthesis}
+    \textbf{Citation:} Shen et al., 2021
+
+    \textbf{Summary:} Generative model to generate 3D shapes via user guided input (such as coarse voxels), hybrid implementation where explicit representation used to allow finer geometric features to be rendered and implicit representation (deformable tetrahedral grid) to allow arbitrary topology.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Uses Marching Tetrahedra algorithm which is differentiable unlike MC, adapts resolution by deforming the tetrahedra grid --- the singularity for linearly interpolating the crossing remains but in practice only evaluated when vertex sign is different
+        \item Input is point cloud or low-res voxel which gets converted into a feature vector of the 3D volume, 
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{Shape as Points: a Differentiable Poisson Solver}
+    \textbf{Citation:} Peng et al., 2021
+
+    \textbf{Summary:} Differentiable Poisson surface reconstruction method used in both optimisation and learning-based scenarios (noisy point clouds). Render using MC.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Optimisation-based: noisy oriented point cloud as input, pass this into DPSR to obtain dense indicator grid, then sample points and use MC to generate a mesh which can be compared using Chamfer to target.
+MC not differentiable, but gradient approximated using inverse normals of mesh sampled vertices (from MeshSDF).
+        \item Learning -based: learns features/normals from a point cloud using an MLP and then calculating an indicator grid using DPSR and comparing it to the target indicator grid to supervise learning.
+        \item Doees not require normals like traditional PSR.
+        \item Poisson equation intuition: the gradient of the indicator function image is approximated by point normals.
+        \item Discretizes allowing inversion of divergence operator, use spectral methods to solve the equation (FFT highly optimised on GPUs).
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{Convolutional Occupancy Networks}
+    \textbf{Citation:} Peng et al., 2020
+
+    \textbf{Summary:} A new implicit representation (compared to MLPs which fail to encode local information/structured reasoning) which uses convolutional encoders and then a decoder to get a probabilistic occupancy grid.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Converts 3D input (point cloud or voxel) into features, run features through CNN onto occupancy grid output
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{Multiview Neural Surface Reconstruction by Disentangling Geometry and Appearance}
+    \textbf{Citation:} Yariv et al., 2020
+
+    \textbf{Summary:} Separates geometry from appearance from camera parameters to construct 3D scenes from input images (even with variable lighting conditions).
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Input: masked 2D images (indicator function approximated using sigmoid for differentiability) and camera estimates.
+        \item Differentiable renderer which is able to render the surface of a shape represented by the zero level set of the implicit representation (neural network).
+        \item Able to render material/colour alongside reconstruct shape.
+        \item Uses differentiable ray casting, parametrises the intersection of the ray with the surface of the object with the geometry and camera parameters.
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{Scene Representation Networks: Continuous 3D-Structure-Aware Neural Scene Representations}
+    \textbf{Citation:} Sitzmann, Zollhofer, Wetzstein
+
+    \textbf{Summary:} Differentiable ray-marching algorithm trained on 2D images and camera poses, outputs 3D scene (geometry + appearance).
+Maps 3D world coordinate to feature representation of scene properties at that coordinate.
+Does not require 3D supervision.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Can generate images without 2D convolutions - not a 3D shape reconstruction, but still involves a 3D implicit representation to generate images from novel views.
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{Differentiable Volumetric Rendering:
+    Learning Implicit 3D Representations without 3D Supervision}
+    \textbf{Citation:} Niemeyer et al., 2020
+
+    \textbf{Summary:} Differentiable renderer for implicit shapes (and texture).
+Depth gradients obtained with implicit differentiation.
+Input 2D posed images.
+Outputs depth map from occupancy network, which is unprojected into 3D and rendered to be compared with ground truth image.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item 2D supervision (depth maps or multi-vew images) attempts to overcome learning-based 3D reconstruction approaches which are restricted to synthetic data (and restricts their prior knowledge) during supervision.
+However many suffer from discretisation artifacts (meshes and voxels).
+        \item Mesh based methods require deformable template mesh or a collection of 3D patches which leads to self-intersections and non-watertight meshes
+        \item uses occupancy network to assign a probability of occupancy to each point, and texture (colour) uses a texture field
+        \item able to learn without 3D ground truth data
+        \item key contribution is analytical gradient to optimise for the surface depth and overcome non-differentiable argmin and iterative secant method
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{Octree Generating Networks: Efficient Convolutional Architectures for High-resolution 3D Outputs}
+    \textbf{Citation:} Tatarchenko, Dosovitskiy, and Brox, 2017
+
+    \textbf{Summary:} Convolutional decoder which outputs higher resolution octree from high-level feature descriptions and reconstructing 3D objects from single images.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Uses hash table rather than storing cell pointers using an index-value pair $(m,v)$ where $m$ generated from spatial coordinate and level using Z-order curves and $v$ generated any signal.
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{PointNet: Deep Learning on Point Sets for 3D Classification and Segmentation}
+    \textbf{Citation:} Qi et al., 2017
+
+    \textbf{Summary:} Takes point clouds and outputs classification, segmentation, and semantic parsing.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Uses symmetric function of max pooling for permutation invariance (as point clouds are a set)
+        \item Network learns set of functions which select informative points of the point cloud and encode reason for selection
+        \item Network should be invariant to ordering and transformations such as rotation or translation
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{DUSt3R: Geometric 3D Vision Made Easy}
+    \textbf{Citation:} Wang et al., 2024
+
+    \textbf{Summary:} Multi-view stereo reconstruction in the wild, from unconstrained image collection, can output pointmaps from which camera parameters, depthmaps, and 3D reconstruction can be recovered
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Unlike most approaches which first attempt to estimate camera parameters, this work performs regression of pointmaps to relax hard constraints of projective camera models
+        \item Unconstrained image-based dense 3D reconstruction form multiple views subsumes many other geometric 3D vision tasks
+        \item Attempts to overcome challenge of existing approaches which are only as good as the quality of input images and camera parameters 
+        \item Jointly processes input images and 3D pointmaps (dense 2D field of 3D points) in order to associate 2D structures with 3D shapes
+        \item Need to apply MC, PSR, or other learning-based methods to extract mesh from pointmap
+        \item From a pair of images, encodes them using ViT (patchify) with shared weights, but then decoded separately, with the second decoder also optimising for relative pose to image 1.
+Outputs point map and confidence map for each head.
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{NeRF: Representing Scenes as Neural Radiance Fields for View Synthesis}
+    \textbf{Citation:} Mildenhall et al., 2020
+
+    \textbf{Summary:} Trains one MLP per scene with input being single 5D coordinate (x,y,z and viewing direction $\theta, \phi$) and output is volume density and view-dependent radiance (colour and lighting).
+Loss between rendered images and observed image inputs.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Uses positional encoding with sine and cosines to take input vectors into a higher frequency dimension, allowing finer details to be generated
+        \item Estimates radiance using quadrature (whatever that means)
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{Pix2Vox++: Multi-scale Context-aware 3D Object Reconstruction from Single and Multiple Images}
+    \textbf{Citation:} Xie et al., 2020
+
+    \textbf{Summary:} Takes multi-view images to produce coarse voxel strucures which are then fused adn refined
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{3D-R2N2: A Unified Approach for Single and Multi-view 3D Object Reconstruction}
+    \textbf{Citation:} Choy et al., 2016
+
+    \textbf{Summary:} Uses Recurrent NN to map image/s to reconstructed 3D shape in the form of a voxelised occupancy grid.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Convolution NN with many LSTMs with restricted connections for encoding, decoder is 3D deconvolution network, uses softmax voxel-wise loss
+        \item 
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{BundleSDF: Neural 6-DoF Tracking and 3D Reconstruction of Unknown Objects}
+    \textbf{Citation:} Wen et al., 2023
+
+    \textbf{Summary:} Near real-time tracking method of objects (6 DoF) from a video sequence whilst performing 3D reconstruction. Only need to be segmented in the first frame.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item 1) Match features between consecutive segmented video frames to obtain rough pose estimates of the object, some which are stored in memory.
+2) Pose graph constructed from memory pool 3) Online optimisation of pose graph with current pose, then stored back in memory pool, 4) All posed frames in memory pool used to train Neural Object Field (geometry is an SDF and also visual appearance/texture)
+        \item Uses octree for (hierarchical) ray sampling
+    \end{itemize}
+\end{ArticleSummary}
+
+\begin{ArticleSummary}{TopoNet: Topology Learning for 3D Reconstruction of Objects of Arbitrary Genus}
+    \textbf{Citation:} Charrada et al., 2022
+
+    \textbf{Summary:} Takes RGB image and reconstructs 3D mesh by deforming a template mesh and then performs face-pruning to refine the mesh and allow arbitrary genus changing using an RL network.
+
+    \textbf{Key Points:}
+    \begin{itemize}
+        \item Conv Graph Net to extract features of initial mesh and deform it
+        \item Does not look watertight bro
+    \end{itemize}
+\end{ArticleSummary}
+
 \end{document}

### 2025-03-18 src/notebooks/utils.py

@@ -1,7 +1,5 @@
 import torch
 import numpy as np
-
-
 import os
 import re
 from collections import defaultdict
@@ -30,21 +28,37 @@ def read_matrices_from_file(filename):
         
     return matrices
 
-def load_camera_matrices(path):
+def load_camera_matrices(path, matrix_types=None):
+    """
+    Loads camera matrices from .npy files in the specified directory.
+
+    Parameters:
+        path (str): Path to the directory containing camera matrix files.
+        matrix_types (set or list, optional): Specifies which matrix types to load (e.g., {'K', 'RT'}).
+                                              If None, all available matrices ('K', 'RT', 'P') will be loaded.
+
+    Returns:
+        dict: A dictionary mapping camera names to their respective matrices.
+    """
     cameras = defaultdict(dict)
     file_pattern = re.compile(r"^(.*)_(K|RT|P)\.npy$")
-    for filename in os.listdir(path):
+    
+    if matrix_types is not None:
+        matrix_types = set(matrix_types)  # Ensure it's a set for quick lookup
+    
+    for filename in sorted(os.listdir(path)):  # Sort filenames alphabetically
         match = file_pattern.match(filename)
         if match:
-            print(filename)
             cam_name, matrix_type = match.groups()
-            filepath = os.path.join(path, filename)
-            cameras[cam_name][matrix_type] = np.load(filepath)
-    return cameras
+            if matrix_types is None or matrix_type in matrix_types:
+                filepath = os.path.join(path, filename)
+                cameras[cam_name][matrix_type] = torch.tensor(np.load(filepath))
     
+    return cameras
+
 
 def load_renders(renders_path):
-    renders = defaultdict(list)
+    renders = defaultdict(dict)
     pattern = re.compile(r"([a-zA-Z]+)(\d+)\.png")
     for filename in sorted(os.listdir(renders_path), key=lambda x: (re.match(pattern, x).group(1), int(re.match(pattern, x).group(2))) if re.match(pattern, x) else (x, float('inf'))):
         match = pattern.match(filename)
@@ -52,11 +66,11 @@ def load_renders(renders_path):
             word, number = match.groups()
             number = int(number)  # Convert number to integer for sorting
 
-            # Load image (PIL) or store filename
             image_path = os.path.join(renders_path, filename)
             image = cv2.imread(image_path)  # Load image using OpenCV
             if image is not None:
-                renders[word].append(image)
+                renders[word][number] = image  # Store image in the nested dictionary
+
     return renders
 
 

### 2025-04-01 docs/logbook.tex
Worked on mesh pipeline and read papers
updated logbook

@@
 Should be good!
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Fri 14/3/2025}{u7013004}
+		Possibly learn the offsets so I can use offset\_verts but that requires a packed tensor I believe. Might as well just start a new set of vertices I reckon.
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Sun 30/3/2025}{u7013004}
+		The edge loss didn't work on Friday. Let's think about why.
+
+		If X is the result after Chamfer loss backwards, then it won't conform to the volume constraints.
+If Y is the malleable mesh, then decreasing SSE between Y and X will decrease the error in positions of Y and X subjecting Y to have constant volume.
+Preserving the edge length constraint can result in Y being as close to X.
+As a result, the solver might just choose to focus on minimising SSE and edge length constraints could assist in that.
+
+		Okay, the issue is that if I do Edge loss by itself, the gradient blows up.
+Why?
+Because the Hessian (double derivative of objective with respect to Y twice) is apparently incredibly ill-conditioned.
+ChatGPT says it's because of
+
+		2.1 Edge Lengths Close to Their Targets (Small Gradients): if $d(Y_i, Y_j)^2 \approx d(X_i, X_j)^2$ then the scaling term $d(Y_i, Y_j)^2 - d(X_i, X_j)^2$ is small.
+This means the Hessian matrix has many small entries, leading to near-zero eigenvalues and an ill-conditioned system.
+
+		2.2 Highly Connected Meshes (Near-Dependent Constraints): If a vertex has many connected neighbors, its position is constrained by multiple edges.
+This causes rows of the Hessian to be nearly linearly dependent, leading to a large condition number.
+
+		2.3 The Rigid Motion Null Space: The edge length term does not penalize rigid transformations, meaning there are six zero eigenvalues (three for translation, three for rotation), making 
+		$f_{YY}$ singular
+	\end{MeetingMinutes}
+
+	\begin{MeetingMinutes}{Tue 1/4/2025}{u7013004}
+		Wtf it's the 1st of April. After the past four days, I've found that the edge length constraint is kinda bad and not worth spending more time on. 
+		The next phase is to keep writing my thesis, and to write scripts for average performance, and also test a little bit of robustness with edges today. I can plan out my day.
+
+		For my thesis, I would like to plan out my roadmap and write my abstract. My dad would like me to write a problem definition, idk if I'll use it but I'll do it to make him happy. 
 	\end{MeetingMinutes}

### 2025-04-01 src/edgedetection.py

@@ -1,16 +1,19 @@
 import cv2
 import sys
 
-def apply_canny(input_path, output_path=None, threshold1=20, threshold2=250):
+def apply_canny(input_path, output_path=None, threshold1=20, threshold2=250, equalize=False):
     # Read the image in grayscale
     img = cv2.imread(input_path, cv2.IMREAD_GRAYSCALE)
     
     if img is None:
         print(f"Error: Unable to read image {input_path}")
         return
-
+    
+    # Apply histogram equalization if the flag is set
+    if equalize:
+        img = cv2.equalizeHist(img)
+    
     # Apply Canny edge detection        
-    img = cv2.equalizeHist(img)
     edges = cv2.Canny(img, threshold1, threshold2)
 
     if output_path:
@@ -25,7 +28,7 @@ def apply_canny(input_path, output_path=None, threshold1=20, threshold2=250):
 
 if __name__ == "__main__":
     if len(sys.argv) < 2:
-        print("Usage: python canny_edge.py <input_image> [output_image] [threshold1] [threshold2]")
+        print("Usage: python canny_edge.py <input_image> [output_image] [threshold1] [threshold2] [--equalize]")
         sys.exit(1)
     
     input_file = sys.argv[1]
@@ -34,5 +37,8 @@ if __name__ == "__main__":
     # Optional thresholds
     thresh1 = int(sys.argv[3]) if len(sys.argv) > 3 else 100
     thresh2 = int(sys.argv[4]) if len(sys.argv) > 4 else 200
-
-    apply_canny(input_file, output_file, thresh1, thresh2)
+    
+    # Check if equalization flag is provided
+    equalize = "--equalize" in sys.argv or "-e" in sys.argv
+    
+    apply_canny(input_file, output_file, thresh1, thresh2, equalize)

### 2025-04-06 src/experiments/chamfer.py
Ported functions to .py files

@@ -1,0 +1,95 @@
+from alpha_shapes import *
+import shapely
+import torch
+from pytorch3d import Meshes, chamfer_distance
+from torch import nn
+
+def get_boundary(projected_pts, alpha=10.0):
+    # Create a detached copy for Alpha_Shaper
+    projected_pts_detached = projected_pts.detach().clone()
+    
+    # Use the detached copy with Alpha_Shaper
+    shaper = Alpha_Shaper(projected_pts_detached)
+    alpha_shape = shaper.get_shape(alpha)
+    while isinstance(alpha_shape, shapely.MultiPolygon) or isinstance(alpha_shape, shapely.GeometryCollection):
+        alpha -= 1
+        alpha_shape = shaper.get_shape(alpha)
+    boundary = torch.tensor(alpha_shape.exterior.coords.xy, dtype=torch.double)
+    
+    # Find indices of boundary points
+    boundary_indices = torch.where(
+        torch.any(torch.isclose(projected_pts_detached[:, None], boundary.T, atol=1e-6).all(dim=-1), dim=1)
+    )[0]
+    
+    # Index back into the original tensor with gradients
+    boundary_pts = projected_pts[boundary_indices]
+    return boundary_pts
+
+class PyTorchChamferLoss(nn.Module):
+    def __init__(self, src: Meshes, tgt: Meshes, projmatrices, edgemap_info):
+        super().__init__()
+        self.src = src  # (B meshes)
+        self.tgt = tgt  # (B meshes)
+        self.projmatrices = projmatrices # (P, 3, 4)
+        self.edgemaps = edgemap_info[0] # (P, max_Ni, 2)
+        self.edgemaps_len = edgemap_info[1] # (P,)
+    
+    def project_vertices(self, vertices):
+        """
+        Projects a set of vertices into multiple views using different projection matrices.
+        Args:
+            vertices: Tensor of shape (N, 3), representing 3D vertex positions.
+        Returns:
+            Tensor of shape (P, N, 2), containing projected 2D points in each view.
+        """
+        V = vertices.shape[0]
+        projection_matrices = self.projmatrices
+
+        ones = torch.ones((V, 1), dtype=vertices.dtype, device=vertices.device)
+        vertices_homogeneous = torch.cat([vertices, ones], dim=1).double()  # Shape: (V, 4)
+
+        # Perform batched matrix multiplication (P, 3, 4) @ (V, 4, 1) -> (P, V, 3)
[…]

### 2025-04-06 src/experiments/functions.py

@@ -1,0 +1,131 @@
+import torch
+from pytorch3d import Meshes, chamfer_distance
+import trimesh
+
+torch.autograd.set_detect_anomaly(True)
+
+def least_squares(u0, tgt_vtxs):
+    """
+    u0 are vertices
+    """
+    if not torch.is_tensor(u0):
+        u0 = torch.tensor(u0)
+    if not torch.is_tensor(tgt_vtxs):
+        tgt_vtxs = torch.tensor(tgt_vtxs)
+    res = torch.square(u0 - tgt_vtxs.flatten()).sum()
+    return res.double()
+
+def least_squares_grad(u0, tgt_vtxs):
+    if torch.is_tensor(u0):
+        u0 = u0.detach().clone()
+    else:
+        u0 = torch.tensor(u0)
+    if torch.is_tensor(tgt_vtxs):
+        tgt_vtxs = tgt_vtxs.detach().clone()
+    else:
+        tgt_vtxs = torch.tensor(tgt_vtxs)
+        
+    # Ensure that u0 requires gradients
+    gradient = 2 * (u0 - tgt_vtxs.flatten())
+    return gradient.double()
+
+
+def calculate_volume(vertices, faces):
+    face_vertices = vertices[faces]  # (F, 3, 3)
+    v0, v1, v2 = face_vertices[:, 0, :], face_vertices[:, 1, :], face_vertices[:, 2, :]
+    
+    # Compute determinant of the 3x3 matrix [v0, v1, v2]
+    face_volumes = torch.det(torch.stack([v0, v1, v2], dim=-1)) / 6.0  # Shape: (F,)
+    volume = face_volumes.sum()
+    return volume.abs()
+
+
+def volume_constraint(x, faces, tgt_vol):
+    """
+    Calculate the volume of a mesh using PyTorch tensors.
+    Args:
+        vertices_torch: Nx3 tensor of vertex coordinates
+        faces: Mx3 array of face indices
+    Returns:
+        volume: Total volume of the mesh as a PyTorch scalar
+    """
+    if not torch.is_tensor(x):
+        x = torch.tensor(x)
+    if not torch.is_tensor(faces):
+        faces = torch.tensor(faces)
+    if not torch.is_tensor(tgt_vol):
+        tgt_vol = torch.tensor(tgt_vol)
+
+    vertices = x.view(-1,3)
+    faces = faces.view(-1,3).int()
+    volume = calculate_volume(vertices, faces)
+    res = volume.abs() - tgt_vol
[…]

### 2025-04-06 src/experiments/io.py

@@ -1,0 +1,109 @@
+import torch
+import numpy as np
+import os
+import re
+from collections import defaultdict
+import cv2
+from cv2.typing import MatLike
+
+def load_camera_matrices(path, matrix_types=None):
+    """
+    Loads camera matrices from .npy files in the specified directory.
+
+    Parameters:
+        path (str): Path to the directory containing camera matrix files.
+        matrix_types (set or list, optional): Specifies which matrix types to load (e.g., {'K', 'RT'}).
+        If None, all available matrices ('K', 'RT', 'P') will be loaded.
+
+    Returns:
+        dict: A dictionary mapping camera numbers to their respective matrices.
+    """
+    cameras = defaultdict(dict)
+    file_pattern = re.compile(r"^Camera_(\d+)_(K|RT|P)\.npy$")
+    
+    if matrix_types is not None:
+        matrix_types = set(matrix_types)  # Ensure it's a set for quick lookup
+    
+    for filename in sorted(os.listdir(path)):  # Sort filenames alphabetically
+        match = file_pattern.match(filename)
+        if match:
+            cam_number, matrix_type = match.groups()
+            cam_number = int(cam_number)  # Convert camera number to integer
+            if matrix_types is None or matrix_type in matrix_types:
+                filepath = os.path.join(path, filename)
+                cameras[cam_number][matrix_type] = torch.tensor(np.load(filepath))
+    
+    return cameras
+
+def load_renders(renders_path):
+    renders = defaultdict(dict)
+    pattern = re.compile(r"([a-zA-Z]+)(\d+)\.png")
+    for filename in sorted(os.listdir(renders_path), key=lambda x: (re.match(pattern, x).group(1), int(re.match(pattern, x).group(2))) if re.match(pattern, x) else (x, float('inf'))):
+        match = pattern.match(filename)
+        if match:
+            word, number = match.groups()
+            number = int(number)  # Convert number to integer for sorting
+
+            image_path = os.path.join(renders_path, filename)
+            image = cv2.imread(image_path)  # Load image using OpenCV
+            if image is not None:
+                renders[word][number] = image  # Store image in the nested dictionary
+
+    return renders
+
+def get_projmats_and_edgemap_info(view_idx, target_mesh: str, matrices, edgemaps, edgemaps_len):
+    """
[…]

### 2025-04-06 src/experiments/node.py

@@ -1,0 +1,127 @@
+import torch
+from pytorch3d import Meshes
+from ddn import EqConstDeclarativeNode, DeclarativeFunction
+from functions import *
+from utils import *
+import scipy.optimize as opt
+
+class ConstrainedProjectionNode(EqConstDeclarativeNode):
+    """
+    Performs a projection of the input points X onto the nearest points Y such that the volume of Y is constant.
+    """
+    def __init__(self, src: Meshes, tgt: Meshes):
+        super().__init__(eps=1.0e-6) # relax tolerance on optimality test 
+        self.src = src # source meshes (B,)
+        # self.tgt = tgt # target meshes (B,)
+        self.b = len(src)
+
+    def objective(self, xs: torch.Tensor, y: torch.Tensor, scatter_add=False):
+        """
+        Calculates sum of squared differences between source and target meshes.
+
+        Args:
+            xs (torch.Tensor): a padded (B, max Vi, 3) tensor of the original vertices
+            y (torch.Tensor): a padded (B, max Vi, 3) tensor of the projected vertices        
+        """
+        src_verts = padded_to_packed(xs, self.src.num_verts_per_mesh()).view(-1,3)
+        tgt_verts = padded_to_packed(y, self.src.num_verts_per_mesh()).view(-1,3)
+
+        sqr_diffs = torch.square(src_verts - tgt_verts).sum(dim=-1) # (sum(V_i))
+        n_batches = len(self.src)
+        sse = torch.zeros(n_batches, dtype=sqr_diffs.dtype)
+        if scatter_add:
+            sse.scatter_add_(0, self.src.verts_packed_to_mesh_idx(), sqr_diffs)
+        else:
+            n_verts_per_mesh = self.src.num_verts_per_mesh()
+            for i in range(n_batches):
+                mesh_to_vert = self.src.mesh_to_verts_packed_first_idx()  # Index of first face per mesh
+                start = mesh_to_vert[i]
+                end = start + n_verts_per_mesh[i]
+                sse[i] = sqr_diffs[start:end].sum()  # Sum over all faces
+        return sse
+
+    def equality_constraints(self, xs, y, scatter_add=False):
+        """
+        Enforces volume constraint of projected points
+        Assumes same number of vertices in each projected mesh currently
+
+        Args:
+            xs (torch.Tensor): a padded (B, max Vi, 3) tensor of the original vertices
[…]

### 2025-04-06 src/experiments/utils.py

@@ -1,0 +1,77 @@
+import torch
+import matplotlib.pyplot as plt
+from chamfer import get_boundary
+
+def padded_to_packed(xs, lengths):
+    packed = []
+    batch_size = xs.size(0)
+    for b in range(batch_size):
+        n = lengths[b]
+        packed.append(xs[b][:n])
+    packed = torch.cat(packed, dim=0)
+    return packed
+
+def create_padded_tensor(vertices, vert2mesh, max_V, B):
+    padded = torch.zeros((B, max_V, 3),device=vertices.device)
+    for i in range(B):
+        mesh_vertices = vertices[vert2mesh == i]
+        num_vertices = mesh_vertices.shape[0]
+        padded[i, :num_vertices, :] = mesh_vertices
+    return padded
+
+def plot_vertices(verts_list):
+    fig = plt.figure(figsize=(5, 5))
+    ax = fig.add_subplot(111, projection='3d')
+    
+    if not isinstance(verts_list, list):
+        x, y, z = verts_list.clone().detach().cpu().squeeze().unbind(1)
+        ax.scatter3D(x, z, -y)
+    else:
+        colors = ['b',  'g', 'r' , 'm', 'c', 'y']  # Define some colors for different sets
+        marker_size = 5  # Make points smaller
+    
+        for i, verts in enumerate(verts_list):
+            x, y, z = verts.clone().detach().cpu().squeeze().unbind(1)
+            ax.scatter3D(x, z, -y, color=colors[i % len(colors)], s=marker_size, label=f"Set {i+1}")    
+    
+    ax.set_xlabel('x')
+    ax.set_ylabel('z')
+    ax.set_zlabel('y')
+    ax.set_aspect("equal")
+    ax.view_init(190, 30)
+    ax.legend()
+    plt.show()
+
+def plot_projections(projverts, projmats, edgemaps):
+    plt.ioff()  # Disable interactive mode
+
+    P, _, _ = projmats.shape
+    edge_coords, edge_lens = edgemaps
+    fig, axes = plt.subplots(1, P, figsize=(2 * P, 2))  # Increase figure size
+    if P == 1:
+        axes = [axes]  # Ensure iterable for a single subplot case
+
+    for i in range(P):
+        proj_2d_hom = (projmats[i] @ torch.cat([projverts, torch.ones(projverts.shape[0], 1)], dim=1).T).T
[…]

### 2025-04-06 src/chamfer.py
Wrote experiment files

@@ -1,7 +1,8 @@
 from alpha_shapes import *
 import shapely
 import torch
-from pytorch3d import Meshes, chamfer_distance
+from pytorch3d.structures import Meshes
+from pytorch3d.loss import chamfer_distance
 from torch import nn
 
 def get_boundary(projected_pts, alpha=10.0):

### 2025-04-06 src/dataloader.py

@@ -1,0 +1,40 @@
+import json
+import os
+from pytorch3d.io import load_obj
+from pytorch3d.structures import Meshes
+from .io import *
+
+class DataLoader:
+    def __init__(self, config_path: str) -> None:
+        with open(config_path, "r") as f:
+            self.cfg = json.load(f)
+
+        # Set up paths and other config settings
+        self.mesh_dir = self.cfg["paths"]["mesh_dir"]
+        self.mesh_res = self.cfg["paths"]["mesh_res"]
+        self.renders_path = self.cfg["paths"]["renders_path"]
+        self.matrices_path = self.cfg["paths"]["matrices_path"]
+        
+        self.edgemap_options = {mesh["name"]: {int(k): v for k,v in mesh["edgemap_options"].items()} for mesh in self.cfg["meshes"]}
+        
+        self.camera_matrices = load_camera_matrices(path=self.matrices_path, matrix_types="P")
+
+        renders = load_renders(self.renders_path)
+        self.edgemaps, self.edgemaps_len = load_edgemaps(renders, self.edgemap_options)
+
+        # Load meshes using PyTorch3D
+        self.meshes = self.load_meshes()
+
+    def load_meshes(self):
+        """
+        Load meshes using PyTorch3D's load_obj function.
+        Returns a dictionary of PyTorch3D Meshes objects.
+        """
+        meshes = {}
+        for mesh_info in self.cfg["meshes"]:
+            mesh_name = mesh_info["name"]
+            mesh_path = os.path.join(self.mesh_dir, f"{mesh_name}_{self.mesh_res}.obj")
+            verts, faces, aux = load_obj(mesh_path)
+            # Create Meshes object in PyTorch3D
+            meshes[mesh_name] = Meshes(verts=[verts], faces=[faces])
+        return meshes

### 2025-04-06 src/experimentrunner.py

@@ -1,0 +1,122 @@
+from .dataloader import DataLoader
+import json
+from .node import *
+from .chamfer import *
+from .utils import *
+from typing import List, Tuple
+import random
+
+class ExperimentRunner:
+    def __init__(self, config_path: str, data_loader: DataLoader) -> None:
+    # Load the experiment config from a JSON file
+        with open(config_path, "r") as f:
+            self.cfg = json.load(f)
+        
+        # Store the experiment details
+        self.experiment_name = self.cfg["name"]
+        self.experiment_description = self.cfg["description"]
+        self.src_mesh = self.cfg["src_mesh"]
+        self.target_meshes = self.cfg["target_meshes"]
+        
+        # Parse views configuration for each target mesh
+        self.views_config = {}
+        for mesh_name, view_config in self.cfg["views"].items():
+            self.views_config[mesh_name] = {
+                "mode": view_config["mode"],
+                "view_idx": view_config["view_idx"],
+                "num_views": view_config["num_views"]
+            }
+        
+        # Training settings
+        self.n_iters = self.cfg["training"]["n_iters"]
+        self.lr = self.cfg["training"]["lr"]
+        self.momentum = self.cfg["training"]["momentum"]
+        
+        # Use the DataLoader to load data
+        self.data_loader = data_loader
+
+    def run(self):
+        return self.pipeline(self.src_mesh, self.target_meshes)
+
+    def pipeline(self, src_name: str, tgt_names: List[str]):
+        """Pipeline function to run the experiment on a source mesh and a list of target meshes."""
+        
+        src_mesh = self.get_mesh(src_name)
+        for tgt_name in tgt_names:
+            tgt_mesh = self.get_mesh(tgt_name)
+
+            # Select the view points and get corresponding projmats
+            projmats, tgt_edgemap_info, view_idx = self.get_projmats_and_edgemap_info(tgt_name)
+            batch_info = ([tgt_edgemap_info[0]], [tgt_edgemap_info[1]])
+
+            # Train the mesh transformation
+            final_verts = self.train_loop(src_mesh, tgt_mesh, projmats, batch_info, n_iters=self.n_iters, lr=self.lr, moment=self.momentum)
+
+            # Update source mesh for warmstarting
[…]

### 2025-04-06 src/functions.py

@@ -1,5 +1,6 @@
 import torch
-from pytorch3d import Meshes, chamfer_distance
+from pytorch3d.structures import Meshes
+from pytorch3d.loss import chamfer_distance
 import trimesh
 
 torch.autograd.set_detect_anomaly(True)

### 2025-04-06 src/io.py

@@ -90,20 +90,33 @@ def canny_edge_map(img: MatLike, options):
     # return edge map
     return edge_map
 
-def get_edgemaps(renders, options):
+def load_edgemaps(renders, options):
     edgemaps = {}
     edgemaps_len = {}
-    for k,v in renders.items():
+
+    # Iterate through the renders dictionary, which is grouped by mesh names
+    for mesh_name, mesh_renders in renders.items():
         views = {}
         views_len = {}
-        for num, img in v.items():
-            edges = canny_edge_map(img, options[num])
-            edge_coords = np.argwhere(edges > 0)
-            edge_coords = edge_coords[:, [1,0]]  
-            views[num] = torch.tensor(edge_coords)
-            views_len[num] = len(edge_coords)
-        edgemaps[k] = views
-        edgemaps_len[k] = views_len
+        
+        # Find the edgemap options for the current mesh name
+        if mesh_name in options:
+            edgemap_options = options[mesh_name]
+
+            # Iterate through each render (view) for the current mesh
+            for num, img in mesh_renders.items():
+                # Apply the corresponding edgemap options for this render number (num)
+                if num in edgemap_options:
+                    edge_option = edgemap_options[num]
+                    edges = canny_edge_map(img, edge_option)
+                    edge_coords = np.argwhere(edges > 0)
+                    edge_coords = edge_coords[:, [1, 0]]  # Convert to (x, y) coordinates
+                    
+                    views[num] = torch.tensor(edge_coords)
+                    views_len[num] = len(edge_coords)
+
+        # Store the results for the current mesh
+        edgemaps[mesh_name] = views
+        edgemaps_len[mesh_name] = views_len
+
     return edgemaps, edgemaps_len
-
-

### 2025-04-06 src/node.py

@@ -1,18 +1,17 @@
 import torch
-from pytorch3d import Meshes
+from pytorch3d.structures import Meshes
 from ddn import EqConstDeclarativeNode, DeclarativeFunction
-from functions import *
-from utils import *
+from .functions import *
+from .utils import *
 import scipy.optimize as opt
 
 class ConstrainedProjectionNode(EqConstDeclarativeNode):
     """
     Performs a projection of the input points X onto the nearest points Y such that the volume of Y is constant.
     """
-    def __init__(self, src: Meshes, tgt: Meshes):
+    def __init__(self, src: Meshes):
         super().__init__(eps=1.0e-6) # relax tolerance on optimality test 
         self.src = src # source meshes (B,)
-        # self.tgt = tgt # target meshes (B,)
         self.b = len(src)
 
     def objective(self, xs: torch.Tensor, y: torch.Tensor, scatter_add=False):

### 2025-04-06 src/test.py

@@ -1,260 +1,0 @@
-import numpy as np
-import torch
-import trimesh
-import matplotlib.pyplot as plt
-from scipy.optimize import minimize
-
-def load_meshes(source_path, target_path):
-    """
-    Load source and target meshes from OBJ files.
-    
-    Args:
-        source_path: Path to source mesh OBJ file
-        target_path: Path to target mesh OBJ file
-    
-    Returns:
-        source_vertices: Vertices of source mesh
-        source_faces: Faces of source mesh
-        target_vertices: Vertices of target mesh
-        target_faces: Faces of target mesh
-    """
-    source_mesh = trimesh.load_mesh(source_path)
-    target_mesh = trimesh.load_mesh(target_path)
-    print(source_mesh.vertices.shape)
-    # Verify that meshes have the same number of vertices
-    if source_mesh.vertices.shape[0] != target_mesh.vertices.shape[0]:
-        raise ValueError("Meshes must have the same number of vertices")
-    
-    # Verify that meshes have the same number of faces
-    if source_mesh.faces.shape[0] != target_mesh.faces.shape[0]:
-        raise ValueError("Meshes must have the same number of faces")
-    return source_mesh.vertices, np.array(source_mesh.faces), target_mesh.vertices, np.array(target_mesh.faces)
-
-def calculate_volume_torch(vertices_torch, faces):
-    """
-    Calculate the volume of a mesh using PyTorch tensors.
-    
-    Args:
-        vertices_torch: Nx3 tensor of vertex coordinates
-        faces: Mx3 array of face indices
-    
-    Returns:
-        volume: Total volume of the mesh as a PyTorch scalar
-    """
-    volume = torch.tensor(0.0, requires_grad=True)
-    
-    for face in faces:
-        v1, v2, v3 = [vertices_torch[idx] for idx in face]
-        # Calculate cross product v2 × v3
-        cross_product = torch.cross(v2, v3)
-        # Calculate dot product v1 · (v2 × v3)
-        triple_product = torch.dot(v1, cross_product)
-        # Volume of tetrahedron = (1/6) * |triple scalar product|
-        tetrahedron_volume = torch.abs(triple_product) / 6.0
-        volume = volume + tetrahedron_volume
-    
-    return volume
-
-def objective_function(x, target_vertices, num_vertices):
-    """
-    Calculate the squared sum of differences between vertices.
-    
[…]

### 2025-04-06 src/utils.py

@@ -1,6 +1,8 @@
 import torch
 import matplotlib.pyplot as plt
 from chamfer import get_boundary
+import plotly.graph_objects as go
+import numpy as np
 
 def padded_to_packed(xs, lengths):
     packed = []
@@ -75,3 +77,58 @@ def plot_projections(projverts, projmats, edgemaps):
     plt.subplots_adjust(wspace=0.1)  # Reduce horizontal space
     plt.show()
     plt.close(fig)
+
+def visualise_meshes(srcmesh, tgtmesh):
+    vertices = np.asarray(srcmesh.verts_packed())
+    faces = np.asarray(srcmesh.faces_packed())
+
+# Create a Plotly 3D mesh
+    fig = go.Figure(data=[go.Mesh3d(
+        x=vertices[:, 0],
+        y=vertices[:, 1],
+        z=vertices[:, 2],
+        i=faces[:, 0],
+        j=faces[:, 1],
+        k=faces[:, 2],
+        opacity=0.5,
+        color="lightblue"
+    )])
+
+    vertices = np.asarray(tgtmesh.verts_packed())
+    faces = np.asarray(tgtmesh.faces_packed())
+
+    fig.add_trace(go.Mesh3d(
+        x=vertices[:, 0],
+        y=vertices[:, 1],
+        z=vertices[:, 2],
+        i=faces[:, 0],
+        j=faces[:, 1],
+        k=faces[:, 2],
+        opacity=0.1,
+        color="red"
+    ))
+
+    # Update layout for better presentation
+    fig.update_layout(
+        scene=dict(
+            xaxis=dict(title="X"),
+            yaxis=dict(title="Y"),
+            zaxis=dict(title="Z"),
+        ),
+        title="3D Mesh Visualization"
+    )
+
+    # Show the figure
+    fig.show()
+
+
+class bcolors:
+    HEADER = '\033[95m'
+    OKBLUE = '\033[94m'
+    OKCYAN = '\033[96m'
+    OKGREEN = '\033[92m'
+    WARNING = '\033[93m'
+    FAIL = '\033[91m'
+    ENDC = '\033[0m'
+    BOLD = '\033[1m'
+    UNDERLINE = '\033[4m'

### 2025-04-08 src/dataloader.py
Set up experiments

@@ -18,7 +18,6 @@ class DataLoader:
         self.edgemap_options = {mesh["name"]: {int(k): v for k,v in mesh["edgemap_options"].items()} for mesh in self.cfg["meshes"]}
         
         self.camera_matrices = load_camera_matrices(path=self.matrices_path, matrix_types="P")
-
         renders = load_renders(self.renders_path)
         self.edgemaps, self.edgemaps_len = load_edgemaps(renders, self.edgemap_options)
 
@@ -36,5 +35,5 @@ class DataLoader:
             mesh_path = os.path.join(self.mesh_dir, f"{mesh_name}_{self.mesh_res}.obj")
             verts, faces, aux = load_obj(mesh_path)
             # Create Meshes object in PyTorch3D
-            meshes[mesh_name] = Meshes(verts=[verts], faces=[faces])
+            meshes[mesh_name] = Meshes(verts=[verts], faces=[faces.verts_idx])
         return meshes

### 2025-04-08 src/experimentrunner.py

@@ -3,16 +3,24 @@ import json
 from .node import *
 from .chamfer import *
 from .utils import *
-from typing import List, Tuple
+from typing import List, Tuple, Union
 import random
+import wandb
+import os
 
 class ExperimentRunner:
-    def __init__(self, config_path: str, data_loader: DataLoader) -> None:
-    # Load the experiment config from a JSON file
-        with open(config_path, "r") as f:
-            self.cfg = json.load(f)
-        
+    def __init__(self, experiment_config: Union[str, dict], data_loader: DataLoader) -> None:
+        # If experiment_config is a path, load from file
+        if isinstance(experiment_config, str) and os.path.exists(experiment_config):
+            with open(experiment_config, "r") as f:
+                self.cfg = json.load(f)
+        elif isinstance(experiment_config, dict):
+            self.cfg = experiment_config
+        else:
+            raise ValueError("Invalid experiment_config. Must be a path or dict.")
+
         # Store the experiment details
+        self.project = self.cfg["project"]
         self.experiment_name = self.cfg["name"]
         self.experiment_description = self.cfg["description"]
         self.src_mesh = self.cfg["src_mesh"]
@@ -31,9 +39,12 @@ class ExperimentRunner:
         self.n_iters = self.cfg["training"]["n_iters"]
         self.lr = self.cfg["training"]["lr"]
         self.momentum = self.cfg["training"]["momentum"]
-        
+        self.verbose = self.cfg["verbose"]
+        self.vis_enabled = self.cfg["vis"]["enabled"]
+        self.vis_freq = self.cfg["vis"]["frequency"]
         # Use the DataLoader to load data
         self.data_loader = data_loader
+
 
     def run(self):
         return self.pipeline(self.src_mesh, self.target_meshes)
@@ -41,41 +52,67 @@ class ExperimentRunner:
     def pipeline(self, src_name: str, tgt_names: List[str]):
         """Pipeline function to run the experiment on a source mesh and a list of target meshes."""
         
+        run = wandb.init(
+            project=self.project,
+            name=self.experiment_name,
+            notes=self.experiment_description,
+            config={
+                "iters":    self.n_iters,
+                "lr":       self.lr,
+                "momentum": self.momentum,
+                "source": src_name,
[…]

### 2025-04-08 src/experiments/script.py

@@ -1,0 +1,13 @@
+from src.dataloader import *
+from src.experimentrunner import *
+import warnings
+warnings.filterwarnings('ignore')
+
+def main():
+    dataloader = DataLoader("/home/jeffrey/honours/src/experiments/skyconfig.json")
+    runner = ExperimentRunner("/home/jeffrey/honours/src/experiments/experiment_template.json", dataloader)
+    runner.run()
+
+
+if __name__ == "__main__":
+    main()

### 2025-04-08 src/node.py

@@ -1,9 +1,10 @@
 import torch
 from pytorch3d.structures import Meshes
-from ddn import EqConstDeclarativeNode, DeclarativeFunction
+from ddn.pytorch.node import *
 from .functions import *
 from .utils import *
 import scipy.optimize as opt
+import wandb
 
 class ConstrainedProjectionNode(EqConstDeclarativeNode):
     """
@@ -13,6 +14,7 @@ class ConstrainedProjectionNode(EqConstDeclarativeNode):
         super().__init__(eps=1.0e-6) # relax tolerance on optimality test 
         self.src = src # source meshes (B,)
         self.b = len(src)
+        self.iter = 0
 
     def objective(self, xs: torch.Tensor, y: torch.Tensor, scatter_add=False):
         """
@@ -73,7 +75,7 @@ class ConstrainedProjectionNode(EqConstDeclarativeNode):
         volumes = volumes.abs()
         return volumes  # Shape: (B,)    
     
-    def solve(self, xs: torch.Tensor):
+    def solve(self, xs: torch.Tensor, ):
         """Projects the vertices onto the target mesh vertices across batches.
 
         Args:
@@ -85,6 +87,7 @@ class ConstrainedProjectionNode(EqConstDeclarativeNode):
         n_batches = len(self.src)
         num_verts_per_mesh = self.src.num_verts_per_mesh()
         results = torch.zeros((n_batches, num_verts_per_mesh.max(), 3), dtype=torch.double)
+        losses = torch.zeros(n_batches, dtype=torch.double)
         for batch in range(n_batches):
             n_verts = num_verts_per_mesh[batch]
             verts = xs[batch][:n_verts].flatten().detach().double().cpu().numpy()
@@ -117,7 +120,15 @@ class ConstrainedProjectionNode(EqConstDeclarativeNode):
             if not res.success:
                 print("FAILED:", res.message)
             results[batch] = torch.tensor(res.x, dtype=torch.double, requires_grad=True).view(-1,3)
+            losses[batch] = res.fun
         results = torch.nn.utils.rnn.pad_sequence(results, batch_first=True)
+        
+        # assume one batch
+        wandb.log(
+            data={"inner/lstsq": losses[0]},
+            step=self.iter,
+        )
+
         return results,None
     
 

### 2025-04-08 src/utils.py

@@ -1,6 +1,6 @@
 import torch
 import matplotlib.pyplot as plt
-from chamfer import get_boundary
+from .chamfer import get_boundary
 import plotly.graph_objects as go
 import numpy as np
 

### 2025-04-08 docs/papers.tex
added nvidia and laplacian smoothing files

@@
 
-    \textbf{Summary:} Uses a neural network classifier to represent the continuous decision boundary of a 3D surface.
+    \textbf{Summary:} Uses a neural network classifier to represent the continuous decision boundary of a 3D surface, enables occupancy to be evaulated at arbitrary resolution.
 
@@
 For boundary voxels, we further subdivide them and repeat, until we reach the desired resolution, before applying Marching Cubes.
+        \item In ShapeNet results, able to construct finer details which get lost in coarser voxel representations.
     \end{itemize}
@@
     \begin{itemize}
-        \item This is unsupervised, so only requires images for training.
+        \item This is unsupervised, so only requires images for training, whereas existing approaches require large amounts of data.
         \item Existing pixel-wise loss functions render images from 3D structures but require rasterisation or pixel-wise interpolation commonly, which they claim don't actually contribute to structure learning
@@
 This ensures the entire foreground is covered and prevents clumping.
+        \item Link to my approach: renderers bridge the gap between 3D and 2D by using 2D loss functions on 3D structure, loss functions which are usually defined on RGB pixelvalues or pixel-wise silhouette coverage.
+Renders predicted 3D structure from specific view angle into an image.
     \end{itemize}

### 2025-04-08 src/experiments/script.py
wrote scripts, sphere to balloon varying number of views

@@ -1,5 +1,5 @@
-from src.dataloader import *
-from src.experimentrunner import *
+from src.framework_mesh.dataloader import *
+from src.framework_mesh.experimentrunner import *
 import warnings
 warnings.filterwarnings('ignore')
 

### 2025-04-08 src/experimentrunner.py
wrote scripts, sphere to balloon varying number of views

@@ -7,6 +7,7 @@ from typing import List, Tuple, Union
 import random
 import wandb
 import os
+from tqdm import trange
 
 class ExperimentRunner:
     def __init__(self, experiment_config: Union[str, dict], data_loader: DataLoader) -> None:
@@ -34,6 +35,8 @@ class ExperimentRunner:
                 "view_idx": view_config["view_idx"],
                 "num_views": view_config["num_views"]
             }
+            self.num_views = view_config["num_views"]
+
         
         # Training settings
         self.n_iters = self.cfg["training"]["n_iters"]
@@ -61,7 +64,8 @@ class ExperimentRunner:
                 "lr":       self.lr,
                 "momentum": self.momentum,
                 "source": src_name,
-                "targets": tgt_names[0] if len(tgt_names) == 1 else tgt_names
+                "targets": tgt_names[0] if len(tgt_names) == 1 else tgt_names,
+                "num_views": self.num_views
             }
         )
         wandb.define_metric("outer/chamfer", summary="min")
@@ -110,7 +114,9 @@ class ExperimentRunner:
         # plot_projections(verts.detach().squeeze().double(), projmats, (a,b))
         min_loss = float("inf")
         best_verts = None
-        for i in range(n_iters):
+    
+        pbar = trange(n_iters, desc="Training", leave=True)  # Always visible
+        for i in pbar:
             optimiser.zero_grad(set_to_none=True)
             node.iter = i + step_offset
             projverts = ConstrainedProjectionFunction.apply(node, verts)
@@ -123,6 +129,8 @@ class ExperimentRunner:
                 colour = bcolors.OKGREEN
             loss.backward()
             optimiser.step()
+
+            pbar.set_description(f"Loss: {loss.item():.4f}")
 
             # log
             wandb.log(

### 2025-04-08 src/experiments/script.py

@@ -4,10 +4,27 @@ import warnings
 warnings.filterwarnings('ignore')
 
 def main():
+    # from source S to target T, vary the number of views randomly
     dataloader = DataLoader("/home/jeffrey/honours/src/experiments/skyconfig.json")
-    runner = ExperimentRunner("/home/jeffrey/honours/src/experiments/experiment_template.json", dataloader)
-    runner.run()
 
+    with open("/home/jeffrey/honours/src/experiments/experiment_num_views.json") as f:
+        base_cfg = json.load(f)
+
+    tgt_mesh_name = base_cfg["target_meshes"][0]
+    total_views = len(base_cfg["views"][tgt_mesh_name]["view_idx"])
+
+    for nviews in range(1, total_views + 1):
+        cfg = base_cfg.copy()
+        cfg["views"][tgt_mesh_name]["num_views"] = nviews
+        base_name = f"{cfg['name']}{nviews}"
+        print(f"\n[INFO] Starting experiments with {nviews} view(s)...")
+
+        for r in range(5):
+            cfg["name"] = f"{base_name}_r{r}"
+            print(f"[INFO] Running: {cfg['name']}")
+            runner = ExperimentRunner(cfg, dataloader)
+            runner.run()
+            print(f"[INFO] Finished: {cfg['name']}")
 
 if __name__ == "__main__":
     main()

### 2025-04-08 src/experimentrunner.py
added edge configs and script

@@ -24,6 +24,7 @@ class ExperimentRunner:
         self.project = self.cfg["project"]
         self.experiment_name = self.cfg["name"]
         self.experiment_description = self.cfg["description"]
+        self.group = self.cfg["group"]
         self.src_mesh = self.cfg["src_mesh"]
         self.target_meshes = self.cfg["target_meshes"]
         
@@ -59,6 +60,7 @@ class ExperimentRunner:
             project=self.project,
             name=self.experiment_name,
             notes=self.experiment_description,
+            group=self.group,
             config={
                 "iters":    self.n_iters,
                 "lr":       self.lr,

### 2025-04-08 src/experiments/script.py

@@ -5,9 +5,9 @@ warnings.filterwarnings('ignore')
 
 def main():
     # from source S to target T, vary the number of views randomly
-    dataloader = DataLoader("/home/jeffrey/honours/src/experiments/skyconfig.json")
+    dataloader = DataLoader("/home/jeffrey/honours/src/experiments/skyconfig_edge1.json")
 
-    with open("/home/jeffrey/honours/src/experiments/experiment_num_views.json") as f:
+    with open("/home/jeffrey/honours/src/experiments/experiment_edges.json") as f:
         base_cfg = json.load(f)
 
     tgt_mesh_name = base_cfg["target_meshes"][0]
@@ -19,7 +19,7 @@ def main():
         base_name = f"{cfg['name']}{nviews}"
         print(f"\n[INFO] Starting experiments with {nviews} view(s)...")
 
-        for r in range(5):
+        for r in range(8):
             cfg["name"] = f"{base_name}_r{r}"
             print(f"[INFO] Running: {cfg['name']}")
             runner = ExperimentRunner(cfg, dataloader)

### 2025-04-09 src/experimentrunner.py
added group attribute to config

@@ -59,13 +59,14 @@ class ExperimentRunner:
             project=self.project,
             name=self.experiment_name,
             notes=self.experiment_description,
+            group="clean",
             config={
                 "iters":    self.n_iters,
                 "lr":       self.lr,
                 "momentum": self.momentum,
                 "source": src_name,
                 "targets": tgt_names[0] if len(tgt_names) == 1 else tgt_names,
-                "num_views": self.num_views
+                "num_views": self.num_views,
             }
         )
         wandb.define_metric("outer/chamfer", summary="min")
@@ -171,9 +172,8 @@ class ExperimentRunner:
             view_idx = self.views_config[mesh_name]["view_idx"]
         else: # random
             num_views = self.views_config[mesh_name]["num_views"]
-            view_idx = random.sample(list(camera_matrices.keys()), num_views)
-        print(view_idx)
-        print([k for k in camera_matrices.keys()])
+            view_idx = random.sample(self.views_config[mesh_name]["view_idx"], num_views)
+        print(f"{view_idx} from {self.views_config[mesh_name]['view_idx']}")
         projmats = torch.stack([camera_matrices[idx]["P"] for idx in view_idx])
         tgt_edgemaps = torch.nn.utils.rnn.pad_sequence([edgemaps[i] for i in view_idx], batch_first=True, padding_value=0.0)
         tgt_edgemaps_len = torch.tensor([edgemaps_len[i] for i in view_idx])

### 2025-04-09 src/experiments/script.py
added spline code

@@ -19,6 +19,10 @@ def main():
     runner = ExperimentRunner("/home/jeffrey/honours/src/experiments/exp_pipeline.json", dataloader)
     runner.run()
 
+    # from source S to target T, vary the number of views randomly
+    dataloader = DataLoader("/home/jeffrey/honours/src/experiments/skyconfig_balloon.json")
+    runner = ExperimentRunner("/home/jeffrey/honours/src/experiments/experiment_spline.json", dataloader)
+    runner.run()
 
 if __name__ == "__main__":
     main()

### 2025-04-09 src/framework_mesh/io.py

@@ -5,6 +5,8 @@ import re
 from collections import defaultdict
 import cv2
 from cv2.typing import MatLike
+from scipy.interpolate import splprep, splev
+
 
 def load_camera_matrices(path, matrix_types=None):
     """
@@ -109,9 +111,32 @@ def load_edgemaps(renders, options):
                 if num in edgemap_options:
                     edge_option = edgemap_options[num]
                     edges = canny_edge_map(img, edge_option)
-                    edge_coords = np.argwhere(edges > 0)
-                    edge_coords = edge_coords[:, [1, 0]]  # Convert to (x, y) coordinates
-                    
+                    if True:
+                        edge_coords = np.argwhere(edges > 0)
+                        edge_coords = edge_coords[:, [1, 0]]  # Convert to (x, y) coordinates
+                    else:
+                        contours, _ = cv2.findContours(edges, cv2.RETR_EXTERNAL, cv2.CHAIN_APPROX_NONE)
+                        all_subpixel_edges = []
+                        print("number of contours", len(contours))
+                        for contour in contours:
+                            if len(contour) < 5:
+                                continue
+                            contour = contour.squeeze()  # Remove single-dim (N,1,2) -> (N,2)
+                            x, y = contour[:, 0], contour[:, 1]
+
+                            # Fit a spline to the contour
+                            try:
+                                tck, u = splprep([x, y], s=1.0)  # s controls smoothing
+                                u_fine = np.linspace(0, 1, len(x)*5)  # More points = higher "resolution"
+                                x_fine, y_fine = splev(u_fine, tck)
+
+                                subpixel_points = np.vstack((x_fine, y_fine)).T
+                                all_subpixel_edges.append(subpixel_points)
+                                edge_coords = np.concatenate(all_subpixel_edges)
+                            except Exception as e:
+                                print(f"Skipping a contour due to error: {e}")
+                                continue
+
                     views[num] = torch.tensor(edge_coords)
                     views_len[num] = len(edge_coords)
 
@@ -120,3 +145,5 @@ def load_edgemaps(renders, options):
         edgemaps_len[mesh_name] = views_len
 
     return edgemaps, edgemaps_len
+
+

### 2025-04-09 src/dataloader.py
added spline code

@@ -15,11 +15,10 @@ class DataLoader:
         self.renders_path = self.cfg["paths"]["renders_path"]
         self.matrices_path = self.cfg["paths"]["matrices_path"]
         
-        self.edgemap_options = {mesh["name"]: {int(k): v for k,v in mesh["edgemap_options"].items()} for mesh in self.cfg["meshes"]}
+        self.edgemap_options = {mesh["name"]: {int(k): v for k,v in mesh.get("edgemap_options", {}).items()} for mesh in self.cfg["meshes"]}
         
         self.camera_matrices = load_camera_matrices(path=self.matrices_path, matrix_types="P")
-        renders = load_renders(self.renders_path)
-        self.edgemaps, self.edgemaps_len = load_edgemaps(renders, self.edgemap_options)
+        self.renders = load_renders(self.renders_path)
 
         # Load meshes using PyTorch3D
         self.meshes = self.load_meshes()

### 2025-04-09 src/experimentrunner.py

@@ -3,6 +3,7 @@ import json
 from .node import *
 from .chamfer import *
 from .utils import *
+from .io import *
 from typing import List, Tuple, Union
 import random
 import wandb
@@ -47,6 +48,8 @@ class ExperimentRunner:
         self.vis_freq = self.cfg["vis"]["frequency"]
         # Use the DataLoader to load data
         self.data_loader = data_loader
+        edgemap_options = {k: v for k,v in data_loader.edgemap_options.items() if k in self.target_meshes}
+        self.edgemaps, self.edgemaps_len = load_edgemaps(data_loader.renders, edgemap_options)
 
 
     def run(self):
@@ -158,8 +161,7 @@ class ExperimentRunner:
         return self.data_loader.meshes[name]
     
     def get_edgemaps(self, name):
-        dataloader = self.data_loader
-        return dataloader.edgemaps[name], dataloader.edgemaps_len[name]
+        return self.edgemaps[name], self.edgemaps_len[name]
 
     def get_camera_matrices(self):
         return self.data_loader.camera_matrices

### 2025-04-09 src/experiments/script.py

@@ -5,26 +5,9 @@ warnings.filterwarnings('ignore')
 
 def main():
     # from source S to target T, vary the number of views randomly
-    dataloader = DataLoader("/home/jeffrey/honours/src/experiments/skyconfig.json")
-
-    with open("/home/jeffrey/honours/src/experiments/experiment_num_views.json") as f:
-        base_cfg = json.load(f)
-
-    tgt_mesh_name = base_cfg["target_meshes"][0]
-    total_views = len(base_cfg["views"][tgt_mesh_name]["view_idx"])
-
-    for nviews in range(1, total_views + 1):
-        cfg = base_cfg.copy()
-        cfg["views"][tgt_mesh_name]["num_views"] = nviews
-        base_name = f"{cfg['name']}{nviews}"
-        print(f"\n[INFO] Starting experiments with {nviews} view(s)...")
-
-        for r in range(5):
-            cfg["name"] = f"{base_name}_r{r}"
-            print(f"[INFO] Running: {cfg['name']}")
-            runner = ExperimentRunner(cfg, dataloader)
-            runner.run()
-            print(f"[INFO] Finished: {cfg['name']}")
+    dataloader = DataLoader("/home/jeffrey/honours/src/experiments/skyconfig_balloon.json")
+    runner = ExperimentRunner("/home/jeffrey/honours/src/experiments/experiment_spline.json", dataloader)
+    runner.run()
 
 if __name__ == "__main__":
     main()

### 2025-05-03 src/framework_ellipsoid/plotting.py
pushing so i don't forget next time

@@ -6,6 +6,7 @@ import plotly.graph_objects as go
 from matplotlib.patches import Polygon
 from framework_ellipsoid.utils import rotation_matrix_3d
 import math
+import numpy as np
 
 def plot_ellipsoid_mpl(a, b, c, yaw, pitch, roll, points, r=0.62, vmin=None, vmax=None, show_gt=False):
     u = torch.linspace(0, 2 * torch.pi, 80)
@@ -26,7 +27,8 @@ def plot_ellipsoid_mpl(a, b, c, yaw, pitch, roll, points, r=0.62, vmin=None, vma
 
     # Normalize residuals for coloring
     if vmin is None or vmax is None:
-        vmin, vmax = residuals.min().item(), residuals.max().item()
+        absmax = torch.abs(residuals).item()
+        vmin, vmax = -absmax, absmax
     norm = colors.Normalize(vmin=vmin, vmax=vmax)
     cmap = plt.colormaps["RdBu_r"]
     facecolors = cmap(norm(residuals))
@@ -92,6 +94,7 @@ def plot_ellipsoid_plotly(a, b, c, yaw, pitch, roll, points, r=0.62, vmin=None, 
     radii = torch.norm(ellipsoid_xyz, dim=0)
     residuals = (radii - r).reshape(rotated.shape[:2]).cpu().numpy()
     if vmin is None or vmax is None:
+        
         vmin, vmax = residuals.min().item(), residuals.max().item()
 
     surface = go.Surface(

### 2025-05-03 src/framework_mesh/gradient.py

@@ -36,12 +36,29 @@ def make_jacobi_hook(edge_src, edge_dst, boundary_mask, k=5, constrained=False, 
         g = grad_in.view(-1, 3).clone()
         for _ in range(k):
             g_nb = torch.zeros_like(g)
-            g_nb.scatter_add_(0, edge_dst[:, None].expand(-1, 3), g[edge_src])
-            deg = torch.zeros(V, 1, device=g.device)
-            deg.scatter_add_(0, edge_dst[:, None], onesE)
+
+            if constrained:
+                # Mask: only use boundary vertices for smoothing
+                mask_src = boundary_mask[edge_src]  # shape (E,)
+                g_src_masked = g[edge_src] * mask_src[:, None].float()
+                g_nb.scatter_add_(0, edge_dst[:, None].expand(-1, 3), g_src_masked)
+
+                # Degree should count only boundary neighbors
+                deg_masked = mask_src.float()
+                deg = torch.zeros(V, 1, device=g.device)
+                deg.scatter_add_(0, edge_dst[:, None], deg_masked[:, None])
+            else:
+                # Use all neighbors
+                g_nb.scatter_add_(0, edge_dst[:, None].expand(-1, 3), g[edge_src])
+                deg = torch.zeros(V, 1, device=g.device)
+                deg.scatter_add_(0, edge_dst[:, None], onesE)
+
+            # Add self-contribution
             g_nb += g
             deg += 1.0
-            g_nb = g_nb / (deg.clamp_min_(1.0)+1)
+            g_nb = g_nb / (deg.clamp_min_(1.0) + 1)
+
+            # Update gradients
             if constrained:
                 g = torch.where(boundary_mask[:, None], g, g_nb)
             else:

### 2025-05-03 src/framework_mesh/script.py

@@ -36,7 +36,7 @@ def main():
                 exp_config['gradient']['constrained'] = constrained
 
                 # Update experiment name to be descriptive
-                exp_config['name'] = f"pipeline_{method}_constrained_{constrained}_res_{mesh_res}"
+                exp_config['name'] = f"jaconlyboundary_pipeline_{method}_constrained_{constrained}_res_{mesh_res}"
 
                 # Create runner and run
                 runner = ExperimentRunner(exp_config, dataloader)

### 2025-05-03 src/framework_mesh/utils.py

@@ -403,7 +403,7 @@ def compute_boundary_distance_heatmap(src: Meshes, boundary_mask, D_all, cmin=No
         heatmap_values = min_dist.float().cpu().numpy()
 
     if cmin is None or cmax is None:
-        max_val = np.percentile(heatmap_values, 95)
+        max_val = max(1,np.percentile(heatmap_values, 95))
         cmin, cmax = 0, max_val
 
     mesh = trimesh.Trimesh(vertices=verts3d.cpu().numpy(), faces=faces.cpu().numpy())

### 2025-05-03 src/framework_ellipsoid/experiments.py
pushing so i don't forget next time

@@ -147,8 +147,7 @@ class EllipsoidExperiment:
         if backend == "mpl":
             fig,vmin,vmax = plot_ellipsoid_mpl(a, b, c, yaw, pitch, roll, points.cpu(), r=r, vmin=self.vmin, vmax=self.vmax)
             wandb.log({f"vis/ellipsoid": wandb.Image(fig)}, step=step)
-            plt.close(fig)
-
+            plt.close()
         elif backend == "plotly":
             fig,vmin,vmax = plot_ellipsoid_plotly(a, b, c, yaw, pitch, roll, points.cpu(), r=r, vmin=self.vmin, vmax=self.vmax)
             wandb.log({f"vis/ellipsoid": wandb.Plotly(fig)}, step=step)
@@ -162,4 +161,4 @@ class EllipsoidExperiment:
             if callable(silhouette):
                 fig = silhouette(step=step)
                 wandb.log({f"vis/silhouettes": wandb.Image(fig)}, step=step)
-                plt.close(fig)
+                plt.close()

### 2025-05-03 src/framework_ellipsoid/loss.py

@@ -154,16 +154,46 @@ class BoundaryProjectionChamferLoss(nn.Module):
         z = torch.cos(theta)
         return torch.stack([x.flatten(), y.flatten(), z.flatten()], dim=0)
 
+    def sample_ellipsoid_surface_uniform(self, A: torch.Tensor, n: int, oversample: float = 2.0):
+        """
+        Sample n points uniformly from the surface {x | xᵀ A x = 1}.
+        """
+        eigvals, eigvecs = torch.linalg.eigh(A)
+        A_inv_sqrt = eigvecs @ torch.diag(1.0 / torch.sqrt(eigvals)) @ eigvecs.T
+
+        samples = []
+        batch = int(n * oversample)
+
+        while sum(p.shape[1] for p in samples) < n:
+            u = torch.randn(3, batch, dtype=torch.double)
+            u /= torch.norm(u, dim=0, keepdim=True)
+            
+            x = A_inv_sqrt @ u
+            # norm_A = torch.sqrt(torch.sum(x_tilde * (A @ x_tilde), dim=0))  # correct A-norm
+            # x = x_tilde / norm_A.unsqueeze(0)
+
+            # weights = torch.norm(x_tilde, dim=0)  # ||M u|| (used for density correction)
+            # probs = 1.0 / weights**3
+            # probs /= probs.max()
+
+            accept = torch.rand(batch) < 1
+            samples.append(x[:, accept])
+
+        result = torch.cat(samples, dim=1)[:, :n]
+        return result  # (3, n)
+
+
     def forward(self, input, p=1.6075):
         B = len(self.rot_mats)
         A = A_from_u_batch(input).double()  # (1, 3, 3)
         Au = A[0]
-        L = torch.linalg.cholesky(Au)  # B in pseudocode: upper-triangular
-        Binv = torch.inverse(L)        # B⁻¹ (3×3)
+        # L = torch.linalg.cholesky(Au)  # B in pseudocode: upper-triangular
+        # Binv = torch.inverse(L)        # B⁻¹ (3×3)
 
-        z = self.sample_unit_sphere(self.sqrt_m, device=input.device)  # (3, M)
-        E = Binv @ z  # (3, M), ellipsoid surface points
-
+        # z = self.sample_unit_sphere(self.sqrt_m, device=input.device)  # (3, M)
+        # E = Binv @ z  # (3, M), ellipsoid surface points
+        n = self.sqrt_m ** 2
[…]

### 2025-05-03 src/framework_ellipsoid/plotting.py

@@ -1,4 +1,6 @@
 import torch
+import matplotlib
+matplotlib.use('Agg') 
 import matplotlib.pyplot as plt
 import matplotlib.cm as cm
 import matplotlib.colors as colors

### 2025-05-03 src/framework_ellipsoid/realscript.py

@@ -1,0 +1,145 @@
+import json
+import argparse
+from copy import deepcopy
+from .experiments import EllipsoidExperiment
+import warnings
+warnings.filterwarnings('ignore')
+from pathlib import Path
+import matplotlib
+matplotlib.use('Agg') 
+
+# ----------------------------- PRESETS ------------------------------------ #
+AXES_PRESETS_02 = {
+    "A1": [0.271, 0.271, 0.271],            # sphere
+    "A2": [0.215, 0.215, 0.431],            # oblate spheroid
+    "A3": [0.171, 0.342, 0.342],            # prolate spheroid
+    "A4": [0.22314, 0.26777, 0.33472],            # triaxial 1:1.2:1.5
+    "A5": [0.18821, 0.28231, 0.37641],            # triaxial 1:1.5:2
+    "A6": [0.14938, 0.29876, 0.44814]           # triaxial 1:2:3
+}
+
+AXES_PRESETS_24 = {
+    "A1": [0.620, 0.620, 0.620],
+    "A2": [0.493, 0.493, 0.986],
+    "A3": [0.391, 0.783, 0.783],
+    "A4": [0.51087, 0.61305, 0.76631],
+    "A5": [0.43089, 0.64633, 0.86177],
+    "A6": [0.342, 0.68399, 1.02599]
+}
+
+ANGLE_PRESETS = {
+    "R1": [0, 0, 0],
+    "R2": [30, 30, 0],
+    "R3": [45, 45, 45],
+    "R4": [45, 60, 80],
+    "R5": [80, 45, 15],
+}
+
+PROBLEM_SETTINGS = [
+    ("axisaligned", 0.02),
+    ("nonaxisaligned", 0.02),
+    ("chamfersampled", 0.24),
+    ("chamferboundary", 0.24),
+    ("volumeconstrained", 0.24),
+]
+
+# Base configuration shared by all runs
+BASE_CFG = {
+    "project": "grid_init_uniform_sampling",
+    "name": "auto-run",
+    "sqrt_m": 30,
+    "seed": 42,
+    "noise": 1e-2,
+    "p": 1.6075,
+    "initial_axes": [0.2, 0.2, 0.2],  # placeholder, will be overwritten
+    "initial_angles": [0, 0, 0],       # placeholder, will be overwritten
+    "target": {"radius": 0.2820947918, "m": 300, "sqrt_m": 30},
+    "view_mode": "angles",
[…]

### 2025-05-03 src/framework_ellipsoid/utils.py

@@ -12,10 +12,10 @@ def sample_ellipsoid_surface(sqrt_m, a, b, c, yaw, pitch, roll, noise_std=1e-4, 
             x = a * torch.sin(t) * torch.cos(u)
             y = b * torch.sin(t) * torch.sin(u)
             z = c * torch.cos(t)
-            return torch.stack([x, y, z], dim=0)  # shape (3, ...)
-        t_min, t_max = 0, torch.pi
+            return torch.stack([x, y, z], dim=0)  # (3, ...)
         u_min, u_max = 0, 2 * torch.pi
-        coords = r_surface(m, ellipsoid_func, t_min, t_max, u_min, u_max, 100, 100)
+        t_min, t_max = 0, torch.pi
+        coords = r_surface(m, ellipsoid_func, t_min, t_max, u_min, u_max, m, m).double()
 
     else:
         phi = 2.0 * math.pi * torch.linspace(0.0, 1.0, sqrt_m).double()
@@ -158,22 +158,20 @@ def build_view_matrices(cfg):
     return R
 
 def r_surface(n, func, t0, t1, u0, u1, t_precision=50, u_precision=50, device='cpu'):
-    # Create t, u grids
-    t_vals = 2.0 * torch.pi * torch.linspace(0,1, t_precision, device=device).double()
-    u_vals = torch.pi * torch.linspace(0,1, u_precision, device=device).double()
-    t, u = torch.meshgrid(t_vals, u_vals, indexing='ij')  # (t_precision, u_precision)
+    theta_vals = torch.linspace(t0, t1, t_precision, device=device).double()
+    phi_vals = torch.linspace(u0, u1, u_precision, device=device).double()
+    theta, phi = torch.meshgrid(theta_vals, phi_vals, indexing='ij')  # (t, u)
 
-    coords = func(t, u)  # (3, t_precision, u_precision)
+    coords = func(theta, phi)  # (3, t, u)
 
-    # Compute differential vectors
+    # Area elements
     dt = torch.zeros_like(coords)
     du = torch.zeros_like(coords)
     dt[:, 1:, :] = coords[:, 1:, :] - coords[:, :-1, :]
     du[:, :, 1:] = coords[:, :, 1:] - coords[:, :, :-1]
 
[…]

### 2025-05-04 src/framework_ellipsoid/experiments.py
committing after the shut down

@@ -36,7 +36,13 @@ class EllipsoidExperiment:
             run = wandb.init(
                 project=self.cfg.get("project", "ellipsoid-exp"),
                 name=self.cfg.get("name", f"{self.cfg['problem']}-{self.cfg.get('target', {}).get('m', 'unknown')}"),
-                config=self.cfg,    
+                config={
+                    "problem": self.cfg["problem"],
+                    # "axes_id": self.cfg["axes_id"],         # letter like "A"
+                    # "rotation_id": self.cfg["rotation_id"], # letter like "R2"
+                    # "trial": self.cfg["trial"],
+                    **self.cfg
+                },
                 reinit=True,
             )
             wandb.define_metric("outer/loss", summary="min")
@@ -55,7 +61,6 @@ class EllipsoidExperiment:
             res = loss_fn(y)
             outer_loss = res.sum()
             outer_loss.backward()
-            optimiser.step()
             pbar.set_description(f"Loss: {outer_loss.item():.4f}")
             if self.wandb:
                 wandb.log({
@@ -72,6 +77,7 @@ class EllipsoidExperiment:
             c_hat = y_detached[2].item()
             if y_detached.numel() >= 6:  # check if angles exist
                     yaw_hat, pitch_hat, roll_hat = torch.rad2deg(y_detached[3:6]).tolist()
+
 
             if self.wandb:
                 self.log_geometry_and_errors(y_detached, self.cfg, i)
@@ -90,6 +96,9 @@ class EllipsoidExperiment:
                         f"has volume {ellipsoid_volume(a_hat, b_hat, c_hat):.3f} and "
                         f"surface area {ellipsoid_surface_area(a_hat, b_hat, c_hat, self.problem.p):.5f}. "
                         f"LR {optimiser.param_groups[0]['lr']}")
+                    
+            optimiser.step()
+
         if self.wandb:
             run.finish()
 
@@ -140,14 +149,13 @@ class EllipsoidExperiment:
             yaw = pitch = roll = 0.0
 
         r = cfg["target"]["radius"]
-        backend = vis_cfg.get("backend", "mpl")
+        backend = vis_cfg.get("backend", "plotly")
 
         points = x[0].detach().view(3,-1)
 
         if backend == "mpl":
[…]

### 2025-05-04 src/framework_ellipsoid/loss.py

@@ -154,16 +154,46 @@ class BoundaryProjectionChamferLoss(nn.Module):
         z = torch.cos(theta)
         return torch.stack([x.flatten(), y.flatten(), z.flatten()], dim=0)
 
+    def sample_ellipsoid_surface_uniform(self, A: torch.Tensor, n: int, oversample: float = 2.0):
+        # eigvals, eigvecs = torch.linalg.eigh(A)
+        # A_inv_sqrt = eigvecs @ torch.diag(1.0 / torch.sqrt(eigvals)) @ eigvecs.T
+        L = torch.linalg.cholesky(A)  # B in pseudocode: upper-triangular
+        A_inv_sqrt = torch.linalg.inv(L)
+
+        accepted = []
+        batch = int(n * oversample)
+
+        while sum(p.shape[1] for p in accepted) < n:
+            u = torch.randn(3, batch, dtype=torch.float64)
+            u /= torch.norm(u, dim=0, keepdim=True)
+
+            x = A_inv_sqrt @ u
+            accepted.append(x)
+            norm_A = torch.sqrt(torch.sum(x * (A @ x), dim=0))  # shape: (N,)
+            x = x / norm_A.unsqueeze(0)
+
+            weights = torch.norm(A_inv_sqrt @ u, dim=0)
+            probs = 1.0 / weights**3
+            probs /= probs.max()
+
+            accept = torch.rand(batch) < probs
+            accepted.append(x[:, accept])
+
+        result = torch.cat(accepted, dim=1)[:, :n]
+        return result
+        
+
     def forward(self, input, p=1.6075):
         B = len(self.rot_mats)
         A = A_from_u_batch(input).double()  # (1, 3, 3)
         Au = A[0]
         L = torch.linalg.cholesky(Au)  # B in pseudocode: upper-triangular
-        Binv = torch.inverse(L)        # B⁻¹ (3×3)
+        Binv = torch.linalg.inv(L)        # B⁻¹ (3×3)
 
-        z = self.sample_unit_sphere(self.sqrt_m, device=input.device)  # (3, M)
-        E = Binv @ z  # (3, M), ellipsoid surface points
-
+        # z = self.sample_unit_sphere(self.sqrt_m, device=input.device)  # (3, M)
+        # E = Binv @ z  # (3, M), ellipsoid surface points
+        n = self.sqrt_m * self.sqrt_m
+        E = self.sample_ellipsoid_surface_uniform(Au, n)
[…]

### 2025-05-04 src/framework_ellipsoid/plotting.py

@@ -6,6 +6,7 @@ import plotly.graph_objects as go
 from matplotlib.patches import Polygon
 from framework_ellipsoid.utils import rotation_matrix_3d
 import math
+import numpy as np
 
 def plot_ellipsoid_mpl(a, b, c, yaw, pitch, roll, points, r=0.62, vmin=None, vmax=None, show_gt=False):
     u = torch.linspace(0, 2 * torch.pi, 80)
@@ -92,7 +93,8 @@ def plot_ellipsoid_plotly(a, b, c, yaw, pitch, roll, points, r=0.62, vmin=None, 
     radii = torch.norm(ellipsoid_xyz, dim=0)
     residuals = (radii - r).reshape(rotated.shape[:2]).cpu().numpy()
     if vmin is None or vmax is None:
-        vmin, vmax = residuals.min().item(), residuals.max().item()
+        absmax = np.abs(residuals).max().item()
+        vmin, vmax = -absmax, absmax
 
     surface = go.Surface(
         x=rotated[..., 0].cpu().numpy(),
@@ -100,9 +102,10 @@ def plot_ellipsoid_plotly(a, b, c, yaw, pitch, roll, points, r=0.62, vmin=None, 
         z=rotated[..., 2].cpu().numpy(),
         surfacecolor=residuals,
         colorscale='RdBu',
+        reversescale=True,
         cmin=vmin,
         cmax=vmax,
-        opacity=0.95,
+        opacity=0.98,
         showscale=True
     )
 
@@ -111,7 +114,7 @@ def plot_ellipsoid_plotly(a, b, c, yaw, pitch, roll, points, r=0.62, vmin=None, 
     scatter = go.Scatter3d(
         x=pts[0], y=pts[1], z=pts[2],
         mode='markers',
-        marker=dict(size=1, color='black', opacity=0.5)
+        marker=dict(size=1.2, color='black', opacity=0.5)
     )
 
     # Optional: GT sphere wireframe
@@ -133,11 +136,11 @@ def plot_ellipsoid_plotly(a, b, c, yaw, pitch, roll, points, r=0.62, vmin=None, 
 
     fig = go.Figure(data=data)
     fig.update_layout(scene=dict(
-        xaxis=dict(range=[-1, 1]),
-        yaxis=dict(range=[-1, 1]),
-        zaxis=dict(range=[-1, 1]),
-        aspectmode='cube'
+        aspectmode='data'
     ))
+    fig.update_layout(
[…]

### 2025-05-04 src/framework_ellipsoid/utils.py

@@ -5,6 +5,7 @@ from torch import cos, sin
 import torch.nn.functional as F
 from scipy.interpolate import interp1d
 
+# Main sampling function
 def sample_ellipsoid_surface(sqrt_m, a, b, c, yaw, pitch, roll, noise_std=1e-4, uniform=True):
     if uniform:
         m = sqrt_m * sqrt_m
@@ -12,20 +13,19 @@ def sample_ellipsoid_surface(sqrt_m, a, b, c, yaw, pitch, roll, noise_std=1e-4, 
             x = a * torch.sin(t) * torch.cos(u)
             y = b * torch.sin(t) * torch.sin(u)
             z = c * torch.cos(t)
-            return torch.stack([x, y, z], dim=0)  # shape (3, ...)
-        t_min, t_max = 0, torch.pi
-        u_min, u_max = 0, 2 * torch.pi
-        coords = r_surface(m, ellipsoid_func, t_min, t_max, u_min, u_max, 100, 100)
-
+            return torch.stack([x, y, z], dim=0)
+        coords = r_surface(m, ellipsoid_func, 0, torch.pi, 0, 2*torch.pi, 100, 100)
     else:
-        phi = 2.0 * math.pi * torch.linspace(0.0, 1.0, sqrt_m).double()
-        theta = math.pi * torch.linspace(0.0, 1.0, sqrt_m).double()
-        phi, theta = torch.meshgrid(phi, theta, indexing='ij')
-        x = a * torch.sin(theta) * torch.cos(phi)
-        y = b * torch.sin(theta) * torch.sin(phi)
-        z = c * torch.cos(theta)
-        coords = torch.stack((x.flatten(), y.flatten(), z.flatten()), dim=0)
-    
+        def ellipsoid_func(theta, phi):
+            x = a * torch.sin(theta) * torch.cos(phi)
+            y = b * torch.sin(theta) * torch.sin(phi)
+            z = c * torch.cos(theta)
+            return torch.stack([x, y, z], dim=0)
+
+        theta = torch.linspace(0, math.pi, sqrt_m).double()
+        phi = torch.linspace(0, 2 * math.pi, sqrt_m).double()
+        theta, phi = torch.meshgrid(theta, phi, indexing='ij')
+        coords = ellipsoid_func(theta, phi).reshape(3, -1)
[…]

### 2025-05-05 src/framework_mesh/dataloader.py
did edgemaps for diffuse

@@ -19,33 +19,71 @@ class DataLoader:
         self.mesh_dir = self.cfg["paths"]["mesh_dir"]
         self.mesh_res = self.cfg["paths"]["mesh_res"]
         self.renders_path = self.cfg["paths"]["renders_path"]
+        self.material = self.cfg["paths"]["material"]
+        self.materials_path = os.path.join(self.renders_path, self.material)
         self.matrices_path = self.cfg["paths"]["matrices_path"]
         
-        self.edgemap_options = {mesh["name"]: {int(k): v for k,v in mesh.get("edgemap_options", {}).items()} for mesh in self.cfg["meshes"]}
-        
-        self.camera_matrices = load_camera_matrices(path=self.matrices_path, matrix_types="P")
-        self.renders = load_renders(self.renders_path)
+        self.edgemap_options = {
+            mesh["name"]: mesh.get("edgemap_options", {})
+            for mesh in self.cfg["meshes"]
+        }
+        # self.camera_matrices = load_camera_matrices(path=self.matrices_path, matrix_types="P")
+        # self.renders = load_renders(self.renders_path)
 
         # Load meshes using PyTorch3D
-        self.meshes, self.gt_meshes = self.load_meshes()
+        # self.meshes, self.gt_meshes = self.load_meshes()
 
-    def load_meshes(self):
+    def load_renders(self, objname):
+        return load_renders(self.materials_path, objname)
+    
+    def load_camera_matrices(self):
+        return load_camera_matrices(self.matrices_path, matrix_types="P")
+        # return cam_id_map  # {int_id: {P: ...}}
+
+
+
+    # def load_meshes(self):
+    #     """
+    #     Load meshes using PyTorch3D's load_obj function.
+    #     Returns a dictionary of PyTorch3D Meshes objects.
+    #     """
+    #     meshes = {}
+    #     gt_meshes = {}
+    #     for mesh_info in self.cfg["meshes"]:
+    #         mesh_name = mesh_info["name"]
+    #         if mesh_name == "sphere":
+    #             mesh_path = os.path.join(self.mesh_dir, f"{mesh_name}_{self.mesh_res}.obj")
+    #             verts, faces, aux = load_obj(mesh_path)
+    #             # Create Meshes object in PyTorch3D
+    #             meshes[mesh_name] = Meshes(verts=[verts], faces=[faces.verts_idx])
+
[…]

### 2025-05-05 src/framework_mesh/experimentrunner.py

@@ -33,7 +33,7 @@ class ExperimentRunner:
         for mesh_name, view_config in self.cfg["views"].items():
             self.views_config[mesh_name] = {
                 "mode": view_config["mode"],
-                "view_idx": view_config["view_idx"],
+                "view_names": view_config["view_names"],
                 "num_views": view_config["num_views"]
             }
             self.num_views = view_config["num_views"]
@@ -48,8 +48,8 @@ class ExperimentRunner:
         self.vis_freq = self.cfg["vis"]["frequency"]
         # Use the DataLoader to load data
         self.data_loader = data_loader
-        edgemap_options = {k: v for k,v in data_loader.edgemap_options.items() if k in self.target_meshes}
-        self.edgemaps, self.edgemaps_len = load_edgemaps(data_loader.renders, edgemap_options)
+        # edgemap_options = {k: v for k,v in data_loader.edgemap_options.items() if k in self.target_meshes}
+        # self.edgemaps, self.edgemaps_len = load_edgemaps(data_loader.renders, edgemap_options)
         
         self.wandb = self.cfg["wandb"]
 
@@ -77,34 +77,45 @@ class ExperimentRunner:
             wandb.define_metric("outer/gt/chamfer", summary="min")
             wandb.define_metric("outer/gt/iou", summary="max")
 
-        view_idxs = {}
+        view_names_used = {}
         step_offset = 0
 
         src_mesh = self.get_mesh(src_name)
+
         for tgt_name in tgt_names:
             tgt_mesh = self.get_gt_mesh(tgt_name)
 
-            # Select the view points and get corresponding projmats
-            projmats, tgt_edgemap_info, view_idx = self.get_projmats_and_edgemap_info(tgt_name)
+            # Use encapsulated loading logic
+            projmats, tgt_edgemap_info, view_ids = self.get_projmats_and_edgemap_info(tgt_name)
             gt_projmats, gt_edgemap_info, _ = self.get_gt_projmats_and_edgemap_info(tgt_name)
 
-            view_idxs[tgt_name] = view_idx
+            # Reverse map view_ids to view_names
+            _, _, cam_id_to_name = self.data_loader.load_camera_matrices()
+            view_names = [cam_id_to_name[i] for i in view_ids]
+            view_names_used[tgt_name] = view_names
+
             edgemap_info = ([tgt_edgemap_info[0]], [tgt_edgemap_info[1]])
 
[…]

### 2025-05-05 src/framework_mesh/io.py

@@ -9,12 +9,9 @@ from scipy.interpolate import splprep, splev
 
 
 def load_camera_matrices(path, matrix_types=None):
-    """
-    Loads camera matrices from .npy files in the specified directory.
-    Returns a nested dict: cameras[obj_name][cam_name][K/RT/P]
-    """
-    cameras = defaultdict(lambda: defaultdict(dict))
-    file_pattern = re.compile(r"^(Cam_([A-Za-z]+_\d+)_(K|RT|P)\.npy$")
+    cameras = defaultdict(dict)
+    cam_names = set()
+    file_pattern = re.compile(r"^Cam_([A-Za-z]+_\d+)_(K|RT|P)\.npy$")
 
     if matrix_types is not None:
         matrix_types = set(matrix_types)
@@ -26,50 +23,29 @@ def load_camera_matrices(path, matrix_types=None):
             if matrix_types is None or matrix_type in matrix_types:
                 filepath = os.path.join(path, filename)
                 cameras[cam_name][matrix_type] = torch.tensor(np.load(filepath))
+                cam_names.add(cam_name)
 
-    return cameras
+    cam_names = sorted(cam_names)
+    cam_name_to_id = {name: idx for idx, name in enumerate(cam_names)}
+    cam_id_to_name = {idx: name for name, idx in cam_name_to_id.items()}
+    cameras_id = {cam_name_to_id[name]: cameras[name] for name in cam_names}
 
-def load_renders(renders_path):
-    renders = defaultdict(lambda: defaultdict(dict))
-    pattern = re.compile(r"^(Oblique_[^_]+)_Cam_([A-Za-z]+_\d+)\.png$")
+    return cameras_id, cam_name_to_id, cam_id_to_name
+
+def load_renders(renders_path, object_name):
+    renders = {}
+    pattern = re.compile(rf"^{object_name}_Cam_(Overhead|Above|Ground)_(\d+)\.png")
 
     for filename in sorted(os.listdir(renders_path)):
         match = pattern.match(filename)
         if match:
-            obj_name, cam_name = match.groups()
+            view_type, view_number = match.groups()
+            cam_name = f"{view_type}_{view_number}"
             image_path = os.path.join(renders_path, filename)
             image = cv2.imread(image_path)
             if image is not None:
-                renders[obj_name][cam_name] = image
-
-    return renders
-
-
-def get_projmats_and_edgemap_info(view_idx, target_mesh: str, matrices, edgemaps, edgemaps_len):
-    """
[…]

### 2025-05-05 src/framework_mesh/script.py

@@ -4,14 +4,14 @@ import warnings
 warnings.filterwarnings('ignore')
 
 def main():
-    data_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/skyconfig_local.json"
+    data_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/obliqueconfig_local.json"
     with open(data_path, 'r') as f:
         config = json.load(f)
 
     # Modify the 'mesh_res' value
     config['paths']['mesh_res'] = 2
     dataloader = DataLoader(config)
-    runner = ExperimentRunner("/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/exp_pipeline.json", dataloader)
+    runner = ExperimentRunner("/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/exp_oblique.json", dataloader)
     runner.run()
 
     # config['paths']['mesh_res'] = 3

### 2025-05-05 src/framework_ellipsoid/base_problem.py
Feat/gradient (#3)
* Pulling

* created experiment setup for ellipsoids:

* added renders and experiment files

* recalculated normals for lstrawberry meshes

* wrote pipeline script

* created experiment setup for ellipsoids:

* continuing cherrypick

* added init and gpu config

* Changed to Plotly visualisation

* removed oblique data

* refactored inner problem nodes

* Added mesh projection outline detection

* testing for longer

* wrote pipeline script

* added plotting

* fixed up coding

* added cuda device

* Revert "added cuda device"

This reverts commit e8cc346dd0c059f7138bd893014

@@ -1,8 +1,20 @@
+<<<<<<< HEAD
+=======
+import torch
+>>>>>>> main
 import abc
 
 class BaseEllipsoidProblem(abc.ABC):
     def __init__(self, cfg):
         self.cfg = cfg
+        self.sqrt_m = cfg["sqrt_m"]
+        self.seed = cfg["seed"]
+        torch.seed = self.seed
+        self.nu = cfg.get("noise", 1e-4)
+        self.p = cfg.get("p", 1.6075)
+        self.initial_axes = cfg["initial_axes"]  # [a, b, c]
+        self.m = self.sqrt_m * self.sqrt_m
+        self.wandb = cfg.get("wandb", False)
 
     @abc.abstractmethod
     def generate_data(self):

### 2025-05-05 src/framework_ellipsoid/experiments.py

@@ -3,60 +3,163 @@ import torch
 import wandb
 from .utils import ellipsoid_volume, ellipsoid_surface_area
 from framework_ellipsoid.base_problem import BaseEllipsoidProblem
+from .plotting import *
+from .problems import ChamferBoundaryProblem, ChamferSampledProblem
+from tqdm import trange
+import math
+import io
+
 
 class EllipsoidExperiment:
     def __init__(self, cfg):
         self.cfg = cfg
         problem_name = cfg["problem"]
+        self.problem_name = problem_name
         if problem_name not in PROBLEM_REGISTRY:
             raise ValueError(f"Unknown problem type: {problem_name}")
         self.problem: BaseEllipsoidProblem = PROBLEM_REGISTRY[problem_name](cfg)
+        self.wandb = cfg.get("wandb", False)
+        self.training = cfg["training"]
+        self.vmin,self.vmax = None,None
 
     def run(self):
-        x = self.problem.generate_data()
+        data = self.problem.generate_data()
+        n_batches = data.size(0)
+        x = data.view(n_batches, -1)
+        x.requires_grad = True
         node = self.problem.get_node()
         loss_fn = self.problem.get_loss()
 
-        x.requires_grad = True
-        y_init = node.solve(x)[0]
+        # y_init = node.solve(x)[0]
+        
+        if self.wandb:
+            run = wandb.init(
+                project=self.cfg.get("project", "ellipsoid-exp"),
+                name=self.cfg.get("name", f"{self.cfg['problem']}-{self.cfg.get('target', {}).get('m', 'unknown')}"),
+                config=self.cfg,    
+                reinit=True,
+            )
+            wandb.define_metric("outer/loss", summary="min")
+            wandb.define_metric("inner/loss", summary="min")
 
-        run = wandb.init(
-            project=self.cfg.get("project", "ellipsoid-exp"),
-            name=self.cfg.get("name", f"{self.cfg['problem']}-{self.cfg.get('target', {}).get('m', 'unknown')}"),
-            config=self.cfg,    
-            reinit=True,
-            mode="offline" if self.cfg.get("offline", False) else "online"
-        )
-        wandb.define_metric("outer/loss", summary="min")
[…]

### 2025-05-05 src/framework_ellipsoid/functions.py

@@ -21,7 +21,7 @@ def vol_constraint_function_grad(u):
     with torch.enable_grad():
         res = vol_constraint_function(u)
     constr_grad = torch.autograd.grad(res, u)[0]
-
+    return constr_grad
 
 def sa_constraint_function(u, p=1.6075):
     if not torch.is_tensor(u):
@@ -40,19 +40,26 @@ def sa_constraint_function_grad(u):
         res = sa_constraint_function(u)
     return torch.autograd.grad(res, u)[0]
 
-def objective_function(u, X):
+def objective_function(u, X, u_prev=None):
     if not torch.is_tensor(u):
         u = torch.tensor(u).double()
     if not torch.is_tensor(X):
         X = torch.tensor(X).double()
-    L = torch.diag(1 / u[:3]**2).double()
-    R = rotation_matrix_3d(u[3:6]).double()
+    L = torch.diag(1/u[:3]**2).double()
+    R = rotation_matrix_3d(u[3:6]).double() # assumes radians
     A = R @ L @ R.T
     XT_AX = torch.einsum('ji,jk,ki->i', X, A, X)
     b = torch.ones(X.shape[1])
-    return torch.sum((XT_AX - b)**2) / 100
+    if u_prev is None:
+        u_prev = u 
+        return torch.sum((XT_AX - b) ** 2)
+    
+    elif not torch.is_tensor(u_prev):   
+        u_prev = torch.tensor(u_prev).double()
+    res = torch.sum((XT_AX - b) ** 2) + torch.norm(u_prev - u)**2
+    return res
 
-def objective_function_grad(u, X):
+def objective_function_grad(u, X, u_prev=None):
     if torch.is_tensor(u):
         u = u.detach().clone()
     else:
@@ -61,10 +68,26 @@ def objective_function_grad(u, X):
         X = X.detach().clone()
     else:
         X = torch.tensor(X).double()
+    if u_prev is not None:
+        if torch.is_tensor(u_prev):
+            u_prev = u_prev.detach().clone()
+        else:
+            u_prev = torch.tensor(u_prev)
+    else:
+        u_prev = u
     u.requires_grad = True
     with torch.enable_grad():
-        res = objective_function(u, X).double()
-    return torch.autograd.grad(res, u)[0].double()
+        res = objective_function(u, X, u_prev).double()
+    obj_grad = torch.autograd.grad(res, u)[0].double()
+    return obj_grad
+
+def get_bounding_box_dims(points):
[…]

### 2025-05-05 src/framework_ellipsoid/loss.py

@@ -4,7 +4,8 @@ import torch.nn as nn
 from pytorch3d.loss import chamfer_distance
 import alpha_shapes
 import numpy as np
-
+from .plotting import *
+from .utils import rotation_matrix_3d_batch
 
 def A_from_u_batch(u):
     Lambda = torch.diag_embed(1 / u[:, :3] ** 2)
@@ -105,15 +106,12 @@ class SampledProjectionChamferLoss(nn.Module):
         ellipses = homogeneous_projection_batch_torch(A, R)  # (N, 2, 2)
         matrix_sqrts = pos_sqrt(ellipses)
         sampled_pts = sample_pts(matrix_sqrts, m=self.m)
-        return self.chamfer(sampled_pts)
-
-    def chamfer(self, sampled_pts):
         res, _ = chamfer_distance(
             sampled_pts.float(), self.target_pts.float(),
             batch_reduction=None,
             point_reduction="mean"
         )
-        return res.sum()
+        return res
 
 def nearest_boundary_points(original_pts, boundary_np, atol=1e-6):
     """
@@ -139,16 +137,17 @@ def nearest_boundary_points(original_pts, boundary_np, atol=1e-6):
 
 
 class BoundaryProjectionChamferLoss(nn.Module):
-    def __init__(self, views, m=50, alpha=0.0):
+    def __init__(self, views, m=50, sqrt_m=25, alpha=0.0):
         super().__init__()
         self.rot_mats, self.target_pts = views  # target_pts: (N, M, 2)
         self.m = m
+        self.sqrt_m = sqrt_m
         self.alpha = alpha
 
-    def sample_unit_sphere(self, m, device):
+    def sample_unit_sphere(self, sqrt_m, device):
         """Sample m points on the unit sphere S²"""
-        phi = 2 * math.pi * torch.linspace(0.0, 1.0, m, device=device)
-        theta = math.acos(1 - 2 * torch.linspace(0.0, 1.0, m, device=device))
+        phi = 2 * math.pi * torch.linspace(0.0, 1.0, sqrt_m, dtype=torch.double, device=device)
+        theta = torch.acos(1 - 2 * torch.linspace(0.0, 1.0, sqrt_m, dtype=torch.double, device=device))
         phi, theta = torch.meshgrid(phi, theta, indexing="ij")
         x = torch.sin(theta) * torch.cos(phi)
         y = torch.sin(theta) * torch.sin(phi)
@@ -162,34 +161,43 @@ class BoundaryProjectionChamferLoss(nn.Module):
[…]

### 2025-05-05 src/framework_ellipsoid/problems/chamfersample.py

@@ -1,139 +1,0 @@
-import torch
-import torch.nn as nn
-import numpy as np
-import scipy.optimize as opt
-from ddn.pytorch.node import EqConstDeclarativeNode, DeclarativeFunction
-from framework_ellipsoid.base_problem import BaseEllipsoidProblem
-from framework_ellipsoid.functions import (
-    objective_function,
-    objective_function_grad,
-    vol_constraint_function,
-    vol_constraint_function_grad,
-    initialise_u
-)
-import wandb
-from framework_ellipsoid.utils import sample_ellipsoid_surface,build_view_matrices
-from framework_ellipsoid.loss import SampledProjectionChamferLoss,build_target_circle 
-
-class EllipsoidConstrainedProjectionNode(EqConstDeclarativeNode):
-    def __init__(self, m):
-        super().__init__(eps=1e-6)
-        self.m = m
-        self.n = (3 * m,)
-        self.u_prev = None
-
-    def objective(self, xs, y):
-        n_batches = xs.size(0)
-        data = xs.view(n_batches, 3, self.m)
-        y = y.view(n_batches, 6)
-        semiaxes = y[:, :3]
-        angles = y[:, 3:]
-        L_diag = 1 / semiaxes ** 2
-        L = torch.diag_embed(L_diag)
-
-        angles_rad = torch.deg2rad(angles)
-        cos = torch.cos(angles_rad)
-        sin = torch.sin(angles_rad)
-
-        R = torch.zeros((n_batches, 3, 3), dtype=torch.double, device=xs.device)
-        cy, cp, cr = cos[:, 0], cos[:, 1], cos[:, 2]
-        sy, sp, sr = sin[:, 0], sin[:, 1], sin[:, 2]
-
-        R[:, 0, 0] = cy * cp
-        R[:, 0, 1] = cy * sp * sr - sy * cr
-        R[:, 0, 2] = cy * sp * cr + sy * sr
-        R[:, 1, 0] = sy * cp
-        R[:, 1, 1] = sy * sp * sr + cy * cr
-        R[:, 1, 2] = sy * sp * cr - cy * sr
-        R[:, 2, 0] = -sp
-        R[:, 2, 1] = cp * sr
-        R[:, 2, 2] = cp * cr
-
-        A = torch.bmm(R, torch.bmm(L, R.transpose(1, 2)))
-        XT_AX = torch.einsum('bji,bjk,bki->bi', data, A, data)
-
[…]

### 2025-05-05 src/framework_ellipsoid/problems/chamfersampled.py

@@ -1,38 +1,29 @@
 import torch
 from framework_ellipsoid.base_problem import BaseEllipsoidProblem
 from framework_ellipsoid.utils import sample_ellipsoid_surface,build_view_matrices
+from framework_ellipsoid.loss import SampledProjectionChamferLoss,build_target_circle 
 from .innernodes import UnitVolConstrainedProjectionNode, EllipseConstrainedProjectionFunction
-from framework_ellipsoid.loss import BoundaryProjectionChamferLoss,build_target_circle 
-from framework_ellipsoid.utils import ellipsoid_surface_area, ellipsoid_volume
 
-class ChamferBoundaryProblem(BaseEllipsoidProblem):
+class ChamferSampledProblem(BaseEllipsoidProblem):
     def __init__(self, cfg):
         super().__init__(cfg)
         self.rot_mats = build_view_matrices(cfg)
-
-        target_cfg = cfg.get("target", {})
-        radius = target_cfg.get("radius", 1.0)
-        m_pts = target_cfg.get("m", 50)
-        self.sqrt_m_pts = target_cfg.get("sqrt_m", 50)
-
-        target_pts = build_target_circle(radius, m_pts).expand(self.rot_mats.size(0), -1, -1)
+        target_cfg = cfg["target"]
+        target_pts = build_target_circle(target_cfg["radius"], target_cfg["m"]).expand(self.rot_mats.size(0), -1, -1)
         self.views = (self.rot_mats, target_pts)
-
         initial_angles_deg = torch.tensor(cfg.get("initial_angles", [0, 0, 0]), dtype=torch.double)
         self.initial_angles = torch.deg2rad(initial_angles_deg)
-        self.m_pts = m_pts
 
     def generate_data(self):
         a, b, c = self.initial_axes
         yaw, pitch, roll = self.initial_angles
-        print(f"Initial volume is {ellipsoid_volume(a,b,c)} and initial surface area {ellipsoid_surface_area(a,b,c)}")
         return sample_ellipsoid_surface(self.sqrt_m, a, b, c, yaw, pitch, roll, self.nu)
 
     def get_node(self):
         return UnitVolConstrainedProjectionNode(self.m, self.wandb)
 
     def get_loss(self):
-        return BoundaryProjectionChamferLoss(self.views, m=self.m_pts, sqrt_m=self.sqrt_m_pts)
+        return SampledProjectionChamferLoss(self.views, m=self.cfg["target"]["m"])
 
[…]

### 2025-05-05 src/framework_ellipsoid/problems/volumeconstrained.py

@@ -1,101 +1,7 @@
 import torch
 import torch.nn as nn
-import numpy as np
-import scipy.optimize as opt
-from ddn.pytorch.node import EqConstDeclarativeNode, DeclarativeFunction
 from framework_ellipsoid.base_problem import BaseEllipsoidProblem
-from framework_ellipsoid.functions import (
-    objective_function,
-    objective_function_grad,
-    vol_constraint_function,
-    vol_constraint_function_grad,
-    initialise_u
-)
-import wandb
-
-class EllipsoidConstrainedProjectionNode(EqConstDeclarativeNode):
-    def __init__(self, m):
-        super().__init__(eps=1e-4)
-        self.m = m
-        self.n = (3 * m,)
-
-    def objective(self, xs: torch.Tensor, y: torch.Tensor) -> torch.Tensor:
-        n_batches = xs.size(0)
-        data = xs.view(n_batches, 3, self.m)
-        y = y.view(n_batches, 6)
-        semiaxes = y[:, :3]
-        angles = y[:, 3:]
-        L_diag = 1 / semiaxes ** 2
-        L = torch.diag_embed(L_diag)
-
-        angles_rad = torch.deg2rad(angles)
-        cos = torch.cos(angles_rad)
-        sin = torch.sin(angles_rad)
-
-        R = torch.zeros((n_batches, 3, 3), dtype=torch.double, device=xs.device)
-        cy, cp, cr = cos[:, 0], cos[:, 1], cos[:, 2]
-        sy, sp, sr = sin[:, 0], sin[:, 1], sin[:, 2]
-
-        R[:, 0, 0] = cy * cp
-        R[:, 0, 1] = cy * sp * sr - sy * cr
-        R[:, 0, 2] = cy * sp * cr + sy * sr
-        R[:, 1, 0] = sy * cp
-        R[:, 1, 1] = sy * sp * sr + cy * cr
-        R[:, 1, 2] = sy * sp * cr - cy * sr
-        R[:, 2, 0] = -sp
-        R[:, 2, 1] = cp * sr
-        R[:, 2, 2] = cp * cr
-
-        A = torch.bmm(R, torch.bmm(L, R.transpose(1, 2)))
-        XT_AX = torch.einsum('bji,bjk,bki->bi', data, A, data)
-        return ((XT_AX - 1) ** 2).sum(dim=1)
-
[…]

### 2025-05-05 src/framework_ellipsoid/script.py

@@ -1,35 +1,42 @@
 import json
 import argparse
 from copy import deepcopy
-from experiments import EllipsoidExperiment
+from .experiments import EllipsoidExperiment
+import warnings
+warnings.filterwarnings('ignore')
 
 def load_config(path):
     with open(path, "r") as f:
         return json.load(f)
 
 def main():
-    parser = argparse.ArgumentParser()
-    parser.add_argument("--config", type=str, required=True, help="Path to base config")
-    parser.add_argument("--param", type=str, required=True, help="Parameter to vary, e.g. target.m")
-    parser.add_argument("--values", type=str, required=True, help="Comma-separated list of values")
-    args = parser.parse_args()
+    # parser = argparse.ArgumentParser()
+    # parser.add_argument("--config", type=str, required=True, help="Path to base config")
+    # parser.add_argument("--param", type=str, required=True, help="Parameter to vary, e.g. target.m")
+    # parser.add_argument("--values", type=str, required=True, help="Comma-separated list of values")
+    # args = parser.parse_args()
 
-    base_cfg = load_config(args.config)
-    values = [eval(v) for v in args.values.split(",")]
+    # base_cfg = load_config(args.config)
+    # values = [eval(v) for v in args.values.split(",")]
 
-    for v in values:
-        cfg = deepcopy(base_cfg)
+    # for v in values:
+    #     cfg = deepcopy(base_cfg)
 
-        # Support nested keys like "target.m"
-        keys = args.param.split(".")
-        sub_cfg = cfg
-        for k in keys[:-1]:
-            sub_cfg = sub_cfg[k]
-        sub_cfg[keys[-1]] = v
-        param_value_str = str(v).replace(" ", "").replace("[", "").replace("]", "").replace(",", "-")
-        cfg["name"] = f"{cfg['problem']}-{args.param.replace('.', '_')}-{param_value_str}"
+    #     # Support nested keys like "target.m"
+    #     keys = args.param.split(".")
+    #     sub_cfg = cfg
+    #     for k in keys[:-1]:
+    #         sub_cfg = sub_cfg[k]
+    #     sub_cfg[keys[-1]] = v
+    #     param_value_str = str(v).replace(" ", "").replace("[", "").replace("]", "").replace(",", "-")
+    #     cfg["name"] = f"{cfg['problem']}-{args.param.replace('.', '_')}-{param_value_str}"
 
-        print(f"Running with {args.param} = {v}")
[…]

### 2025-05-05 src/framework_ellipsoid/utils.py

@@ -1,22 +1,36 @@
 import torch
 import math
-from .utils import rotation_matrix_3d
 import numpy as np
 from torch import cos, sin
 import torch.nn.functional as F
-from framework_ellipsoid.loss import rotation_matrix_3d_batch
+from scipy.interpolate import interp1d, interp2d
 
-def sample_ellipsoid_surface(sqrt_m, a, b, c, yaw, pitch, roll, noise_std=1e-4):
-    phi = 2.0 * math.pi * torch.linspace(0.0, 1.0, sqrt_m).double()
-    theta = math.pi * torch.linspace(0.0, 1.0, sqrt_m).double()
-    phi, theta = torch.meshgrid(phi, theta, indexing='ij')
-    x = a * torch.sin(theta) * torch.cos(phi)
-    y = b * torch.sin(theta) * torch.sin(phi)
-    z = c * torch.cos(theta)
-    coords = torch.stack((x.flatten(), y.flatten(), z.flatten()), dim=0)
-    R = rotation_matrix_3d(torch.tensor([yaw, pitch, roll]))
-    rotated = R @ coords
-    noisy = rotated + noise_std * torch.randn_like(rotated)
+def sample_ellipsoid_surface(sqrt_m, a, b, c, yaw, pitch, roll, noise_std=1e-4, uniform=True):
+    if uniform:
+        m = sqrt_m * sqrt_m
+        def ellipsoid_func(t, u):
+            x = a * torch.sin(t) * torch.cos(u)
+            y = b * torch.sin(t) * torch.sin(u)
+            z = c * torch.cos(t)
+            return torch.stack([x, y, z], dim=0)  # (3, ...)
+        t_min, t_max = 0, torch.pi
+        u_min, u_max = 0, 2 * torch.pi
+        coords = r_surface(m, ellipsoid_func, t_min, t_max, u_min, u_max, 100, 100).double()
+
+    else:
+        phi = 2.0 * math.pi * torch.linspace(0.0, 1.0, sqrt_m).double()
+        theta = math.pi * torch.linspace(0.0, 1.0, sqrt_m).double()
+        phi, theta = torch.meshgrid(phi, theta, indexing='ij')
+        x = a * torch.sin(theta) * torch.cos(phi)
+        y = b * torch.sin(theta) * torch.sin(phi)
+        z = c * torch.cos(theta)
[…]

### 2025-05-05 src/framework_mesh/dataloader.py

@@ -5,7 +5,7 @@ from pytorch3d.structures import Meshes
 from .io import *
 
 class DataLoader:
-    def __init__(self, config) -> None:
+    def __init__(self, config, device=torch.device("cpu")) -> None:
 
         if isinstance(config, str) and os.path.exists(config):
             with open(config, "r") as f:
@@ -14,7 +14,9 @@ class DataLoader:
             self.cfg = config
         else:
             raise ValueError("Invalid experiment_config. Must be a path or dict.")
-
+        
+        self.device = device
+        
         # Set up paths and other config settings
         self.mesh_dir = self.cfg["paths"]["mesh_dir"]
         self.mesh_res = self.cfg["paths"]["mesh_res"]
@@ -73,7 +75,7 @@ class DataLoader:
         if not os.path.exists(mesh_path):
             return None
 
-        verts, faces, _ = load_obj(mesh_path)
+        verts, faces, _ = load_obj(mesh_path, device=self.device)
         return Meshes(verts=[verts], faces=[faces.verts_idx])
 
     def get_gt_mesh(self, mesh_name):
@@ -85,5 +87,5 @@ class DataLoader:
         if not os.path.exists(gt_path):
             return None
 
-        verts, faces, _ = load_obj(gt_path)
+        verts, faces, _ = load_obj(gt_path, device=self.device)
         return Meshes(verts=[verts], faces=[faces.verts_idx])

### 2025-05-05 src/framework_mesh/experimentrunner.py

@@ -4,11 +4,13 @@ from .node import *
 from .chamfer import *
 from .utils import *
 from .io import *
+from .gradient import *
 from typing import List, Tuple, Union
 import random
 import wandb
 import os
 from tqdm import trange
+
 
 class ExperimentRunner:
     def __init__(self, experiment_config: Union[str, dict], data_loader: DataLoader) -> None:
@@ -20,7 +22,6 @@ class ExperimentRunner:
             self.cfg = experiment_config
         else:
             raise ValueError("Invalid experiment_config. Must be a path or dict.")
-
         # Store the experiment details
         self.project = self.cfg["project"]
         self.experiment_name = self.cfg["name"]
@@ -53,10 +54,20 @@ class ExperimentRunner:
         
         self.wandb = self.cfg["wandb"]
 
-    def run(self):
-        return self.pipeline(self.src_mesh, self.target_meshes)
+        self.smoothing = self.cfg["gradient"]["smoothing"]
+        self.smoothing_method = self.cfg["gradient"]["method"]
+        self.smoothing_k = self.cfg["gradient"]["k"]
+        self.smoothing_constrained = self.cfg["gradient"]["constrained"]
+        self.smoothing_debug = self.cfg["gradient"]["debug"]
+        if self.smoothing_debug:
+            print("Smoothing debug is on")
 
-    def pipeline(self, src_name: str, tgt_names: List[str]):
+    def run(self):
+        device = self.data_loader.device
+        print(f"Running on device: {device}")
+        return self.pipeline(self.src_mesh, self.target_meshes, device)
+
+    def pipeline(self, src_name: str, tgt_names: List[str], device: torch.device):
         """Pipeline function to run the experiment on a source mesh and a list of target meshes."""
         if self.wandb:
             run = wandb.init(
@@ -65,12 +76,8 @@ class ExperimentRunner:
                 notes=self.experiment_description,
                 group="clean",
                 config={
-                    "iters":    self.n_iters,
-                    "lr":       self.lr,
-                    "momentum": self.momentum,
-                    "source": src_name,
-                    "targets": tgt_names[0] if len(tgt_names) == 1 else tgt_names,
-                    "num_views": self.num_views,
+                    "data":    self.data_loader.cfg,
+                    "experiment":       self.cfg
                 }
             )
[…]

### 2025-05-05 src/framework_mesh/gradient.py

@@ -1,0 +1,166 @@
+import torch
+from collections import deque
+
+def build_edge_lists(faces, device=None):
+    e = torch.vstack([faces[:, [0, 1]], faces[:, [1, 2]], faces[:, [2, 0]]])
+    e = torch.cat([e, e.flip(1)]).unique(dim=0)
+    if device: e = e.to(device)
+    return e[:, 0], e[:, 1]
+
+def bfs_hop_distance(V, edge_src, edge_dst, boundary_idx, k_max=10):
+    B = boundary_idx.numel()
+    D = torch.full((V, B), k_max + 1, dtype=torch.int16)
+    frontier = [deque([b.item()]) for b in boundary_idx]
+    visited  = [set([b.item()])   for b in boundary_idx]
+
+    for dist in range(k_max + 1):
+        for b, q in enumerate(frontier):
+            for _ in range(len(q)):
+                v = q.popleft()
+                D[v, b] = dist
+                for u in edge_dst[edge_src == v].tolist():
+                    if u not in visited[b]:
+                        visited[b].add(u)
+                        q.append(u)
+    return D
+
+
+# -------------------------------
+# 1. Jacobi smoother
+# -------------------------------
+def make_jacobi_hook(edge_src, edge_dst, boundary_mask, k=5, constrained=False, debug=False):
+    V = boundary_mask.size(0)
+    onesE = torch.ones(edge_src.size(0), 1, device=edge_src.device)
+
+    def hook(grad_in):
+        g = grad_in.view(-1, 3).clone()
+        for _ in range(k):
+            g_nb = torch.zeros_like(g)
+            g_nb.scatter_add_(0, edge_dst[:, None].expand(-1, 3), g[edge_src])
+            deg = torch.zeros(V, 1, device=g.device)
+            deg.scatter_add_(0, edge_dst[:, None], onesE)
+            g_nb += g
+            deg += 1.0
+            g_nb = g_nb / (deg.clamp_min_(1.0)+1)
+            if constrained:
+                g = torch.where(boundary_mask[:, None], g, g_nb)
+            else:
+                g = g_nb
+
+        if debug:
+            g_diff = (g - grad_in.view(-1, 3)).abs()
+            print("[Jacobi] Δg > 1e-6:", (g_diff.norm(dim=-1) > 1e-6).sum().item(),
+                  "| mean Δg:", g_diff.mean().item())
+
+        return g.view_as(grad_in)
+
+    return hook
+
+
[…]

### 2025-05-05 src/framework_mesh/io.py

@@ -8,10 +8,12 @@ from cv2.typing import MatLike
 from scipy.interpolate import splprep, splev
 
 
-def load_camera_matrices(path, matrix_types=None):
+def load_camera_matrices(path, matrix_types=None, device=torch.device("cpu")):
     cameras = defaultdict(dict)
     cam_names = set()
     file_pattern = re.compile(r"^Cam_([A-Za-z]+_\d+)_(K|RT|P)\.npy$")
+
+    device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
 
     if matrix_types is not None:
         matrix_types = set(matrix_types)
@@ -22,7 +24,7 @@ def load_camera_matrices(path, matrix_types=None):
             cam_name, matrix_type = match.groups()
             if matrix_types is None or matrix_type in matrix_types:
                 filepath = os.path.join(path, filename)
-                cameras[cam_name][matrix_type] = torch.tensor(np.load(filepath))
+                cameras[cam_name][matrix_type] = torch.tensor(np.load(filepath), device=device)
                 cam_names.add(cam_name)
 
     cam_names = sorted(cam_names)
@@ -58,7 +60,7 @@ def canny_edge_map(img: MatLike, options):
     # return edge map
     return edge_map
 
-def load_edgemaps(renders: dict, edgemap_options: dict):
+def load_edgemaps(renders: dict, edgemap_options: dict, device=torch.device("cpu")):
     edgemaps = {}
     edgemaps_len = {}
 

### 2025-05-05 src/framework_mesh/script.py

@@ -1,24 +1,46 @@
-from .dataloader import *
-from .experimentrunner import *
+import torch
+import json
+from itertools import product
+from .dataloader import DataLoader
+from .experimentrunner import ExperimentRunner
 import warnings
-warnings.filterwarnings('ignore')
+warnings.filterwarnings("ignore")
+
 
 def main():
-    data_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/obliqueconfig_local.json"
+    device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
+    device = torch.device("cpu")
+    
+    data_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/skyconfig_local.json"
+    exp_config_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/exp_pipeline.json"
+
     with open(data_path, 'r') as f:
-        config = json.load(f)
+        data_config = json.load(f)
 
-    # Modify the 'mesh_res' value
-    config['paths']['mesh_res'] = 2
-    dataloader = DataLoader(config)
-    runner = ExperimentRunner("/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/exp_oblique.json", dataloader)
-    runner.run()
+    with open(exp_config_path, 'r') as f:
+        exp_config = json.load(f)
 
-    # config['paths']['mesh_res'] = 3
-    # dataloader = DataLoader(config)
-    # runner = ExperimentRunner("/home/jeffrey/honours/src/experiments/exp_pipeline.json", dataloader)
-    # runner.run()
+    methods = ["jacobi"]
+    constrained_options = [False]
+    mesh_resolutions = [2]
 
+    for mesh_res in mesh_resolutions:
+        # Update mesh resolution
+        data_config['paths']['mesh_res'] = mesh_res
+        dataloader = DataLoader(data_config, device)
+
+        for method in methods:
+            for constrained in constrained_options:
+                # Update gradient method and constraint setting
+                exp_config['gradient']['method'] = method
+                exp_config['gradient']['constrained'] = constrained
+
+                # Update experiment name to be descriptive
+                exp_config['name'] = f"pipeline_{method}_constrained_{constrained}_res_{mesh_res}"
+
+                # Create runner and run
+                runner = ExperimentRunner(exp_config, dataloader)
+                runner.run()
 
 if __name__ == "__main__":
     main()

### 2025-05-05 src/framework_mesh/utils.py

@@ -42,28 +42,28 @@ def create_padded_tensor(vertices, vert2mesh, max_V, B):
         padded[i, :num_vertices, :] = mesh_vertices
     return padded
 
-# def plot_vertices(verts_list):
-#     fig = plt.figure(figsize=(5, 5))
-#     ax = fig.add_subplot(111, projection='3d')
+def plot_vertices(verts_list):
+    fig = plt.figure(figsize=(5, 5))
+    ax = fig.add_subplot(111, projection='3d')
     
-#     if not isinstance(verts_list, list):
-#         x, y, z = verts_list.clone().detach().cpu().squeeze().unbind(1)
-#         ax.scatter3D(x, z, -y)
-#     else:
-#         colors = ['b',  'g', 'r' , 'm', 'c', 'y']  # Define some colors for different sets
-#         marker_size = 5  # Make points smaller
+    if not isinstance(verts_list, list):
+        x, y, z = verts_list.clone().detach().cpu().squeeze().unbind(1)
+        ax.scatter3D(x, z, -y)
+    else:
+        colors = ['b',  'g', 'r' , 'm', 'c', 'y']  # Define some colors for different sets
+        marker_size = 5  # Make points smaller
     
-#         for i, verts in enumerate(verts_list):
-#             x, y, z = verts.clone().detach().cpu().squeeze().unbind(1)
-#             ax.scatter3D(x, z, -y, color=colors[i % len(colors)], s=marker_size, label=f"Set {i+1}")    
+        for i, verts in enumerate(verts_list):
+            x, y, z = verts.clone().detach().cpu().squeeze().unbind(1)
+            ax.scatter3D(x, z, -y, color=colors[i % len(colors)], s=marker_size, label=f"Set {i+1}")    
     
-#     ax.set_xlabel('x')
-#     ax.set_ylabel('z')
-#     ax.set_zlabel('y')
-#     ax.set_aspect("equal")
-#     ax.view_init(190, 30)
-#     ax.legend()
-#     plt.show()
+    ax.set_xlabel('x')
+    ax.set_ylabel('z')
+    ax.set_zlabel('y')
+    ax.set_aspect("equal")
+    ax.view_init(190, 30)
+    ax.legend()
+    plt.show()
 
 def plot_projections(projverts, projmats, edgemaps):
     plt.ioff()  # Disable interactive mode
[…]

### 2025-05-05 src/framework_mesh/experimentrunner.py
ready to run experiments:

@@ -89,6 +89,7 @@ class ExperimentRunner:
 
         src_mesh = self.get_mesh(src_name).to(device)
         for tgt_name in tgt_names:
+            print(f"Target: {tgt_name}")
             tgt_mesh = self.get_gt_mesh(tgt_name).to(device)
 
             # Use encapsulated loading logic

### 2025-05-05 src/framework_mesh/script.py

@@ -11,15 +11,15 @@ def main():
     device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
     device = torch.device("cpu")
     
-    data_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/skyconfig_local.json"
-    exp_config_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/exp_pipeline.json"
+    data_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/obliqueconfig_local.json"
+    exp_config_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/exp_oblique.json"
 
     with open(data_path, 'r') as f:
         data_config = json.load(f)
 
     with open(exp_config_path, 'r') as f:
         exp_config = json.load(f)
-
+    
     methods = ["jacobi"]
     constrained_options = [False]
     mesh_resolutions = [2]

### 2025-05-05 src/framework_mesh/script.py
ready to run on cupertino

@@ -1,11 +1,9 @@
 import torch
 import json
-from itertools import product
 from .dataloader import DataLoader
 from .experimentrunner import ExperimentRunner
 import warnings
 warnings.filterwarnings("ignore")
-
 
 def main():
     device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
@@ -19,28 +17,29 @@ def main():
 
     with open(exp_config_path, 'r') as f:
         exp_config = json.load(f)
-    
-    methods = ["jacobi"]
-    constrained_options = [False]
-    mesh_resolutions = [2]
+
+    smoothing = exp_config.get("gradient", {}).get("smoothing", False)
+    mesh_resolutions = [2, 3]
 
     for mesh_res in mesh_resolutions:
-        # Update mesh resolution
         data_config['paths']['mesh_res'] = mesh_res
         dataloader = DataLoader(data_config, device)
 
-        for method in methods:
-            for constrained in constrained_options:
-                # Update gradient method and constraint setting
-                exp_config['gradient']['method'] = method
-                exp_config['gradient']['constrained'] = constrained
+        if not smoothing:
+            exp_config['name'] = f"pipeline_nosmooth_res_{mesh_res}"
+            print(f"[RUN] smoothing=False | mesh_res={mesh_res}")
+            runner = ExperimentRunner(exp_config, dataloader)
+            runner.run()
 
-                # Update experiment name to be descriptive
-                exp_config['name'] = f"pipeline_{method}_constrained_{constrained}_res_{mesh_res}"
-
-                # Create runner and run
-                runner = ExperimentRunner(exp_config, dataloader)
-                runner.run()
+        else:
+            method = "jacobi"
+            constrained = False
+            exp_config['gradient']['method'] = method
+            exp_config['gradient']['constrained'] = constrained
+            exp_config['name'] = f"pipeline_{method}_constrained_{constrained}_res_{mesh_res}"
+            print(f"[RUN] smoothing=True | method={method} | constrained={constrained} | mesh_res={mesh_res}")
+            runner = ExperimentRunner(exp_config, dataloader)
+            runner.run()
 
 if __name__ == "__main__":
     main()

### 2025-05-05 src/framework_mesh/functions.py
added stuff for oblique gpu

@@ -6,11 +6,10 @@ import numpy as np
 
 torch.autograd.set_detect_anomaly(True)
 
-def least_squares(u0, tgt_vtxs):
+def least_squares(u0, tgt_vtxs, device=torch.device("cpu")):
     """
     u0 are vertices
     """
-    device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
     if not torch.is_tensor(u0):
         u0 = torch.tensor(u0, device=device)
     if not torch.is_tensor(tgt_vtxs):
@@ -18,8 +17,7 @@ def least_squares(u0, tgt_vtxs):
     res = torch.square(u0 - tgt_vtxs.flatten()).sum()
     return res.double()
 
-def least_squares_grad(u0, tgt_vtxs):
-    device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
+def least_squares_grad(u0, tgt_vtxs, device=torch.device("cpu")):
     if torch.is_tensor(u0):
         u0 = u0.detach().clone()
     else:
@@ -44,7 +42,7 @@ def calculate_volume(vertices, faces):
     return volume.abs()
 
 
-def volume_constraint(x, faces, tgt_vol):
+def volume_constraint(x, faces, tgt_vol, device=torch.device("cpu")):
     """
     Calculate the volume of a mesh using PyTorch tensors.
     Args:
@@ -53,8 +51,6 @@ def volume_constraint(x, faces, tgt_vol):
     Returns:
         volume: Total volume of the mesh as a PyTorch scalar
     """
-    device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
-
     if not torch.is_tensor(x):
         x = torch.tensor(x, device=device)
     if not torch.is_tensor(faces):
@@ -68,8 +64,7 @@ def volume_constraint(x, faces, tgt_vol):
     res = volume.abs() - tgt_vol
     return res.double()
 
-def volume_constraint_grad(x, faces):
-    device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
+def volume_constraint_grad(x, faces, device=torch.device("cpu")):
     if torch.is_tensor(x):
         x = x.detach().clone()
     else:

### 2025-05-05 src/framework_mesh/io.py

@@ -13,8 +13,6 @@ def load_camera_matrices(path, matrix_types=None, device=torch.device("cpu")):
     cam_names = set()
     file_pattern = re.compile(r"^Cam_([A-Za-z]+_\d+)_(K|RT|P)\.npy$")
 
-    device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
-
     if matrix_types is not None:
         matrix_types = set(matrix_types)
 

### 2025-05-05 src/framework_mesh/chamfer.py
switched to single direction chamfer

@@ -118,7 +118,8 @@ class PyTorchChamferLoss(nn.Module):
                         x_lengths=boundary_lengths[b].long(),
                         y_lengths=self.edgemaps_len[b].long(),
                         batch_reduction="mean",
-                        point_reduction="mean")
+                        point_reduction="mean",
+                        single_directional=True)
             chamfer_loss[b] = res.sum()
 
         return chamfer_loss.double()

### 2025-05-05 src/framework_mesh/dataloader.py

@@ -76,6 +76,8 @@ class DataLoader:
             return None
 
         verts, faces, _ = load_obj(mesh_path, device=self.device)
+        verts = torch.stack([verts[:, 0], -verts[:, 2], verts[:, 1]], dim=1)
+
         return Meshes(verts=[verts], faces=[faces.verts_idx])
 
     def get_gt_mesh(self, mesh_name):
@@ -88,4 +90,6 @@ class DataLoader:
             return None
 
         verts, faces, _ = load_obj(gt_path, device=self.device)
+        verts = torch.stack([verts[:, 0], -verts[:, 2], verts[:, 1]], dim=1)
+
         return Meshes(verts=[verts], faces=[faces.verts_idx])

### 2025-05-05 src/framework_mesh/script.py

@@ -26,7 +26,7 @@ def main():
         dataloader = DataLoader(data_config, device)
 
         if not smoothing:
-            exp_config['name'] = f"pipeline_nosmooth_res_{mesh_res}"
+            exp_config['name'] = f"nosmooth_res_{mesh_res}"
             print(f"[RUN] smoothing=False | mesh_res={mesh_res}")
             runner = ExperimentRunner(exp_config, dataloader)
             runner.run()
@@ -36,7 +36,7 @@ def main():
             constrained = False
             exp_config['gradient']['method'] = method
             exp_config['gradient']['constrained'] = constrained
-            exp_config['name'] = f"pipeline_{method}_constrained_{constrained}_res_{mesh_res}"
+            exp_config['name'] = f"smooth_{method}_constrained_{constrained}_res_{mesh_res}"
             print(f"[RUN] smoothing=True | method={method} | constrained={constrained} | mesh_res={mesh_res}")
             runner = ExperimentRunner(exp_config, dataloader)
             runner.run()

### 2025-05-05 src/framework_mesh/experimentrunner.py
committing before long run

@@ -134,12 +134,12 @@ class ExperimentRunner:
         V_max = src.num_verts_per_mesh().max().item()
         boundary_mask = torch.zeros(V_max, dtype=torch.bool, device=device)
 
-        if self.smoothing:
-            faces = src[0].faces_packed().to(device)
-            edge_src, edge_dst = build_edge_lists(faces, device)
+        faces = src[0].faces_packed().to(device)
+        edge_src, edge_dst = build_edge_lists(faces, device)
+        all_idx = torch.arange(V_max, device=device)
+        D_all = bfs_hop_distance(V_max, edge_src, edge_dst, all_idx, k_max=10)
 
-            all_idx = torch.arange(V_max, device=device)
-            D_all = bfs_hop_distance(V_max, edge_src, edge_dst, all_idx, k_max=10)
+        if self.smoothing:
             hook = select_hook(
                 method=self.smoothing_method,         # "jacobi", "invhop", or "khop"
                 edge_src=edge_src,
@@ -158,6 +158,8 @@ class ExperimentRunner:
         a,b = a[0], b[0]
         projectionplot = plot_projections(verts.detach().squeeze().double(), gt_projmats, gt_edgemap_info)
         cmin,cmax = None,None
+        bmin,bmax = None,None
+
         initheatmap,cmin,cmax = create_heatmap(Meshes(verts=verts.detach(), faces=src[0].faces_packed().unsqueeze(0)), tgt[0], cmin, cmax)
         if self.wandb:
             wandb.log({"plt/projections": wandb.Image(projectionplot),
@@ -166,7 +168,7 @@ class ExperimentRunner:
         min_loss = float("inf")
         best_verts = None
     
-        pbar = trange(n_iters, desc="Training", leave=True)  # Always visible
+        pbar = trange(n_iters, desc="Training", leave=True, mininterval=5)  # Always visible
         for i in pbar:
             optimiser.zero_grad(set_to_none=True)
             node.iter = i + step_offset
@@ -200,13 +202,22 @@ class ExperimentRunner:
                 print(f"{i:4d} Loss: {colour}{loss.item():.3f}{bcolors.ENDC} Volume: {calculate_volume(projverts[0], src[0].faces_packed()):.3f}")
                 print(f"GT Chamfer: [{', '.join(f'{x:.3f}' for x in chamfer_gt(projverts, src, tgt))}] "
                     f"GT IoU: [{', '.join(f'{x:.3f}' for x in iou_gt(projverts, src, tgt))}]")
[…]

### 2025-05-05 src/framework_mesh/gradient.py

@@ -34,18 +34,23 @@ def make_jacobi_hook(edge_src, edge_dst, boundary_mask, k=5, constrained=False, 
 
     def hook(grad_in):
         g = grad_in.view(-1, 3).clone()
+        og_g = grad_in.view(-1, 3).clone()
         for _ in range(k):
             g_nb = torch.zeros_like(g)
             g_nb.scatter_add_(0, edge_dst[:, None].expand(-1, 3), g[edge_src])
             deg = torch.zeros(V, 1, device=g.device)
             deg.scatter_add_(0, edge_dst[:, None], onesE)
-            g_nb += g
+            if constrained:
+                g_nb += og_g
+            else:
+                g_nb += g
             deg += 1.0
             g_nb = g_nb / (deg.clamp_min_(1.0)+1)
-            if constrained:
-                g = torch.where(boundary_mask[:, None], g, g_nb)
-            else:
-                g = g_nb
+            g = g_nb
+            # if constrained:
+            #     g = torch.where(boundary_mask[:, None], g, g_nb)
+            # else:
+            #     g = g_nb
 
         if debug:
             g_diff = (g - grad_in.view(-1, 3)).abs()

### 2025-05-05 src/framework_mesh/script.py

@@ -16,30 +16,33 @@ def main():
         data_config = json.load(f)
 
     with open(exp_config_path, 'r') as f:
-        exp_config = json.load(f)
+        exp_config_base = json.load(f)
 
-    smoothing = exp_config.get("gradient", {}).get("smoothing", False)
     mesh_resolutions = [2, 3]
+    smoothing_options = [False, True]
 
     for mesh_res in mesh_resolutions:
-        data_config['paths']['mesh_res'] = mesh_res
-        dataloader = DataLoader(data_config, device)
+        for smoothing in smoothing_options:
+            data_config['paths']['mesh_res'] = mesh_res
+            dataloader = DataLoader(data_config, device)
 
-        if not smoothing:
-            exp_config['name'] = f"nosmooth_res_{mesh_res}"
-            print(f"[RUN] smoothing=False | mesh_res={mesh_res}")
-            runner = ExperimentRunner(exp_config, dataloader)
-            runner.run()
+            exp_config = json.loads(json.dumps(exp_config_base))  # deep copy
+            exp_config.setdefault("gradient", {})["smoothing"] = smoothing
 
-        else:
-            method = "jacobi"
-            constrained = False
-            exp_config['gradient']['method'] = method
-            exp_config['gradient']['constrained'] = constrained
-            exp_config['name'] = f"smooth_{method}_constrained_{constrained}_res_{mesh_res}"
-            print(f"[RUN] smoothing=True | method={method} | constrained={constrained} | mesh_res={mesh_res}")
-            runner = ExperimentRunner(exp_config, dataloader)
-            runner.run()
+            if not smoothing:
+                exp_config['name'] = f"nosmooth_res_{mesh_res}"
+                print(f"[RUN] smoothing=False | mesh_res={mesh_res}")
+                runner = ExperimentRunner(exp_config, dataloader)
+                runner.run()
+            else:
+                method = "jacobi"
+                for constrained in [False, True]:
+                    exp_config['gradient']['method'] = method
+                    exp_config['gradient']['constrained'] = constrained
+                    exp_config['name'] = f"smooth_{method}_constrained_{constrained}_res_{mesh_res}"
+                    print(f"[RUN] smoothing=True | method={method} | constrained={constrained} | mesh_res={mesh_res}")
+                    runner = ExperimentRunner(exp_config, dataloader)
+                    runner.run()
 
 if __name__ == "__main__":
     main()

### 2025-05-05 src/framework_mesh/utils.py

@@ -385,3 +385,53 @@ def create_heatmap(src, tgt, cmin=None, cmax=None):
     # plt.show()
 
     return fig, cmin, cmax
+
+
+def compute_boundary_distance_heatmap(src: Meshes, boundary_mask, D_all, cmin=None, cmax=None):
+     """
+     Computes a 3D heatmap coloring each vertex in `src` based on its geodesic distance to the nearest boundary vertex.
+     """
+     verts3d = src.verts_padded()[0]  # (V, 3)
+     V = verts3d.shape[0]
+     faces = src.faces_packed()
+ 
+     with torch.no_grad():
+         boundary_ids = torch.where(boundary_mask[:V])[0]
+         D = D_all[:V, :V]
+         boundary_dists = D[:, boundary_ids]  # (V, B)
+         min_dist, _ = torch.min(boundary_dists, dim=1)
+         heatmap_values = min_dist.float().cpu().numpy()
+ 
+     if cmin is None or cmax is None:
+         max_val = np.percentile(heatmap_values, 95)
+         cmin, cmax = 0, max_val
+ 
+     mesh = trimesh.Trimesh(vertices=verts3d.cpu().numpy(), faces=faces.cpu().numpy())
+     i, j, k = mesh.faces.T
+ 
+     fig = go.Figure(data=[
+         go.Mesh3d(
+             x=mesh.vertices[:, 0],
+             y=mesh.vertices[:, 1],
+             z=mesh.vertices[:, 2],
+             i=i, j=j, k=k,
+             intensity=heatmap_values,
+             colorscale='Greens',
+             cmin=cmin, cmax=cmax,
+             colorbar=dict(title='Hop Distance to Boundary'),
+             showscale=True,
+             lighting=dict(ambient=0.8, diffuse=0.9),
+             lightposition=dict(x=100, y=200, z=0),
+             opacity=1.0
+         )
+     ])
+     fig.update_layout(
+         # title='Mesh X Colored by Signed Distance from Mesh Y',
+         scene=dict(
+             aspectmode='data',
+             xaxis=dict(title="X"),
+             yaxis=dict(title="Y"),
+             zaxis=dict(title="Z")),         
+     )
+     fig.update_layout(title='Distance to Nearest Boundary (3D Geodesic)', scene=dict(aspectmode='data'))
+     return fig, cmin, cmax

### 2025-05-05 src/framework_mesh/experimentrunner.py
added velocity

@@ -62,10 +62,20 @@ class ExperimentRunner:
         if self.smoothing_debug:
             print("Smoothing debug is on")
 
+        velocity_cfg = self.cfg.get("velocity", {})
+        self.velocity_enabled = velocity_cfg.get("enabled", False)
+        self.velocity_k = velocity_cfg.get("k", 0)
+        self.beta = velocity_cfg.get("beta", 1)
+
+# ============================================================================================================
+
     def run(self):
         device = self.data_loader.device
         print(f"Running on device: {device}")
         return self.pipeline(self.src_mesh, self.target_meshes, device)
+
+# ============================================================================================================
+
 
     def pipeline(self, src_name: str, tgt_names: List[str], device: torch.device):
         """Pipeline function to run the experiment on a source mesh and a list of target meshes."""
@@ -87,6 +97,8 @@ class ExperimentRunner:
         view_names_used = {}
         step_offset = 0
 
+        prev_verts = None
+        prev_displacement = None
         src_mesh = self.get_mesh(src_name).to(device)
         for tgt_name in tgt_names:
             print(f"Target: {tgt_name}")
@@ -103,6 +115,17 @@ class ExperimentRunner:
 
             edgemap_info = ([tgt_edgemap_info[0]], [tgt_edgemap_info[1]])
 
+            if self.velocity_enabled and prev_displacement is not None:
+                faces = src_mesh[0].faces_packed().to(prev_verts.device)
+                edge_src, edge_dst = build_edge_lists(faces, device=prev_verts.device)
+                smoothed_disp = smooth_displacement_jacobi(prev_displacement, edge_src, edge_dst, k=self.velocity_k)
+                extrapolated_verts = prev_verts[0] + self.beta * smoothed_disp
+                src_mesh = Meshes(verts=[extrapolated_verts], faces=src_mesh.faces_packed())
+
+                mean_disp = (self.beta * smoothed_disp).norm(dim=1).mean().item()
+                if self.wandb:
+                    wandb.log({"velocity/mean_displacement": mean_disp}, step=step_offset)
+
             # Train the mesh transformation
             final_verts = self.train_loop(src_mesh, tgt_mesh,
                                         projmats, edgemap_info,
@@ -113,17 +136,19 @@ class ExperimentRunner:
                                         moment=self.momentum,
                                         device=device)
             
[…]

### 2025-05-05 src/framework_mesh/script.py

@@ -8,9 +8,9 @@ warnings.filterwarnings("ignore")
 def main():
     device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
     device = torch.device("cpu")
-    
-    data_path = "/home/jeffrey/honours/src/framework_mesh/obliqueconfig_gpu.json"
-    exp_config_path = "/home/jeffrey/honours/src/framework_mesh/exp_oblique.json"
+
+    data_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/obliqueconfig_local.json"
+    exp_config_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/exp_oblique.json"
 
     with open(data_path, 'r') as f:
         data_config = json.load(f)
@@ -19,30 +19,48 @@ def main():
         exp_config_base = json.load(f)
 
     mesh_resolutions = [2, 3]
-    smoothing_options = [False, True]
+    grad_smoothing_options = [False, True]
+    velocity_ks = [1, 3]
+    velocity_betas = [0.5, 1.0]
 
     for mesh_res in mesh_resolutions:
-        for smoothing in smoothing_options:
-            data_config['paths']['mesh_res'] = mesh_res
-            dataloader = DataLoader(data_config, device)
+        data_config['paths']['mesh_res'] = mesh_res
+        dataloader = DataLoader(data_config, device)
 
-            exp_config = json.loads(json.dumps(exp_config_base))  # deep copy
-            exp_config.setdefault("gradient", {})["smoothing"] = smoothing
+        for grad_smoothing in grad_smoothing_options:
+            for constrained in ([False, True] if grad_smoothing else [False]):
+                for k in velocity_ks:
+                    for beta in velocity_betas:
+                        # Deep copy experiment config
+                        exp_config = json.loads(json.dumps(exp_config_base))
 
-            if not smoothing:
-                exp_config['name'] = f"nosmooth_res_{mesh_res}"
-                print(f"[RUN] smoothing=False | mesh_res={mesh_res}")
-                runner = ExperimentRunner(exp_config, dataloader)
-                runner.run()
-            else:
-                method = "jacobi"
-                for constrained in [False, True]:
-                    exp_config['gradient']['method'] = method
-                    exp_config['gradient']['constrained'] = constrained
-                    exp_config['name'] = f"smooth_{method}_constrained_{constrained}_res_{mesh_res}"
[…]

### 2025-05-06 src/framework_mesh/experimentrunner.py
fixed sizing of extrpaolated_Verts

@@ -116,11 +116,12 @@ class ExperimentRunner:
             edgemap_info = ([tgt_edgemap_info[0]], [tgt_edgemap_info[1]])
 
             if self.velocity_enabled and prev_displacement is not None:
+                print("Accounting for velocity")
                 faces = src_mesh[0].faces_packed().to(prev_verts.device)
                 edge_src, edge_dst = build_edge_lists(faces, device=prev_verts.device)
                 smoothed_disp = smooth_displacement_jacobi(prev_displacement, edge_src, edge_dst, k=self.velocity_k)
                 extrapolated_verts = prev_verts[0] + self.beta * smoothed_disp
-                src_mesh = Meshes(verts=[extrapolated_verts], faces=src_mesh.faces_packed())
+                src_mesh = Meshes(verts=extrapolated_verts[None, ...], faces=src_mesh.faces_padded())
 
                 mean_disp = (self.beta * smoothed_disp).norm(dim=1).mean().item()
                 if self.wandb:
@@ -309,7 +310,7 @@ def smooth_displacement_jacobi(displacement, edge_src, edge_dst, k=1):
         d_nb = torch.zeros_like(d)
         d_nb.scatter_add_(0, edge_dst[:, None].expand(-1, 3), d[edge_src])
         deg = torch.zeros(V, 1, device=d.device)
-        deg.scatter_add_(0, edge_dst[:, None], torch.ones_like(edge_dst[:, None], dtype=d.dtype))
+        deg.scatter_add_(0, edge_dst[:, None], torch.ones_like(edge_dst[:, None], dtype=deg.dtype))
         d_nb += d
         deg += 1.0
         d = d_nb / deg

### 2025-05-06 src/framework_mesh/dataloader.py
fixed volume constraint in node and made life easier with configs

@@ -18,12 +18,16 @@ class DataLoader:
         self.device = device
         
         # Set up paths and other config settings
-        self.mesh_dir = self.cfg["paths"]["mesh_dir"]
-        self.mesh_res = self.cfg["paths"]["mesh_res"]
-        self.renders_path = self.cfg["paths"]["renders_path"]
-        self.material = self.cfg["paths"]["material"]
+        env = self.cfg.get("env", "local")  # default to "local" if not specified
+        path_cfg = self.cfg["paths"][env]
+
+        # Set up paths and other config settings
+        self.mesh_dir = path_cfg["mesh_dir"]
+        self.renders_path = path_cfg["renders_path"]
+        self.material = path_cfg["material"]
         self.materials_path = os.path.join(self.renders_path, self.material)
-        self.matrices_path = self.cfg["paths"]["matrices_path"]
+        self.matrices_path = path_cfg["matrices_path"]
+        self.mesh_res = self.cfg["paths"]["mesh_res"]  # shared outside "env" dict
         
         self.edgemap_options = {
             mesh["name"]: mesh.get("edgemap_options", {})

### 2025-05-06 src/framework_mesh/script.py

@@ -9,8 +9,8 @@ def main():
     device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
     device = torch.device("cpu")
 
-    data_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/obliqueconfig_local.json"
-    exp_config_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/exp_oblique.json"
+    data_path = "./framework_mesh/obliqueconfig.json"
+    exp_config_path = "./framework_mesh/exp_oblique.json"
 
     with open(data_path, 'r') as f:
         data_config = json.load(f)

### 2025-05-06 src/framework_mesh/dataloader.py
dunno

@@ -19,6 +19,7 @@ class DataLoader:
         
         # Set up paths and other config settings
         env = self.cfg.get("env", "local")  # default to "local" if not specified
+        print(f"env is {env}")
         path_cfg = self.cfg["paths"][env]
 
         # Set up paths and other config settings

### 2025-05-06 src/framework_mesh/script.py
edited script

@@ -9,8 +9,8 @@ def main():
     device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
     device = torch.device("cpu")
 
-    data_path = "./framework_mesh/obliqueconfig.json"
-    exp_config_path = "./framework_mesh/exp_oblique.json"
+    data_path = "./framework_mesh/data_translation.json"
+    exp_config_path = "./framework_mesh/exp_translation.json"
 
     with open(data_path, 'r') as f:
         data_config = json.load(f)

### 2025-05-06 src/framework_mesh/experimentrunner.py
made changes

@@ -125,9 +125,13 @@ class ExperimentRunner:
                 extrapolated_verts = prev_verts[0] + self.beta * smoothed_disp
                 src_mesh = Meshes(verts=extrapolated_verts[None, ...], faces=src_mesh.faces_padded())
 
-                mean_disp = (self.beta * smoothed_disp).norm(dim=1).mean().item()
+                mean_disp = (self.beta * smoothed_disp).mean(dim=0)  # Shape: (3,)
                 if self.wandb:
-                    wandb.log({"velocity/mean_displacement": mean_disp}, step=step_offset)
+                    wandb.log({
+                        "velocity/mean_disp_x": mean_disp[0].item(),
+                        "velocity/mean_disp_y": mean_disp[1].item(),
+                        "velocity/mean_disp_z": mean_disp[2].item(),
+                    }, step=step_offset)
 
             # Train the mesh transformation
             final_verts = self.train_loop(

### 2025-05-06 src/framework_mesh/script.py

@@ -18,7 +18,7 @@ def main():
     with open(exp_config_path, 'r') as f:
         exp_config_base = json.load(f)
 
-    mesh_resolutions = [2, 3]
+    mesh_resolutions = [3]
     grad_smoothing_options = [True]
     velocity_ks = [1]
     velocity_betas = [1.0]

### 2025-05-07 src/framework_mesh/chamfer.py
updated chamfer to sample from surface

@@ -45,7 +45,7 @@ def get_boundary(projected_pts: torch.Tensor, alpha: float = 12.0):
     return boundary_pts, b_mask
 
 class PyTorchChamferLoss(nn.Module):
-    def __init__(self, src: Meshes, tgt: Meshes, projmatrices, edgemap_info, boundary_mask=None):
+    def __init__(self, src: Meshes, tgt: Meshes, projmatrices, edgemap_info, boundary_mask=None, doublesided=False):
         super().__init__()
         self.src = src  # (B meshes)
         self.tgt = tgt  # (B meshes)
@@ -53,6 +53,7 @@ class PyTorchChamferLoss(nn.Module):
         self.edgemaps = edgemap_info[0] # (P, max_Ni, 2)
         self.edgemaps_len = edgemap_info[1] # (P,)
         self.boundary_mask = boundary_mask    # keep the reference
+        self.doublesided = doublesided
 
     def project_vertices(self, vertices):
         """
@@ -119,7 +120,7 @@ class PyTorchChamferLoss(nn.Module):
                         y_lengths=self.edgemaps_len[b].long(),
                         batch_reduction="mean",
                         point_reduction="mean",
-                        single_directional=True)
+                        single_directional=not self.doublesided)
             chamfer_loss[b] = res.sum()
 
         return chamfer_loss.double()

### 2025-05-07 src/framework_mesh/experimentrunner.py

@@ -187,7 +187,7 @@ class ExperimentRunner:
             )
             verts.register_hook(hook)
 
-        chamfer_loss = PyTorchChamferLoss(src, tgt, projmats, edgemap_info, boundary_mask=boundary_mask)
+        chamfer_loss = PyTorchChamferLoss(src, tgt, projmats, edgemap_info, boundary_mask=boundary_mask, doublesided = self.cfg["chamfer"]["doublesided"])
         optimiser = torch.optim.SGD([verts], lr=lr, momentum=moment)
         a,b = edgemap_info
         a,b = a[0], b[0]
@@ -219,15 +219,17 @@ class ExperimentRunner:
             loss.backward()
             optimiser.step()
 
-            pbar.set_description(f"Loss: {loss.item()*10:.4f}")
+            pbar.set_description(f"Loss: {loss.item():.4f}")
 
+
+            tmp_mesh = Meshes(verts=projverts.detach(), faces=src[0].faces_packed().unsqueeze(0))
             # log
             if self.wandb:
                 wandb.log(
                     data = {
                     "outer/chamfer": loss * 10,
                     "outer/vol": calculate_volume(projverts[0], src[0].faces_packed()),
-                    "outer/gt/chamfer": chamfer_gt(projverts, src, tgt)[0] * 10,
+                    "outer/gt/chamfer": chamfer_gt(tmp_mesh, tgt)[0],
                     "outer/gt/iou": iou_gt(projverts, src, tgt)[0]
                     },
                     step = i + step_offset
@@ -235,13 +237,12 @@ class ExperimentRunner:
 
             if self.verbose:
                 print(f"{i:4d} Loss: {colour}{loss.item():.3f}{bcolors.ENDC} Volume: {calculate_volume(projverts[0], src[0].faces_packed()):.3f}")
-                print(f"GT Chamfer: [{', '.join(f'{x:.3f}' for x in chamfer_gt(projverts, src, tgt))}] "
+                print(f"GT Chamfer: [{', '.join(f'{x:.3f}' for x in chamfer_gt(tmp_mesh, tgt))}] "
                     f"GT IoU: [{', '.join(f'{x:.3f}' for x in iou_gt(projverts, src, tgt))}]")
             def should_log(i):
                 return i < 50 or (i % self.vis_freq == self.vis_freq-1)  # True at i = 1, 2, 4, 8, 16, ...
             if self.vis_enabled and should_log(i):
[…]

### 2025-05-07 src/framework_mesh/functions.py

@@ -1,6 +1,7 @@
 import torch
 from pytorch3d.structures import Meshes
 from pytorch3d.loss import chamfer_distance
+from pytorch3d.ops import sample_points_from_meshes
 import trimesh
 import numpy as np
 
@@ -92,15 +93,26 @@ def volume_constraint_grad(x, faces, device=torch.device("cpu")):
     analytical_grad = grad_verts.flatten()
     return analytical_grad 
 
-def chamfer_gt(mesh, src:Meshes, tgt:Meshes):
-    res,_ = chamfer_distance(x=mesh.detach().float(), 
-                             y=tgt.verts_padded().float().detach(),
-                             x_lengths=src.num_verts_per_mesh(),
-                             y_lengths=tgt.num_verts_per_mesh(),
-                             batch_reduction=None,
-                             point_reduction="mean")
-    # print("Chamfer", res.size())
-    return res.tolist() # (B,)
+def chamfer_gt(src:Meshes, tgt:Meshes, num_samples=10000):
+    # res,_ = chamfer_distance(x=mesh.detach().float(), 
+    #                          y=tgt.verts_padded().float().detach(),
+    #                          x_lengths=src.num_verts_per_mesh(),
+    #                          y_lengths=tgt.num_verts_per_mesh(),
+    #                          batch_reduction=None,
+    #                          point_reduction="mean")
+    # # print("Chamfer", res.size())
+    # return res.tolist() # (B,)
+    src_pts = sample_points_from_meshes(src, num_samples=num_samples)
+    tgt_pts = sample_points_from_meshes(tgt, num_samples=num_samples)
+
+    res, _ = chamfer_distance(
+        x=src_pts, 
+        y=tgt_pts, 
+        batch_reduction=None,  # per-mesh loss
+        point_reduction="mean"
+    )
+    return res.tolist()  # (B,)
+
 
 
 def sse_gt(mesh, src:Meshes, tgt:Meshes):

### 2025-05-07 src/framework_mesh/script.py

@@ -5,62 +5,96 @@ from .experimentrunner import ExperimentRunner
 import warnings
 warnings.filterwarnings("ignore")
 
-def main():
+def run_with_configs(data_path, exp_path, mesh_res_list, enable_doublesided=False):
     device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
     device = torch.device("cpu")
-
-    data_path = "./framework_mesh/obliqueconfig.json"
-    exp_config_path = "./framework_mesh/exp_oblique.json"
 
     with open(data_path, 'r') as f:
         data_config = json.load(f)
 
-    with open(exp_config_path, 'r') as f:
+    with open(exp_path, 'r') as f:
         exp_config_base = json.load(f)
 
-    mesh_resolutions = [3]
-    grad_smoothing_options = [True]
-    velocity_ks = [1]
-    velocity_betas = [1.0]
+    grad_smoothing = True
+    velocity_k = 1
+    velocity_beta = 1.0
 
-    for mesh_res in mesh_resolutions:
+    for mesh_res in mesh_res_list:
         data_config['paths']['mesh_res'] = mesh_res
         dataloader = DataLoader(data_config, device)
 
-        for grad_smoothing in grad_smoothing_options:
-            for constrained in ([False, True] if grad_smoothing else [False]):
-                for k in velocity_ks:
-                    for beta in velocity_betas:
-                        # Deep copy experiment config
-                        exp_config = json.loads(json.dumps(exp_config_base))
+        for constrained in [False, True] if mesh_res != 3 else [False]:  # for res=3 only run constrained=False unless explicitly stated
+            exp_config = json.loads(json.dumps(exp_config_base))
+            exp_config["velocity"] = {
+                "enabled": True,
+                "k": velocity_k,
+                "beta": velocity_beta
+            }
 
-                        # Velocity (displacement) smoothing
-                        exp_config["velocity"] = {
-                            "enabled": True,
-                            "k": k,
-                            "beta": beta
-                        }
+            exp_config["gradient"] = {
+                "smoothing": grad_smoothing,
+                "method": "jacobi",
+                "k": 5,
+                "constrained": constrained,
+                "debug": False
+            }
 
-                        # Gradient smoothing
-                        exp_config["gradient"] = {
-                            "smoothing": grad_smoothing,
-                            "method": "jacobi",
[…]

### 2025-05-07 src/framework_mesh/functions.py
updated chamfer gt

@@ -102,8 +102,8 @@ def chamfer_gt(src:Meshes, tgt:Meshes, num_samples=10000):
     #                          point_reduction="mean")
     # # print("Chamfer", res.size())
     # return res.tolist() # (B,)
-    src_pts = sample_points_from_meshes(src, num_samples=num_samples)
-    tgt_pts = sample_points_from_meshes(tgt, num_samples=num_samples)
+    src_pts = sample_points_from_meshes(src, num_samples=num_samples).float()
+    tgt_pts = sample_points_from_meshes(tgt, num_samples=num_samples).float()
 
     res, _ = chamfer_distance(
         x=src_pts, 

### 2025-05-07 src/framework_mesh/script.py

@@ -5,9 +5,13 @@ from .experimentrunner import ExperimentRunner
 import warnings
 warnings.filterwarnings("ignore")
 
-def run_with_configs(data_path, exp_path, mesh_res_list, enable_doublesided=False):
+
+def main():
     device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
     device = torch.device("cpu")
+
+    data_path = "./framework_mesh/data_noground.json"
+    exp_path = "./framework_mesh/exp_oblique.json"
 
     with open(data_path, 'r') as f:
         data_config = json.load(f)
@@ -15,86 +19,52 @@ def run_with_configs(data_path, exp_path, mesh_res_list, enable_doublesided=Fals
     with open(exp_path, 'r') as f:
         exp_config_base = json.load(f)
 
-    grad_smoothing = True
+    mesh_res = 2
     velocity_k = 1
     velocity_beta = 1.0
 
-    for mesh_res in mesh_res_list:
-        data_config['paths']['mesh_res'] = mesh_res
-        dataloader = DataLoader(data_config, device)
-
-        for constrained in [False, True] if mesh_res != 3 else [False]:  # for res=3 only run constrained=False unless explicitly stated
-            exp_config = json.loads(json.dumps(exp_config_base))
-            exp_config["velocity"] = {
-                "enabled": True,
-                "k": velocity_k,
-                "beta": velocity_beta
-            }
-
-            exp_config["gradient"] = {
-                "smoothing": grad_smoothing,
-                "method": "jacobi",
-                "k": 5,
-                "constrained": constrained,
-                "debug": False
-            }
-
-            # Enable double-sided Chamfer if requested
-            if enable_doublesided:
-                exp_config["chamfer"] = {
-                    "doublesided": True
-                }
-
-            name = f"vbeta_{velocity_beta}_vk_{velocity_k}_gradsmooth_{grad_smoothing}_constrained_{constrained}_res_{mesh_res}"
-            exp_config["name"] = name
-
-            print(f"[RUN] {name}")
-            runner = ExperimentRunner(exp_config, dataloader)
-            runner.run()
-
-def main():
-    # First set: obliqueconfig, res=2
-    run_with_configs(
-        "./framework_mesh/obliqueconfig.json",
-        "./framework_mesh/exp_oblique.json",
-        mesh_res_list=[2],
-        enable_doublesided=False
-    )
-
[…]

### 2025-05-07 src/framework_mesh/functions.py
fixed chamfer gt float dtype

@@ -93,25 +93,26 @@ def volume_constraint_grad(x, faces, device=torch.device("cpu")):
     analytical_grad = grad_verts.flatten()
     return analytical_grad 
 
-def chamfer_gt(src:Meshes, tgt:Meshes, num_samples=10000):
-    # res,_ = chamfer_distance(x=mesh.detach().float(), 
-    #                          y=tgt.verts_padded().float().detach(),
-    #                          x_lengths=src.num_verts_per_mesh(),
-    #                          y_lengths=tgt.num_verts_per_mesh(),
-    #                          batch_reduction=None,
-    #                          point_reduction="mean")
-    # # print("Chamfer", res.size())
-    # return res.tolist() # (B,)
-    src_pts = sample_points_from_meshes(src, num_samples=num_samples).float()
-    tgt_pts = sample_points_from_meshes(tgt, num_samples=num_samples).float()
+def chamfer_gt(src: Meshes, tgt: Meshes, num_samples=10000):
+    # Convert to float32 manually
+    def cast_to_float32(mesh):
+        verts = [v.to(torch.float32) for v in mesh.verts_list()]
+        faces = mesh.faces_list()
+        return Meshes(verts=verts, faces=faces)
+
+    src = cast_to_float32(src)
+    tgt = cast_to_float32(tgt)
+
+    src_pts = sample_points_from_meshes(src, num_samples=num_samples)
+    tgt_pts = sample_points_from_meshes(tgt, num_samples=num_samples)
 
     res, _ = chamfer_distance(
-        x=src_pts, 
-        y=tgt_pts, 
-        batch_reduction=None,  # per-mesh loss
+        x=src_pts,
+        y=tgt_pts,
+        batch_reduction=None,
         point_reduction="mean"
     )
-    return res.tolist()  # (B,)
+    return res.tolist()
 
 
 

### 2025-05-07 src/framework_mesh/script.py

@@ -40,7 +40,7 @@ def main():
         exp_config["gradient"] = {
             "smoothing": True,
             "method": "jacobi",
-            "k": 4,
+            "k": 5,
             "constrained": constrained,
             "debug": False
         }

### 2025-05-07 src/framework_mesh/experimentrunner.py
flowchart code

@@ -197,21 +197,37 @@ class ExperimentRunner:
                 print(f"{i:4d} Loss: {colour}{loss.item():.3f}{bcolors.ENDC} Volume: {calculate_volume(projverts[0], src[0].faces_packed()):.3f}")
                 print(f"GT Chamfer: [{', '.join(f'{x:.3f}' for x in chamfer_gt(projverts, src, tgt))}] "
                     f"GT IoU: [{', '.join(f'{x:.3f}' for x in iou_gt(projverts, src, tgt))}]")
-            if self.vis_enabled and i % self.vis_freq == self.vis_freq-1:
-                projectionplot = plot_projections(projverts.detach().squeeze().double(), gt_projmats, gt_edgemap_info)
-                tmp_mesh = Meshes(verts=projverts.detach(), faces=src[0].faces_packed().unsqueeze(0))
-                heatmap,cmin,cmax = create_heatmap(tmp_mesh, tgt[0], cmin, cmax)
+            if self.vis_enabled and i % self.vis_freq == self.vis_freq - 1:
+                faces = src[0].faces_packed().unsqueeze(0)
+                mesh_proj = Meshes(verts=projverts.detach(), faces=faces)
+                mesh_verts = Meshes(verts=verts.detach(), faces=faces)
+
+                # Heatmaps
+                heatmap_proj, cmin, cmax = create_heatmap(mesh_proj, tgt[0], cmin, cmax)
+                heatmap_verts, _, _ = create_heatmap(mesh_verts, tgt[0], cmin, cmax)
+
+                # Volume logging
+                vol_proj = calculate_volume(projverts[0], src[0].faces_packed())
+                vol_verts = calculate_volume(verts[0], src[0].faces_packed())
+
+                # 2D scatter projection
+                camera_path = "/Users/jeffreyliang/Documents/Honours/Blender/blurred/cameras/Cam_Ground_3_P.npy"
+                projectionplot = plot_projverts_scatter2d(projverts[0], camera_path, f"Step {i}")
+
+                # Optional: Boundary heatmap if still needed
                 boundary_fig, bmin, bmax = compute_boundary_distance_heatmap(
-                    tmp_mesh, boundary_mask, D_all, bmin, bmax
+                    mesh_proj, boundary_mask, D_all, bmin, bmax
                 )
 
                 if self.wandb:
                     wandb.log({
-                        "plt/projections": wandb.Image(projectionplot),
-                        "plt/heatmap": wandb.Plotly(heatmap),
[…]

### 2025-05-07 src/framework_mesh/script.py

@@ -11,8 +11,8 @@ def main():
     device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
     device = torch.device("cpu")
     
-    data_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/skyconfig_local.json"
-    exp_config_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/exp_pipeline.json"
+    data_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/skyconfig_balloon.json"
+    exp_config_path = "/Users/jeffreyliang/Documents/Honours/honours/src/framework_mesh/exp_gradients.json"
 
     with open(data_path, 'r') as f:
         data_config = json.load(f)
@@ -20,27 +20,9 @@ def main():
     with open(exp_config_path, 'r') as f:
         exp_config = json.load(f)
 
-    methods = ["jacobi"]
-    constrained_options = [False]
-    mesh_resolutions = [2]
-
-    for mesh_res in mesh_resolutions:
-        # Update mesh resolution
-        data_config['paths']['mesh_res'] = mesh_res
-        dataloader = DataLoader(data_config, device)
-
-        for method in methods:
-            for constrained in constrained_options:
-                # Update gradient method and constraint setting
-                exp_config['gradient']['method'] = method
-                exp_config['gradient']['constrained'] = constrained
-
-                # Update experiment name to be descriptive
-                exp_config['name'] = f"jaconlyboundary_pipeline_{method}_constrained_{constrained}_res_{mesh_res}"
-
-                # Create runner and run
-                runner = ExperimentRunner(exp_config, dataloader)
-                runner.run()
+    dataloader = DataLoader(data_config, device)
+    runner = ExperimentRunner(exp_config, dataloader)
+    runner.run()
 
 if __name__ == "__main__":
     main()

### 2025-05-07 src/framework_mesh/utils.py

@@ -435,3 +435,35 @@ def compute_boundary_distance_heatmap(src: Meshes, boundary_mask, D_all, cmin=No
     )
     fig.update_layout(title='Distance to Nearest Boundary (3D Geodesic)', scene=dict(aspectmode='data'))
     return fig, cmin, cmax
+
+
+def project_vertices(verts3d, camera_matrix_path):
+    P = torch.tensor(np.load(camera_matrix_path), dtype=verts3d.dtype, device=verts3d.device)
+    ones = torch.ones((verts3d.shape[0], 1), dtype=verts3d.dtype, device=verts3d.device)
+    verts_h = torch.cat([verts3d, ones], dim=1)
+    proj_h = (P @ verts_h.T).T
+    proj_2d = proj_h[:, :2] / proj_h[:, 2:]
+    return proj_2d.detach().cpu().numpy() 
+
+def plot_projverts_scatter2d(projverts, camera_matrix_path, title, save_path=None):
+    proj_2d = project_vertices(projverts, camera_matrix_path)
+
+    x = proj_2d[:, 0]
+    y = proj_2d[:, 1]
+    fig = go.Figure(data=go.Scatter(x=x, y=-y, mode='markers',
+                                    marker=dict(size=3), name='Projected verts'))
+
+    fig.update_layout(
+        title=title,
+        xaxis=dict(visible=False, scaleanchor='y'),
+        yaxis=dict(visible=False, autorange='reversed'),
+        showlegend=False,
+        plot_bgcolor='rgba(0,0,0,0)',
+        paper_bgcolor='rgba(0,0,0,0)',
+        width=512, height=512
+    )
+
+    if save_path:
+        fig.write_json(save_path)
+
+    return fig
